### 2021-02-09 evaluation/template-plot-forecasts.Rmd

```{r prediction-plots, echo = FALSE, results='asis', fig.width = 10, fig.height=20}

for (forecast_date in forecast_dates) {
  cat(paste0("\n\n## ", forecast_date), "{.tabset}\n\n")
  for (target_type in target_types) {
    cat("\n\n###", target_type, "\n\n")

    filter_both <- list(paste0("target_type %in% '", target_type, "'"))
    filter_truth <- list(paste0("target_end_date > '", as.Date(forecast_date) - 7 * 10, "'"), 
                         paste0("target_end_date <= '", as.Date(forecast_date) + 7 * 4, "'"))
    filter_forecasts <- list(paste0("forecast_date == '", as.Date(forecast_date), "'"))

    plot <- scoringutils::plot_predictions(data,
                                           x = "target_end_date",
                                           filter_both = filter_both,
                                           filter_truth = filter_truth,
                                           filter_forecasts = filter_forecasts,
                                           facet_wrap_or_grid = "grid",
                                           facet_formula = model ~ location_name,
                                           allow_truth_without_pred = FALSE,
                                           scales = "free") +
      ggplot2::ggtitle(paste0("Predictions for incident ", target_type,  "s")) + 
      ggplot2::theme(legend.position = "bottom")

    print(plot)
  }
}

```