### 2021-02-09 evaluation/template-wis-components.Rmd

```{r wis-component-plot, results = 'asis', fig.width=10}
scores <- scoringutils::eval_forecasts(data,
                                       summarise_by = c("model", "location_name", "target_type"), 
                                       compute_relative_skill = FALSE)
plot <- scoringutils::wis_components(scores, 
                                     facet_wrap_or_grid = "grid",
                                     relative_contributions = TRUE,
                                     facet_formula = . ~ target_type + location_name) + 
  ggplot2::coord_flip() + 
  ggplot2::theme(legend.position = "bottom")

print(plot)
```