### 2021-02-11 evaluation/template-ranking-over-time.Rmd

```{r ranking-time, results = 'asis', fig.width=12, fig.height=10}
scores <- scoringutils::eval_forecasts(data,
                                         summarise_by = c("model", 
                                                          "location_name", 
                                                          "target_type",
                                                          "forecast_date"), 
                                       compute_relative_skill = FALSE)

ranked_scores <- scores %>%
  dplyr::mutate(forecast_date = as.Date(forecast_date)) %>%
  dplyr::group_by(forecast_date, location_name, target_type) %>%
  dplyr::mutate(num_forecasts = dplyr::n(), 
                unstandardised_rank = rank(interval_score, ties.method = "average",
                                           na.last = NA), 
                rank = round((1 - (unstandardised_rank - 1) / (num_forecasts - 1)) * 100)) %>%
  dplyr::ungroup()

plot <- score_heatmap(ranked_scores, metric = "rank", 
                      x = "forecast_date",
                      facet_wrap_or_grid = "grid",
                      facet_formula = location_name ~ target_type) + 
  ggplot2::theme(legend.position = "bottom") +
    ggplot2::scale_fill_gradient(high = "white", low = "coral")

print(plot)

```