### 2021-02-13 crowd-forecast/email-templates/email-template.Rmd
update performance board and add code to send emails

---
  output: 
    blastula::blastula_email
  title: "Crowd Forecast Updates"
  fig_caption: false
---

```{r echo=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)

library(here)
library(scoringutils)
library(dplyr)
library(data.table)
library(covid.german.forecasts)
library(kableExtra)
```

```{r load-data, echo=FALSE, message = FALSE, warning = FALSE}
# load all data ----------------------------------------------------------------
root_dir <- here::here("crowd-forecast", "processed-forecast-data")
file_paths_forecast <- here::here(root_dir, list.files(root_dir))

prediction_data <- purrr::map_dfr(file_paths_forecast, 
                                  .f = function(x) {
                                    data <- data.table::fread(x) %>%
                                      dplyr::mutate(target_end_date = as.Date(target_end_date), 
                                                    submission_date = as.Date(submission_date), 
                                                    forecast_date = as.Date(forecast_date))
                                  }) %>%
  dplyr::mutate(target_type = ifelse(grepl("death", target), "death", "case")) %>%
  dplyr::rename(prediction = value) %>%
  dplyr::mutate(forecast_date = as.Date(submission_date)) %>%
  dplyr::rename(model = board_name) %>%
  dplyr::filter(type == "quantile", 
                location_name %in% c("Germany", "Poland")) %>%
  dplyr::select(location, location_name, forecast_date, quantile, prediction, model, target_end_date, horizon, target, target_type)

# filter forecasters such that only those are kept that have submitted forecasts
# in the last four weeks
keep_fc <- prediction_data %>%
  dplyr::filter(forecast_date > (Sys.Date() - 4 * 7)) %>%
  dplyr::pull(model) %>%
  unique()

prediction_data <- dplyr::filter(prediction_data, 
                                 model %in% keep_fc)

files <- list.files(here::here("data-raw"))
file_paths <- here::here("data-raw", files[grepl("weekly-incident", files)])
names(file_paths) <- c("case", "death")

truth_data <- purrr::map_dfr(file_paths, readr::read_csv, .id = "target_type") %>%
  dplyr::rename(true_value = value) %>%
  dplyr::mutate(target_end_date = as.Date(target_end_date)) %>%
  dplyr::arrange(location, target_type, target_end_date) %>%
  dplyr::filter(location_name %in% c("Germany", "Poland"))

data <- scoringutils::merge_pred_and_obs(prediction_data, truth_data, 
                                         join = "full")

```

### Dear `r name`,

Good to have you as a forecaster! You're receiving this message because you gave us your email address when you registered for the [Epiforecasts Crowd Forecast App](https://cmmid-lshtm.shinyapps.io/crowd-forecast/).

It has been a long time, but finally we managed to get this email newsletter running! From now on you will receive a weekly email with

- announcements and other information
- a reminder to submit a forecast
- some information about last weeks forecast and your performance

If you don't want to receive future emails simply write me a quick message. We're working on a way that allows you to unsubscribe from within the app, but that is work in progress. Baby steps, but we're getting there.

If you want to lean more about this project, have a look at our [github repo](https://github.com/epiforecasts/covid.german.forecasts). If you want to learn more about the German and Polish Forecast Hub, where we submit the forecasts to, have a look [here](https://kitmetricslab.github.io/forecasthub/forecast).

<br>

### Reminder for the weekly forecasts

We **update our data every Saturday at 6.30pm CET**. You can then make a forecast until **2pm CET (1pm UK time) on Tuesday** using our [forecast app](https://cmmid-lshtm.shinyapps.io/crowd-forecast/). We would be very happy to have a forecast from you this week!

<br>

### Forecast and Performance overview

But now, let's have a look at last weeks forecasts and performance!

In the following we put together a short overview for you. You can find more detailed information on our performance board [here](https://epiforecasts.io/covid.german.forecasts).

*Here are the latest EpiExpert ensemble forecasts and next to them your own forecasts
if you have submitted any.*

```{r plot-forecasts, echo=FALSE, out.width="49%", fig.show="hold", fig.cap = "Plot of the forecasts"}

forecast_date <- covid.german.forecasts::latest_weekday(Sys.Date())

filter_truth <- list('target_end_date >= Sys.Date() - 14 * 7')
models <- c("anonymous_Human", "EpiExpert-ensemble")
filter_forecasts <- list("model %in% 'EpiExpert-ensemble'", 
                         paste0("forecast_date == '", forecast_date, "'"))

plot_predictions(data = data, 
                 x = "target_end_date",
                 filter_truth = filter_truth, 
                 filter_forecasts = filter_forecasts,
                 facet_formula = target_type ~ location_name, 
                 facet_wrap_or_grid = "grid") + 
  ggplot2::labs(x = "Date", title = "Ensemble Forecasts") + 
  ggplot2::theme(legend.position = "bottom")

forecast_date <- covid.german.forecasts::latest_weekday(Sys.Date())

filter_truth <- list('target_end_date >= Sys.Date() - 14 * 7')
filter_forecasts <- list(paste0("model %in% '", board_name, "'"), 
                         paste0("forecast_date == '", forecast_date, "'"))

p <- plot_predictions(data = data, 
                 x = "target_end_date",
                 filter_truth = filter_truth, 
                 filter_forecasts = filter_forecasts,
                 facet_formula = target_type ~ location_name, 
                 facet_wrap_or_grid = "grid") 
if (!is.null(p)) {
  p <- p + 
  ggplot2::labs(x = "Date", title = "Your Forecasts") + 
  ggplot2::theme(legend.position = "bottom")
  print(p)
}

```

<br>

*Here is the list with the current top 5 forecasters (based on the last 5 weeks): *

```{r rank-table, echo=FALSE, fig.align="center"}
cut_off <- covid.german.forecasts::latest_weekday(Sys.Date()) - 5 * 7
df <- data %>%
  filter(forecast_date >= cut_off)

table <- eval_forecasts(data, summarise_by = c("model"), 
                        compute_relative_skill = TRUE)

table %>%
    arrange(relative_skill) %>%
    dplyr::select(-scaled_rel_skill) %>%
    mutate_if(is.numeric, round, 2) %>%
    dplyr::rename(wis = interval_score,
                  underpred = underprediction,
                  overpred = overprediction,
                  cvrage_dev = coverage_deviation,
                  rel_skill = relative_skill) %>%
  select(model, rel_skill) %>%
  head(5) %>%
  knitr::kable(format = "html") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)

```

<br>

*And here is a plot with the rank by date and target of all active forecasters over time:*

```{r rank-over-time, echo=FALSE, fig.align = "center", fig.width=10, fig.height=8, out.width="80%", fig.cap = "Ranking over time"}
scores <- scoringutils::eval_forecasts(data,
                                         summarise_by = c("model", 
                                                          "location_name", 
                                                          "target_type",
                                                          "forecast_date"), 
                                       compute_relative_skill = FALSE)

ranked_scores <- scores %>%
  dplyr::mutate(forecast_date = as.Date(forecast_date)) %>%
  dplyr::group_by(forecast_date, location_name, target_type) %>%
  dplyr::mutate(num_forecasts = dplyr::n(), 
                rank = rank(interval_score, ties.method = "average",
                                           na.last = NA), 
                standard_rank = round((1 - (rank - 1) / (num_forecasts - 1)) * 100)) %>%
  dplyr::ungroup()

plot <- score_heatmap(ranked_scores, metric = "rank", 
                      x = "forecast_date",
                      facet_wrap_or_grid = "grid",
                      facet_formula = location_name ~ target_type) + 
  ggplot2::theme(legend.position = "bottom") +
    ggplot2::scale_fill_gradient(low = "white", high = "coral")

print(plot)
```

<br>

If you have any questions or feedback, please don't hesitate to reach out. You can drop us an email or submit an issue to our [github repo](https://github.com/epiforecasts/covid.german.forecasts)

Thank you very much for your support and participation. We very much appreciate it!

Best wishes,

Nikos for the Epiforecasts team