### 2021-02-15 evaluation/template-scores-and-truth-time.Rmd

```{r scores-and-truth1, results = 'asis', fig.width=10, fig.height = 5}

library(ggplot2)
library(dplyr)
library(scoringutils)

scores <- eval_forecasts(data, 
                         summarise_by = c("model", "target_type", 
                                          "location_name", "forecast_date"), 
                         compute_relative_skill = FALSE)

p1 <- data %>%
  select(target_end_date, true_value, location_name, target_type) %>%
  filter(target_end_date > "2020-10-12", 
         !is.na(true_value)) %>%
  unique() %>%
  ggplot(aes(x = target_end_date, y = true_value)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  theme(legend.position = "bottom") + 
  labs(y = "true observed values", x = "date", title = "Observed data")

cat("\n\n")
print(p1)
cat("\n\n")
```

```{r scores-and-truth2, results = 'asis', fig.width=10, fig.height = 7}

p2 <- scores %>%
  ggplot(aes(x = forecast_date, y = interval_score, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  theme(legend.position = "bottom") +
  labs(y = "weighted interval score", x = "date", title = "Forecaster Scores")

cat("\n\n")
print(p2)
cat("\n\n")
```