### 2021-02-22 crowd-forecast/email-templates/email-template.Rmd
update email template

---
  output: 
    blastula::blastula_email
  title: "Crowd Forecast Updates"
  fig_caption: false
---

```{r echo=FALSE, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)

library(here)
library(scoringutils)
library(dplyr)
library(data.table)
library(covid.german.forecasts)
library(kableExtra)
```

```{r load-data, echo=FALSE, message = FALSE, warning = FALSE}
# load all data ----------------------------------------------------------------
root_dir <- here::here("crowd-forecast", "processed-forecast-data")
file_paths_forecast <- here::here(root_dir, list.files(root_dir))

prediction_data <- purrr::map_dfr(file_paths_forecast, 
                                  .f = function(x) {
                                    data <- data.table::fread(x) %>%
                                      dplyr::mutate(target_end_date = as.Date(target_end_date), 
                                                    submission_date = as.Date(submission_date), 
                                                    forecast_date = as.Date(forecast_date))
                                  }) %>%
  dplyr::mutate(target_type = ifelse(grepl("death", target), "death", "case")) %>%
  dplyr::rename(prediction = value) %>%
  dplyr::mutate(forecast_date = as.Date(submission_date)) %>%
  dplyr::rename(model = board_name) %>%
  dplyr::filter(type == "quantile", 
                location_name %in% c("Germany", "Poland")) %>%
  dplyr::select(location, location_name, forecast_date, quantile, prediction, model, target_end_date, horizon, target, target_type)

# filter forecasters such that only those are kept that have submitted forecasts
# in the last four weeks
keep_fc <- prediction_data %>%
  dplyr::filter(forecast_date > (Sys.Date() - 4 * 7)) %>%
  dplyr::pull(model) %>%
  unique()

prediction_data <- dplyr::filter(prediction_data, 
                                 model %in% keep_fc)

files <- list.files(here::here("data-raw"))
file_paths <- here::here("data-raw", files[grepl("weekly-incident", files)])
names(file_paths) <- c("case", "death")

truth_data <- purrr::map_dfr(file_paths, readr::read_csv, .id = "target_type") %>%
  dplyr::rename(true_value = value) %>%
  dplyr::mutate(target_end_date = as.Date(target_end_date)) %>%
  dplyr::arrange(location, target_type, target_end_date) %>%
  dplyr::filter(location_name %in% c("Germany", "Poland"))

data <- scoringutils::merge_pred_and_obs(prediction_data, truth_data, 
                                         join = "full")

```

### Dear `r name`,

Good to have you as a forecaster! Here is this weeks reminder and an update about what has happened lately. You're receiving this message because you gave us your email address when you registered for the [Epiforecasts Crowd Forecast App](https://cmmid-lshtm.shinyapps.io/crowd-forecast/).

If you want to lean more about this project, have a look at our [github repo](https://github.com/epiforecasts/covid.german.forecasts). If you want to learn more about the German and Polish Forecast Hub, where we submit the forecasts to, have a look [here](https://kitmetricslab.github.io/forecasthub/forecast).

<br>

### Reminder for the weekly forecasts

We **update our data every Saturday at 6.30pm CET**. You can then make a forecast until **2pm CET (1pm UK time) on Tuesday** using our [forecast app](https://cmmid-lshtm.shinyapps.io/crowd-forecast/). We would be very happy to have a forecast from you this week!

<br>

### Forecast and Performance overview

But now, let's have a look at last weeks forecasts and performance!

In the following we put together a short overview for you. You can find more detailed information on our performance board [here](https://epiforecasts.io/covid.german.forecasts).

*Here are the latest EpiExpert ensemble forecasts and next to them your own forecasts
if you have submitted any.*

```{r plot-forecasts, echo=FALSE, out.width="49%", fig.show="hold"}

if (weekdays(Sys.Date()) == "Monday") {
  forecast_date <- covid.german.forecasts::latest_weekday(Sys.Date() - 7)
} else {
  forecast_date <- covid.german.forecasts::latest_weekday(Sys.Date())
}

filter_truth <- list('target_end_date >= Sys.Date() - 14 * 7')
filter_forecasts <- list("model %in% 'EpiExpert-ensemble'", 
                         paste0("forecast_date == '", forecast_date, "'"))

plot_predictions(data = data, 
                 x = "target_end_date",
                 filter_truth = filter_truth, 
                 filter_forecasts = filter_forecasts,
                 facet_formula = target_type ~ location_name, 
                 facet_wrap_or_grid = "grid") + 
  ggplot2::labs(x = "Date", title = "Ensemble Forecasts") + 
  ggplot2::theme(legend.position = "bottom")

filter_truth <- list('target_end_date >= Sys.Date() - 14 * 7')
filter_forecasts <- list(paste0("model %in% '", board_name, "'"), 
                         paste0("forecast_date == '", forecast_date, "'"))

p <- plot_predictions(data = data, 
                 x = "target_end_date",
                 filter_truth = filter_truth, 
                 filter_forecasts = filter_forecasts,
                 facet_formula = target_type ~ location_name, 
                 facet_wrap_or_grid = "grid") 
if (!is.null(p)) {
  p <- p + 
  ggplot2::labs(x = "Date", title = "Your Forecasts") + 
  ggplot2::theme(legend.position = "bottom")
  print(p)
}

```

<br>

*Here is the list with the current top 5 forecasters (based on the last 5 weeks): *

```{r rank-table, echo=FALSE, fig.align="center"}
cut_off <- covid.german.forecasts::latest_weekday(Sys.Date()) - 5 * 7
df <- data %>%
  filter(forecast_date >= cut_off)

table <- eval_forecasts(data, summarise_by = c("model"), 
                        compute_relative_skill = TRUE)

table %>%
    arrange(relative_skill) %>%
    dplyr::select(-scaled_rel_skill) %>%
    mutate_if(is.numeric, round, 2) %>%
    dplyr::rename(wis = interval_score,
                  underpred = underprediction,
                  overpred = overprediction,
                  cvrage_dev = coverage_deviation,
                  rel_skill = relative_skill) %>%
  select(model, rel_skill) %>%
  head(5) %>%
  knitr::kable(format = "html") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)

```

<br>

*And here is a plot with the rank by date and target of all active forecasters over time:*

```{r rank-over-time, echo=FALSE, fig.align = "center", fig.width=10, fig.height=8, out.width="80%"}
scores <- scoringutils::eval_forecasts(data,
                                         summarise_by = c("model", 
                                                          "location_name", 
                                                          "target_type",
                                                          "forecast_date"), 
                                       compute_relative_skill = FALSE)

ranked_scores <- scores %>%
  dplyr::mutate(forecast_date = as.Date(forecast_date)) %>%
  dplyr::group_by(forecast_date, location_name, target_type) %>%
  dplyr::mutate(num_forecasts = dplyr::n(), 
                rank = rank(interval_score, ties.method = "average",
                                           na.last = NA), 
                standard_rank = round((1 - (rank - 1) / (num_forecasts - 1)) * 100)) %>%
  dplyr::ungroup()

plot <- score_heatmap(ranked_scores, metric = "rank", 
                      x = "forecast_date",
                      facet_wrap_or_grid = "grid",
                      facet_formula = location_name ~ target_type) + 
  ggplot2::theme(legend.position = "bottom") +
    ggplot2::scale_fill_gradient(low = "white", high = "coral")

print(plot)
```

<br>

### Quota issues

One last thing. As you may have experienced, we sometimes experience issues when too many people try to log in at the same time. We are using Google drive to store and load the forecasts and are getting an error that says "Quoata exceeded". Usually, this shouldn't happen at our usage level. We made a Stackoverflow question about it [here](https://stackoverflow.com/questions/66186332/googlesheets-quota-limit-issues-possible-failure-to-use-api-key). In case you have thoughts on that or know people who might, please reach out!

<br>

If you have any further questions or feedback, please don't hesitate to reach out. You can drop us an email or submit an issue to our [github repo](https://github.com/epiforecasts/covid.german.forecasts)

Thank you very much for your support and participation. We very much appreciate it!

Best wishes,

Nikos for the Epiforecasts team

<br>

PS: If you don't want to receive future emails simply write me a quick message. We're working on a way that allows you to unsubscribe from within the app, but that is work in progress. Baby steps, but we're getting there.