### 2021-10-14 analysis/analysis.Rmd
clean code for first Figure

---
title: "Analysis paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
options(knitr.kable.NA = '')
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)
library(covid.german.forecasts)
library(grates)
library(ggdist)
library(ggthemes)
library(scales)
library(dplyr)
library(tidyr)
library(forcats)

# define colors for deaths and cases
colcases <- brewer.pal(name = "Accent", n = 8)[-c(1, 4)]
coldeaths <- brewer.pal(name = "Accent", n = 8)[-4]
colensemble <- c("#FDC086", brewer.pal(name = "Set2", n = 8))

# helper function for labeling plots
scale_fn <- function(x) {
  ifelse(x >= 5000,
         paste0(x / 1000, "k"), x)
}
```

# Visualisation of aggregate performance across horizons

``` {r}
# helper function to score forecasts
score_helper <- function(data, summarise_by) {
  scores <- eval_forecasts(data, 
                           summarise_by = summarise_by, 
                           quantiles = 0.5, 
                           sd = TRUE)

  # for cases, filter out redundant ensembles as convolution model only does 
  # deaths (e.g. ensemble with all is identical to realised ensemble for cases)
  scores <- scores %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case")
    )

  setnames(scores, old = c("model", "sharpness"), 
           new = c("Model", "dispersion"))
}

# helper function to create the WIS plot with its sub-components in panels A and B
component_plot <- function(scores, type, colour_scheme, x = x, rel = relative,
                           facet_formula, vals = NULL) {

  # define colour values for plot
  # for cases and ensembles, filter out identical ensembles
  if (colour_scheme == "ensemble" & type == "case") {
    scores <- scores %>%
      filter(!(Model %in% c("Hub-ensemble-with-all", 
                            "Hub-ensemble-with-convolution")))
    vals <- vals[-c(3, 4)]
  } else if (type == "case") {
    vals <- vals[-1]
  }

  # pivot scores and rename WIS components
  scores <- scores %>%
    select(Model, target_type, dispersion, interval_score,
           underprediction, overprediction, !!x) %>%
    rename("Dispersion" = dispersion, 
           "Over-prediction" = overprediction, 
           "Under-prediction" = underprediction) %>%
    filter(target_type == type) %>%
    pivot_longer(names_to = "WIS components",
                 cols = c(`Dispersion`, `Over-prediction`, `Under-prediction`)) %>%
    mutate(`WIS components` = fct_relevel(`WIS components`, 
                                          c("Dispersion", "Under-prediction", 
                                            "Over-prediction")))

  # compute relative version of the scores
  if (rel) {
    scores <- scores %>%
      arrange(!!sym(x), target_type, interval_score) %>%
      group_by(!!sym(x), target_type) %>%
      mutate_at("value", ~ . / interval_score[Model %in% c("Hub-ensemble", 
                                                           "Hub-ensemble-mean")])
  }

  # create plot
  p <- scores %>%
    ggplot(aes(x = reorder(Model, (interval_score)), y = value, fill = Model, 
               group = Model, alpha = `WIS components`))

  if (rel) {
    p <- p + 
      geom_hline(yintercept = 1, linetype = "dashed")
  }

  p <- p + 
    scale_alpha_manual(values = c(1, 0.1, 0.6)) +
    scale_fill_manual(values = vals) + 
    geom_bar(stat = "identity", position = "stack",
             color = "black",
             size = 0.1,
             width = 0.8) +
    facet_wrap(as.formula(paste("~", x)), ncol = 4,
               strip.position="bottom") + 
    theme_minimal() + 
    theme(axis.text.x = element_blank(), 
          plot.title = element_text(hjust = 0.5, 
                                    size = 8), 
          legend.position = c(0.12, 0.83)) + 
    labs(x = NULL, title = type, y = "WIS") + 
    guides(fill = "none")

  return(p)
}

# helper function to make subplots E - H
plot_helper <- function(tmp = scores, xvar = x, y, rel = relative, 
                        facet_formula, colour) {
  if (rel) {
    tmp <- tmp %>%
      arrange(target_type, !!sym(xvar)) %>%
      group_by(target_type, !!sym(xvar)) %>%
      mutate_at(y, ~ . / .[Model %in% c("Hub-ensemble", 
                                        "Hub-ensemble-mean")])
  }

  tmp %>%
    ggplot(aes_string(x = xvar, y = y, group = "Model", color = "Model")) +
    geom_point(position = position_dodge(width = 0.1)) +
    geom_line(position = position_dodge(width = 0.1)) +
    facet_wrap(facet_formula, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_color_manual(values = colour) + 
    scale_y_continuous(labels = scale_fn)
}

# function to create Figure 1
aggregate_performance <- function(data, 
                                  x = "horizon",
                                  xlabel = "Forecast horizon",
                                  output_dir = here("analysis", "plots"), 
                                  plotname,
                                  relative = FALSE,
                                  facet_formula = ~ target_type, 
                                  colour_scheme = c("normal", "ensemble"),
                                  vals = NULL) {

  # calculate scores
  scores <- score_helper(data, c("model", "target_type", x))

  # define colour values for the plot
  if (colour_scheme[1] == "normal") {
    vals <- coldeaths
  } else {
    vals <- colensemble
  }

  # plot with wis components (panels A and B) ----------------------------------
  component_case <- component_plot(scores, "case", x = x, vals = vals, 
                                   facet_formula = facet_formula, 
                                   rel = relative, 
                                   colour_scheme = colour_scheme[1])
  component_death <- component_plot(scores, "death", x = x, vals = vals, 
                                    facet_formula = facet_formula, 
                                    rel = relative, 
                                    colour_scheme = colour_scheme[1])

  # plot with dispersion of forecasts in panel E -------------------------------
  p_sharpness <- 
    plot_helper(scores, y = "dispersion", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Dispersion", x = NULL)

  # bias plot in panel F -------------------------------------------------------
  # define appropriate y limits
  bias_limits <- max(abs(scores$bias)) * c(-1.1, 1.1) %>% 
    round(2)
  p_bias <- 
    plot_helper(scores, y = "bias", rel = FALSE, colour = vals, xvar = x,
                facet_formula = facet_formula) + 
    labs(y = "Bias", x = NULL) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    expand_limits(y=bias_limits)

  # plot with absolute error of performance in panel G -------------------------
  p_aem <-
    plot_helper(scores, y = "aem", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Absolute error", x = NULL)

  # plot with sd of performance in panel H -------------------------------------
  p_wis_sd <- 
    plot_helper(scores, y = "interval_score_sd", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "WIS - sd", x = NULL)

  # coverage plots (panel C and D) ---------------------------------------------
  # calculate scores that also include the interval range
  scores <- score_helper(data,  c("model", "target_type", x, "range"))

  # take difference to Hub ensemble if relative
  if (relative) {
    scores <- scores %>%
      arrange(target_type, range, !!sym(x)) %>%
      group_by(target_type, range, !!sym(x)) %>%
      mutate_at("coverage", ~ . - .[Model %in%  c("Hub-ensemble", 
                                                  "Hub-ensemble-mean")])
  }

  p_cov50 <- scores %>%
    filter(range == c(50)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                colour = vals, xvar = x) +
    labs(y = "Diff. cov. - 50% PI", x = "") + 
    scale_y_continuous(labels = scales::percent)

  # if not relative, add a dashed line at 50%
  if (!relative) {
    p_cov50 <- p_cov50 + 
      expand_limits(y=c(0, 1)) + 
      labs(y = "Coverage - 50% PI", x = "") + 
      geom_hline(yintercept = 0.5, linetype = "dashed") + 
      scale_y_continuous(labels = scales::percent)
    }

  p_cov90 <-  scores %>%
    filter(range == c(90)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                xvar = x, colour = vals) + 
    labs(y = "Diff. cov. - 90% PI") + 
    scale_y_continuous(labels = scales::percent)

  # if not relative, add a dashed line at 90%
  if (!relative) {
    p_cov90 <- p_cov90 + 
      expand_limits(y=c(0, 1)) + 
      labs(y = "Coverage - 90% PI", x = "") + 
      geom_hline(yintercept = 0.9, linetype = "dashed") + 
      scale_y_continuous() + 
      scale_y_continuous(labels = scales::percent)
  }

  # put all plots together -----------------------------------------------------
  p_combined4 <- 
    (component_case + component_death) / 
    (p_cov50 + p_cov90) /
    (p_sharpness + p_bias) /
    (p_aem + p_wis_sd) +
    plot_layout(guides = "collect", 
                heights = c(1.3, 1, 1, 1)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

    ggsave(here(output_dir, paste0(plotname, "-v4.png")), 
           height = 9.1, width = 10)
}

# create plots with aggregate performance --------------------------------------
# regular plot for Figure 1
filtered_data[model %in% regular_models] %>%
  aggregate_performance(plotname = "aggregate-performance-all")

# plot with restricted time period for sensitivity analysis for SI
filtered_data[model %in% regular_models] %>%
  filter(!(forecast_date %in% as.character(forecast_dates$death_not_scored))) %>%
  aggregate_performance(plotname = "aggregate-performance-all-late-period")

# plot version that compares locations for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all", 
                        x = "location_name", 
                        xlabel = NULL)

# plot version that compares locations in relative terms for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all-rel", 
                        x = "location_name", 
                        xlabel = NULL, relative = TRUE)

# plot for different median ensembles (in relative terms) for SI
filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  # filter(!(model == "Hub-ensemble-with-all" & target_type == "case")) %>%
  aggregate_performance(plotname = "aggregate-performance-rel-ensemble", 
                        relative = TRUE, colour_scheme = "ensemble")

# plot for different mean ensembles (in relative terms) for SI
filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  # filter(!(model == "Hub-ensemble-with-all" & target_type == "case")) %>%
  aggregate_performance(plotname = "aggregate-performance-rel-ensemble-mean", 
                        relative = TRUE, colour_scheme = "ensemble")

```

--------------------------------------------------------------------------------
## Plot with Visualsiation of daily data

```{r}
df <- copy(dailytruth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
][, target_end_date := as.Date(target_end_date)
]

df[weekdays(target_end_date) == "Sunday", sundays := target_end_date]

weeklytruth <- copy(truth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
]

df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  # geom_point(size = 0.3, alpha = 0.4) + 
  # geom_line(size = 0.5, alpha = 0.4) + 
  geom_col(fill = "lightsteelblue", color = "white") + 
  geom_line(data = weeklytruth, aes(y = true_value / 7)) + 
  geom_point(data = weeklytruth, aes(y = true_value / 7), size = 0.3) +
  # geom_vline(aes(xintercept = sundays), linetype = "dashed", alpha = 0.2) + 
  facet_wrap( ~ target_location, scales = "free", ncol = 1) + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                          max(forecast_dates$full_hub_period + 21))) + 
  theme_minimal()

ggsave(here("analysis", "plots", "daily_truth.png"), width = 10, height = 8)
```

--------------------------------------------------------------------------------
## Plot with Visualsiation of x week ahead forecasts and scores

```{r truth-vs-forecasts, fig.height=12, fig.width=15}

# create helper functions for the plot 
# ------------------------------------------------------------------------------
# truth_and_pred_plot = helper function for the visualisation truth and forecast
# scores_over_time_plot = helper function for plotting scores over time
# forecasts_and_scores = function that organises creation of the plots

truth_and_pred_plot <- function(df, ttype = "case") {

  if (ttype == "case") {
    dodgewidth = 3.6
  } else {
    dodgewidth = 3.7
  }

  setnames(df, old = c("model"), new = c("Model"))

  trend <- 
    copy(epitrend)[, location_target := paste0(str_to_sentence(target_type), 
                                               "s", " in ", location_name)]
  setnames(trend, old = c("classification"), new = c("Classification"))

  limits <- min(forecast_dates$full_hub_period) + 5 + 1:21 * 7

  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = Model), 
                   size = 1.1,
                   alpha = 1,
                   position = position_dodge(width = dodgewidth)) + 
    geom_point(data = trend[target_type == ttype], size = 0.7) +
    geom_line(data = unfiltered_data[target_type == ttype & horizon == 2]) + 
    theme_minimal() + 
    theme(panel.grid.major.x = element_blank()) +
    facet_wrap(~ location_target, scales = "free", ncol = 1) + 
    theme(legend.position = "bottom") + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) + 
      scale_y_continuous(labels = scale_fn)
  return(p)
}

scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {
  limits <- min(forecast_dates$full_hub_period) + 5 + 1:21 * 7

  p <- scores %>%
    ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
    geom_line(aes(colour = model)) + 
    geom_point(aes(colour = model, shape = classification)) + 
    theme_minimal() + 
    facet_wrap( ~ location_target, ncol = 1, scales = "free") +
    theme(legend.position = "bottom") +
    theme(panel.grid.major.x = element_blank()) +
    scale_y_continuous(labels = scale_fn) + # , 
                       # trans = log_trans()) + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) 
  return(p)
}

forecasts_and_scores <- function(data, 
                                 output_dir = here("analysis", "plots"), 
                                 plotname = "figure-forecasts", 
                                 ensembles = FALSE) {

  if (ensembles) {
    colcases <- c(colcases[1:2], colcases[4], colcases[3])
    coldeaths <- colcases
  }

  # create data.frame for plotting predictions ---------------------------------
  df <- data[quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")
  df <- merge(
    df[, c("classification", "speed") := NULL], 
    copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
    by = c("target_end_date", "location_name", "target_type")
  )

  # create data.frame for plotting scores --------------------------------------
  summarise_by <- c("model", "horizon", "location_name", "target_type", 
                    "forecast_date", "target_end_date", "location_target")

  scores_target_forecastdate <- eval_forecasts(
    data = data,
    summarise_by = summarise_by
  )[, forecast_date := as.Date(forecast_date)]

  # copy epitrend data.frame and only keep necessary columns
  epi_tmp <- copy(epitrend)
  epi_tmp[, forecast_date := (as.Date(forecast_date))]
  epi_tmp <- 
    epi_tmp[, .(location_name, forecast_date, classification, target_type)] %>% 
    unique()

  # create plots for all horizons ----------------------------------------------
  for (i in 1:4) {

    # truth plot for cases 
    truth_vs_forecast_cases <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "case") + 
      scale_color_manual(values = colcases) + 
      labs(y = "Forecasts and observed values", x = "")  + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm")) + 
      scale_y_continuous(labels = scale_fn)

    # truth plot for deaths
    truth_vs_forecast_deaths <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "death")  + 
      scale_color_manual(values = coldeaths) + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Forecasts and observed values", x = "") +
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm"))

    # scores plot for cases
    scores_over_time_cases <- 
      scores_target_forecastdate[horizon == i & target_type == "case"] %>%
      merge(epi_tmp) %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = colcases) + 
      labs(y = "Weighted interval score", x = "") + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none")

    # scores plot for deaths
    scores_over_time_deaths <- 
      scores_target_forecastdate[horizon == i & target_type == "death"] %>%
      merge(epi_tmp) %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = coldeaths) + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Weighted interval score", x = "")

    # putting them all together
    p <- (truth_vs_forecast_cases + scores_over_time_cases) / 
      (truth_vs_forecast_deaths + scores_over_time_deaths) + 
      plot_layout(guides = "collect") &
      plot_layout(widths = c(1.9, 1)) &
      plot_annotation(tag_levels = 'A') &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin = margin())

    ggsave(here(output_dir, 
                paste0(plotname, "-", i, ".png")), width = 10, height = 11)

  }
}

# create plots -----------------------------------------------------------------
# regular forecasts
unfiltered_data[model %in% c(regular_models)] %>%
  forecasts_and_scores()

data <- unfiltered_data[model %in% c(regular_models, "Renewal-retrospective", "Convolution-retrospective")]

unfiltered_data$model %>% unique()

df$model %>% unique()
# ensemble variants

# models <- c("Hub-ensemble-all", "Crowd forecast", "Hub-ensemble", "Renewal")
# unfiltered_data[model %in% models] %>%
#   forecasts_and_scores(plotname = "figure-forecasts-ensembles", 
#                        ensembles = TRUE)

```

--------------------------------------------------------------------------------
## Summary tables

```{r}

# create helper function to make table
# ------------------------------------------------------------------------------
score_table <- function(scores, 
                        group = TRUE, 
                        full_width = TRUE) {
  # copy data, change names, round column values and reduce to sign. digits
  scores <- as.data.table(scores)
  setnames(scores, 
           old = c("horizon", "model", "target_type", "interval_score", "aem", 
                   "relative_skill", "scaled_rel_skill", "sharpness", 
                   "underprediction", "overprediction", "bias", 
                   "50", "90", "location_target", "target_phase"), 
           new = c("Horizon", "Model", "Target", "WIS", "Abs. error", 
                   "rel. skill", "WIS - rel.", "Dispersion", 
                   "Underpred.", "Overpred.", "Bias", 
                   "50%-Cov.", "90%-Cov.", "Target", "Target"), 
           skip_absent = TRUE)

  # remove some cols, change column order, sort table and rename horizon
  if (all(c("coverage_deviation", "rel. skill") %in% names(scores))) {
    scores <- scores[, .SD,  .SDcols = !c("coverage_deviation", "rel. skill")]
  }

  if (all(c("WIS", "Horizon") %in% names(scores))) {
    wis <- names(scores)[grepl("WIS", names(scores))]
    setcolorder(scores, 
                neworder = c("Target", "Horizon", "Model", wis))  
  }

  if ("Horizon" %in% names(scores)) {
    scores <- scores[order(Target, Horizon)]
    scores[, Horizon := paste(Horizon, "wk ahead")]
  }

  # calculate where to group rows
  case_rows <- nrow(scores[Target == "case"])

  table <- scores[,!c("Target")] %>%
    kable(format = "latex", booktabs = TRUE, 
          align = c("l", "l", rep("c", 9)),
          col.names = c(" ", names(scores)[-(1:2)])) %>%
    collapse_rows(columns = 1) %>%
    kable_styling(latex_options = c("scale_down", 
                                    "hold_position")) %>%
    column_spec(1, background = "white")

  if (group) {
    table <- table %>%
      pack_rows("Cases", 1, case_rows, indent = FALSE, 
                hline_after = TRUE) %>% 
      pack_rows("Deaths", case_rows + 1, nrow(scores), indent = FALSE, 
                hline_after = TRUE)
  }

  return(table)
}

make_table <- function(data, 
                       plotname = "table_scores",
                       output_dir = here("analysis", "plots")) {

  coverage <- eval_forecasts(
    data = data, 
    summarise_by = c("horizon", "model", "target_type", "range"), 
  )[range %in% c(50, 90), 
    .(model, target_type, coverage, range, horizon)] %>%
    dcast(formula = ... ~ range, value.var = "coverage") %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case"))

  df <- eval_forecasts(
    data, 
    summarise_by = c("model", "horizon", "target_type"), 
    compute_relative_skill = FALSE,
    sd = TRUE) %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case")) %>%
    merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

  # round digits
  df <- df[, lapply(.SD, FUN = function(x) {
    if (is.numeric(x)) {
      return(round(signif(x, 3), 2))
    } else {
      return(x)
    }
  })]

  percentage_cols <- c("interval_score", 
                       "interval_score_sd", "sharpness", "underprediction", 
                       "overprediction", "aem")

  # no idea how to do that in data.table: 
  df <- df %>%
    dplyr::group_by(horizon, target_type) %>%
    dplyr::mutate_at(percentage_cols, 
                     ~ paste0(., " (", 
                              round(./.[model %in% c("Hub-ensemble", 
                                                     "Hub-ensemble-mean")], 2), 
                              ")"))

  setDT(df)

  del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                          !(grepl("interval", names(df)))]
  del_cols <- c(del_cols, "coverage_deviation")
  df[, (del_cols) := NULL]

  setnames(df, old = c("interval_score_sd, sharpness"), 
           new = c("WIS - sd", "dispersion"))

  # tables for 1-2 weeks ahead and 3-4 weeks ahead
  table <- score_table(df[horizon <= 2])

  saveRDS(object = table, 
          file = here("analysis", "plots", paste0(plotname, "_", 2, "_ahead.rds")))

  table <- score_table(df[horizon > 2]) 
  saveRDS(object = table, 
          file = here("analysis", "plots", paste0(plotname, "_", 4, "_ahead.rds")))
}

# create tables
# ------------------------------------------------------------------------------

filtered_data[model %in% regular_models] %>%
  make_table()

filtered_data[model %in% regular_models] %>%
  filter(!(forecast_date %in% as.character(forecast_dates$death_not_scored))) %>%
  make_table(plotname = "table-late-period")

filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_mean-ensemble_scores")

filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_median-ensemble_scores")

```

### Tables for different phases of the epidemic

```{r fig.height=10, fig.width=5}

coverage <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("horizon", "model", "target_type", "target_phase", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, target_phase, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "horizon", "target_type", "target_phase"), 
  metrics = c("interval_score", "coverage_deviation"),
  sd = TRUE, 
  quantiles = 0.5
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_phase", "target_type"))

del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                        !(grepl("interval", names(df)))]
del_cols <- c(del_cols, "coverage_deviation")
df[, (del_cols) := NULL]

setnames(df, old = c("interval_score_0.5", "interval_score_sd"), 
         new = c("WIS - median", 
                 "WIS - sd"))

tables <- list()
table_dfs <- list()
targets <- unique(df$target_phase)

# for horizon <= 2
labels <- df[horizon <= 2][order(target_phase, horizon)]$target_phase
table <- score_table(df[horizon <= 2, !"target_type"], group = FALSE) %>%
  pack_rows(index = table(labels), indent = FALSE)
save_kable(table, 
           file = here("analysis", "plots", "table-phases-2.png"))

# for horizon > 2
labels <- df[horizon > 2][order(target_phase, horizon)]$target_phase
table <- score_table(df[horizon > 2, !"target_type"], group = FALSE) %>%
  pack_rows(index = table(labels), indent = FALSE)
save_kable(table, 
           file = here("analysis", "plots", "table-phases-4.png"))

```

### Look at ensemble forecasts

#### Number of members in the ensemble
```{r}
ens <- ensemble_members %>%
  melt(id.vars = c("Date", "model"))

epinow2_included <- ens[model == "epiforecasts-EpiNow2", "epinow" := value]
crowd_included <- ens[model == "epiforecasts-EpiExpert", "crowd" := value]

targetvars <- data.frame(
  variable = c("GM_inc_case", "GM_inc_death", "PL_inc_case", "PL_inc_death"), 
  target = c("Cases - Germany", "Deaths - Germany", "Cases - Poland", "Deaths - Poland")
)

ens <- ens[, .(value = sum(value), 
             epinow = sum(epinow, na.rm = TRUE), 
             crowd = sum(crowd, na.rm = TRUE)), by = c("variable", "Date")] %>%
  merge(targetvars)

# number of submissions from crowd and epinow2
ens[, .(crowd = sum(crowd) / 4, 
        epinow = sum(epinow) / 4)]

# how many models included in the ensemble? 
ens %>%
  summary()

p <- ens %>%
  ggplot(aes(y = value, x = Date)) + 
  facet_wrap(~ target) + 
  # geom_col(aes(fill = variable), position = position_dodge()) + 
  geom_point(size = 0.7) + 
  geom_line() + 
  scale_color_manual(values = colcases) + 
  theme_minimal() + 
  labs(y = "Number of ensemble members", x = "Date")

ggsave(here("analysis", "plots", "ensemble-members.png"), 
       plot = p)

```

### Comparing ensembles with and without a model

```{r}

comparison_ensembles <- function(models_to_plot) {
  df <- unfiltered_data[model %in% models_to_plot &
                          quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")
  df <- merge(
    df[, c("classification", "speed") := NULL], 
    copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
    by = c("target_end_date", "location_name", "target_type")
  )

  p <- df[grepl(2, target)] %>%
    ggplot(aes(y = `0.5`, x = target_end_date, color = model)) +
    geom_line() +
    geom_line(aes(y = true_value), color = "black") + 
    theme_minimal() + 
    facet_wrap(~ location_target, scales = "free", ncol = 2) + 
    theme(legend.position = "bottom") + 
    # theme(panel.background = element_rect(fill = "#e1f0fe")) + 
    scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 14, 
                            max(forecast_dates$full_hub_period))) + 
    labs(y = "Median forecast", x = "Date")
  return(p)
}

models_plot <- c("Crowd forecast", "Hub-ensemble-with-crowd", "Hub-ensemble")
p <- comparison_ensembles(models_plot)
ggsave(here("analysis", "plots", "ensemble-with-and-without-crowd.png"), 
       plot = p)

models_plot <- c("Renewal", "Hub-ensemble-with-renewal", "Hub-ensemble")
p <- comparison_ensembles(models_plot)
ggsave(here("analysis", "plots", "ensemble-with-and-without-renewal.png"), 
       plot = p)

models_plot <- c("Renewal", "Crowd forecast", "Hub-ensemble-all", "Hub-ensemble")
p <- comparison_ensembles(models_plot)
ggsave(here("analysis", "plots", "ensemble-with-and-without-both.png"), 
       plot = p)

```

------------------------------------------------------------------------

## Analyse distribution of scores

```{r}
scores <- eval_forecasts(
  filtered_data[(model %in% regular_models)], 
  summarise_by = c("model", "target_type", "location_name", "horizon", 
                   "forecast_date", "target_phase",
                   "location_target", "classification"),
  compute_relative_skill = FALSE
)

scores[, .(sd = sd(interval_score)), by = c("model", "target_type")]

scores[, target_type := paste0(str_to_title(target_type), "s overall")]

# scores for different phases of the epidemic
scores[, target_phase := factor(target_phase, 
                                levels = str_to_title(c("Cases - decreasing phase", 
                                                        "Cases - unclear phase", 
                                                        "Cases - increasing phase", 
                                                        "Deaths - decreasing phase", 
                                                        "Deaths - unclear phase", 
                                                        "Deaths - increasing phase")))]

distribution_plot_wis <- function(df, x = "interval_score", ranks = FALSE) {

  xlabel = "Weighted interval score"
  dotsize = 1

  if (ranks) {
    df <- df %>%
      group_by(location_name, target_type, horizon, forecast_date) %>%
      arrange(location_name, target_type, horizon, forecast_date, interval_score) %>%
      mutate(!!x := rank(get(x))) %>%
      ungroup() 
    xlabel <- "Rank in terms of WIS"
    dotsize = 0.1

  }

  df %>%
    rename(Model = model) %>%
    ggplot(aes_string(y = "Model", x = x, 
                      group = "Model", fill = "Model")) + 
    ggdist::stat_slab(alpha = 0.3, scale = 0.8, normalize = "panels") +
    ggdist::stat_dots(aes(color = Model),
                      dotsize = dotsize) + 
    ggdist::stat_pointinterval(point_interval = mean_qi, size = 3, 
                               shape = 16, fill = "black", .width = NA) + 
    ggdist::stat_pointinterval(point_interval = median_qi, size = 3, 
                               shape = 15, fill = "black", .width = NA) + 
        facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) +
    scale_fill_manual(values = coldeaths) + 
    scale_color_manual(values = coldeaths) + 
    theme(legend.position = "bottom") + 
    labs(y = "", x = xlabel) + 
  scale_x_continuous(labels = scale_fn) + 
    theme_minimal()
}

for (i in 1:4) {
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score") +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

  p_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

  p_phases_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
    facet_wrap( ~ target_phase, scales = "free_x")

  # pl_wis <- wrap_plots(p_combined_wis, p_wis, widths = c(1, 2)) / p_phases_wis + 
  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here("analysis", "plots", 
              paste0("distribution_scores_wis-", i, ".png")), 
         plot = pl_wis, height = 6, width = 9)

}

# plot with rank distributions
for (i in 1:4) {
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score", 
                                          ranks = TRUE) +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

  p_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                 ranks = TRUE) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

  p_phases_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
    facet_wrap( ~ target_phase, scales = "free_x")

  # pl_wis <- wrap_plots(p_combined_wis, p_wis, widths = c(1, 2)) / p_phases_wis + 
  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here("analysis", "plots", 
              paste0("distribution_scores_wis-", i, "-ranks.png")), 
         plot = pl_wis, height = 6, width = 9)

}

# same plot for absolute error
p_combined_ae <- distribution_plot_wis(df = scores, 
                                        x = "aem") +
  facet_wrap(~ target_type, ncol = 1, scales = "free_x")

p_ae <- distribution_plot_wis(df = scores, x = "aem") + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

```

```{r}
scores[, .(target_count = .N), by = c("target_phase", "model")]

# p_phases <- distribution_plot(df = scores[model != "Baseline"]) + 
#   facet_wrap( ~ target_phase)

p_phases_ae <- distribution_plot_wis(df = scores, x = "aem") + 
  facet_wrap( ~ target_phase, scales = "free_x")

scores[, .(n = .N), by = "target_phase"]

```

```{r fig.height=10, fig.width=15}
pl_ae <- wrap_plots(p_combined_ae, p_ae, widths = c(1, 2)) / p_phases_wis + 
  plot_layout(guides = "collect") & 
  plot_annotation(tag_levels = 'A') &
  theme(legend.position = 'bottom')

ggsave("plots/distribution_scores_ae.png", pl_ae, height = 9)

```

------------------------------------------------------------------------

## WIS contributions

```{r}
# df <- data[grepl("2", target)] %>%
#   merge(epitrend, 
#         by = c("target_end_date", "location_name", "target_type"))

scores_target <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "target_type", "classification"), 
  compute_relative_skill = TRUE
)

scores_target[, target_phase := paste0(target_type, "s - ", 
                                       classification, " phase")]

wis_components(scores_target, 
               facet_formula = ~ target_phase, 
               scales = "free_x", 
               relative_contributions = TRUE,
               x_text_angle = 0) + 
  theme(legend.position = "bottom") + 
  coord_flip() + 
  labs(x = NULL)

ggsave(here("analysis", "plots", "wis-components.png"))

```

### Distribution of Bias over time

```{r}
# df <- data[grepl("2", target)] %>%
#   merge(epitrend, 
#         by = c("target_end_date", "location_name", "target_type"))

scores_target <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "target_type", "classification", "forecast_date")
)

scores_target[, target_phase := paste0(target_type, "s - ", 
                                       classification, " phase")]

p <- scores_target[model %in% regular_models] %>%
  distribution_plot_wis(x = "bias") + 
  scale_x_continuous() + 
  facet_wrap(~ target_phase, ncol = 3) + 
  geom_vline(xintercept = 0, color = "black", size = 1.4) + 
  labs(y = "Bias")

ggsave(here("analysis", "plots", "distribution_bias_phases.png"), plot = p)
```

```{r, eval = TRUE}
scores_full <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "location_target", "horizon", "forecast_date"), 
  compute_relative_skill = FALSE
)

means <- scores_full[, .(meanval = mean(interval_score), 
                         medianval = median(interval_score)), 
                     by =c("model", "location_target", "horizon")]

scores_full %>%
  ggplot(aes(y = interval_score, x = horizon, 
             fill = model, color = model)) +
  geom_boxplot(outlier.shape = 21, coef = 1.5, 
               width=0.5, alpha = 1, 
               position = position_dodge2(width = 0.5)) + 
  # geom_line(data = means, 
  #           aes(group = model, y = meanval, color = model), 
  #           position = position_dodge2(width = 0.5), 
  #           size = 2) +
  # theme(panel.background = element_rect(fill = "#e1f0fe")) +
  scale_color_manual(values = coldeaths) + 
  scale_fill_manual(values = coldeaths) + 
  geom_point(data = means, 
             mapping = aes(y = meanval, fill = model),
             color = "black",
             # fill = "red",
             shape = 21,
             size = 2,
             position = position_dodge2(width = 0.5)) + 
    geom_point(data = means, 
             mapping = aes(y = medianval, fill = model),
             color = "black",
             # fill = "red",
             shape = 22,
             size = 2,
             position = position_dodge2(width = 0.5)) + 
  scale_y_log10() + 
  facet_wrap( ~ location_target, scales = "free") + 
  theme(legend.position = "bottom")

ggsave("plots/scores_horizons.png")

```

------------------------------------------------------------------------

## Calibration

```{r, eval = TRUE}
scores_target <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "target_type", "classification", "range", "quantile"), 
  compute_relative_skill = FALSE, 
  pit_plots = TRUE
)

interval_coverage(scores_target, 
                  facet_formula = ~ target_type + classification)

```

```{r, eval = FALSE}
scores <- list()
# scores for restricted data set
scores_full <- eval_forecasts(filtered_data, 
                              summarise_by = c("model", "target_type", "target"), 
                              compute_relative_skill = TRUE)

scores[["full"]] <- 
  scores_full[, .(model, target_type, scaled_rel_skill, interval_score, target,
                  scenario = "full")]

res <- rbindlist(scores)
res %>%
  ggplot(aes(x = scenario, y = scaled_rel_skill, color = model, group = model)) + 
  geom_point() + 
  facet_wrap( ~ target, scales = "free", ncol = 2) + 
  theme(legend.position = "bottom")
```

## Number of expert forecasters

```{r num-forecasters0, results = 'asis'}
# number of experts and non-experts --------------------------------------------
crowdforecast_data %>%
  select(model, expert) %>%
  unique() %>%
  summarise(experts = sum(expert), 
            nonexpert = sum(!expert))
```

```{r num-forecasters, results = 'asis'}
# check number of available forecasters
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                               "EpiNow2_secondary", 
                               "EpiExpert-ensemble", 
                               "EpiNow2")), 
                .(`number of forecasters` = length(unique(model))), , 
                by = c("forecast_date", "location_name", "target_type")
][order(forecast_date)][
  !is.na(forecast_date)]

dt[, .(sd = sd(`number of forecasters`), 
                             mean = mean(`number of forecasters`), 
                             min  = min(`number of forecasters`), 
                             max = max(`number of forecasters`), 
                             median = median(`number of forecasters`)), 
   by = c("location_name", "target_type")]

dt[, .(sd = sd(`number of forecasters`), 
       mean = mean(`number of forecasters`), 
       min  = min(`number of forecasters`), 
       max = max(`number of forecasters`), 
       median = median(`number of forecasters`))]

```

```{r num-forecasters2, results = 'asis'}
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                                        "EpiNow2_secondary", 
                                        "EpiExpert-ensemble", 
                                        "EpiNow2")), 
                         .(`number of forecasters` = length(unique(model))), , 
                         by = c("forecast_date", "location_name", "target_type")
][order(forecast_date)][
  !is.na(forecast_date)]

dt[, location_target := paste0(str_to_title(target_type), 
                               "s", " in ", location_name)]

dt %>%
  ggplot(aes(y = `number of forecasters`, x = forecast_date)) + 
  geom_line() + 
  facet_wrap(~ location_target) + 
  theme_minimal() + 
  labs(y = "Number of forecasters", x = "Date")

ggsave(filename = here("analysis", "plots", "number-forecasters.png"))
```

```{r, results = 'asis'}
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                                        "EpiNow2_secondary", 
                                        "EpiExpert-ensemble", 
                                        "EpiNow2")) & forecast_date %in% forecast_dates$unfiltered]

num_fc <- dt[, .(n_forecasts = length(unique(forecast_date))),
             by = c("model")][order(n_forecasts)]

median(num_fc$n_forecasts)
mean(num_fc$n_forecasts)

num_fc %>%
  ggplot(aes(x = n_forecasts)) + 
  geom_bar() + 
  theme_minimal() + 
  labs(x = "Number of available submissions", 
       y = "Number of forecasters")

ggsave("plots/crowdforecast_regularity.png")

```

## Number of expert forecasters

```{r, results = 'asis'}

dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                                        "EpiNow2_secondary", 
                                        "EpiNow2")) & forecast_date %in% forecast_dates$unfiltered]

crowddata <- merge_pred_and_obs(dt, truth_data)

scores <- eval_forecasts(
  data = crowddata[grepl("2", target)],
  summarise_by = c("model", "target_type"), 
  compute_relative_skill = TRUE, 
  baseline = "EpiExpert-ensemble"
)

plot_df <- merge(num_fc, scores, by = "model")

plot_df %>%
  ggplot(aes(y = scaled_rel_skill, x = n_forecasts)) + 
  geom_point() + 
  geom_line() + 
  theme_minimal() + 
  facet_wrap(~ target_type) + 
  labs(y = "Scaled relative skill", x = "Number of available submissions") + 
  geom_smooth(method='lm', formula= y~x)

ggsave("plots/skill-vs-forecast-freq.png")

```

## Plots with all forecasts

```{r}
# only select every second forecast date to stratify
df <- unfiltered_data[model %in% regular_models &
                        quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
  dcast(... ~ quantile, value.var = "prediction")
df <- merge(
  df[, c("classification", "speed") := NULL], 
  copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
  by = c("target_end_date", "location_name", "target_type")
)

df[, forecast_date := as.Date(forecast_date)]
df[, i := as.numeric(forecast_date - as.Date(min(forecast_date))) %% 2 + 1]

plot_forecast <- function(df, m = "Renewal") {
  min_date <- min(forecast_dates$unfiltered)
  plot_dates <- c((min_date - c(28, 21, 14, 7)), forecast_dates$unfiltered)
  plot_dates <- plot_dates - 2
  d <- copy(truth_data)[, location_target := paste0(str_to_sentence(target_type), 
                                                    "s", " in ", location_name)]
  daily <- copy(dailytruth_data)[, location_target := paste0(str_to_sentence(target_type), 
                                                             "s", " in ", location_name)]
  daily[, week := as_yearweek(target_end_date, firstday = 7L)]
  daily[, sum_val := sum(true_value), by = c("week", "location_target")]
  daily[, scaled_val := true_value - sum_val / 7]

  df <- copy(df)[, forecast_date := as.factor(forecast_date)]

  df[model == m] %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_line() + 
    geom_line(data = d[target_end_date %in% plot_dates],
              aes(y = true_value, x = target_end_date)) +
    geom_line(data = daily[target_end_date >= min(plot_dates) & target_end_date <= max(plot_dates)],
              aes(y = scaled_val, x = target_end_date), alpha = 0.4) +
    geom_vline(aes(xintercept = as.Date(forecast_date), 
                   color = forecast_date), alpha = 0.3, size = 0.4) + 
    ggplot2::geom_ribbon(ggplot2::aes(ymin = `0.025`, ymax = `0.975`, 
                                      group = forecast_date, 
                                      fill = forecast_date), alpha = 0.3) +
    ggplot2::geom_ribbon(ggplot2::aes(ymin = `0.25`, ymax = `0.75`, 
                                      group = forecast_date, 
                                      fill = forecast_date), alpha = 0.5) +
    ggplot2::geom_line(aes(y = `0.5`,group = forecast_date, 
                           colour = forecast_date), alpha = 1) +
    geom_point() + 
    theme_minimal() + 
    facet_grid(location_target ~ i, scale = "free_y") + 
    theme(legend.position = "bottom") 
    # theme(panel.background = element_rect(fill = "#e1f0fe")) + 
    # scale_y_log10() 
}

for (m in c("Renewal", "Convolution", "Crowd forecast")) {
  p <- plot_forecast(df[horizon <= 2], m = m)

  saveRDS(p, file = here("analysis", "plots", 
                         paste0("all-forecasts-", m,".RDS")))

  ggsave(here("analysis", "plots", 
                         paste0("all-forecasts-", m,".png")), 
         height = 15, width = 10)

}

```