### 2021-10-14 analysis/analysis.Rmd
clean code

---
title: "Analysis paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
options(knitr.kable.NA = '')
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)
library(covid.german.forecasts)
library(grates)
library(ggdist)
library(ggthemes)
library(scales)
library(dplyr)
library(tidyr)
library(forcats)

# define colors for deaths and cases
colcases <- brewer.pal(name = "Accent", n = 8)[-c(1, 4)]
coldeaths <- brewer.pal(name = "Accent", n = 8)[-4]
colensemble <- c("#FDC086", brewer.pal(name = "Set2", n = 8))

# helper function for labeling plots
scale_fn <- function(x) {
  ifelse(x >= 5000,
         paste0(x / 1000, "k"), x)
}
```

# Visualisation of aggregate performance across horizons

``` {r}
# helper function to score forecasts
score_helper <- function(data, summarise_by) {
  scores <- eval_forecasts(data, 
                           summarise_by = summarise_by, 
                           quantiles = 0.5, 
                           sd = TRUE)

  # for cases, filter out redundant ensembles as convolution model only does 
  # deaths (e.g. ensemble with all is identical to realised ensemble for cases)
  scores <- scores %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case")
    )

  setnames(scores, old = c("model", "sharpness"), 
           new = c("Model", "dispersion"))
}

# helper function to create the WIS plot with its sub-components in panels A and B
component_plot <- function(scores, type, colour_scheme, x = x, rel = relative,
                           facet_formula, vals = NULL) {

  # define colour values for plot
  # for cases and ensembles, filter out identical ensembles
  if (colour_scheme == "ensemble" & type == "case") {
    scores <- scores %>%
      filter(!(Model %in% c("Hub-ensemble-with-all", 
                            "Hub-ensemble-with-convolution")))
    vals <- vals[-c(3, 4)]
  } else if (type == "case") {
    vals <- vals[-1]
  }

  # pivot scores and rename WIS components
  scores <- scores %>%
    select(Model, target_type, dispersion, interval_score,
           underprediction, overprediction, !!x) %>%
    rename("Dispersion" = dispersion, 
           "Over-prediction" = overprediction, 
           "Under-prediction" = underprediction) %>%
    filter(target_type == type) %>%
    pivot_longer(names_to = "WIS components",
                 cols = c(`Dispersion`, `Over-prediction`, `Under-prediction`)) %>%
    mutate(`WIS components` = fct_relevel(`WIS components`, 
                                          c("Dispersion", "Under-prediction", 
                                            "Over-prediction")))

  # compute relative version of the scores
  if (rel) {
    scores <- scores %>%
      arrange(!!sym(x), target_type, interval_score) %>%
      group_by(!!sym(x), target_type) %>%
      mutate_at("value", ~ . / interval_score[Model %in% c("Hub-ensemble", 
                                                           "Hub-ensemble-mean")])
  }

  # create plot
  p <- scores %>%
    ggplot(aes(x = reorder(Model, (interval_score)), y = value, fill = Model, 
               group = Model, alpha = `WIS components`))

  if (rel) {
    p <- p + 
      geom_hline(yintercept = 1, linetype = "dashed")
  }

  p <- p + 
    scale_alpha_manual(values = c(1, 0.1, 0.6)) +
    scale_fill_manual(values = vals) + 
    geom_bar(stat = "identity", position = "stack",
             color = "black",
             size = 0.1,
             width = 0.8) +
    facet_wrap(as.formula(paste("~", x)), ncol = 4,
               strip.position="bottom") + 
    theme_minimal() + 
    theme(axis.text.x = element_blank(), 
          plot.title = element_text(hjust = 0.5, 
                                    size = 8), 
          legend.position = c(0.12, 0.83)) + 
    labs(x = NULL, title = type, y = "WIS") + 
    guides(fill = "none")

  return(p)
}

# helper function to make subplots E - H
plot_helper <- function(tmp = scores, xvar = x, y, rel = relative, 
                        facet_formula, colour) {
  if (rel) {
    tmp <- tmp %>%
      arrange(target_type, !!sym(xvar)) %>%
      group_by(target_type, !!sym(xvar)) %>%
      mutate_at(y, ~ . / .[Model %in% c("Hub-ensemble", 
                                        "Hub-ensemble-mean")])
  }

  tmp %>%
    ggplot(aes_string(x = xvar, y = y, group = "Model", color = "Model")) +
    geom_point(position = position_dodge(width = 0.1)) +
    geom_line(position = position_dodge(width = 0.1)) +
    facet_wrap(facet_formula, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_color_manual(values = colour) + 
    scale_y_continuous(labels = scale_fn)
}

# function to create Figure 1
aggregate_performance <- function(data, 
                                  x = "horizon",
                                  xlabel = "Forecast horizon",
                                  output_dir = here("analysis", "plots"), 
                                  plotname,
                                  relative = FALSE,
                                  facet_formula = ~ target_type, 
                                  colour_scheme = c("normal", "ensemble"),
                                  vals = NULL) {

  # calculate scores
  scores <- score_helper(data, c("model", "target_type", x))

  # define colour values for the plot
  if (colour_scheme[1] == "normal") {
    vals <- coldeaths
  } else {
    vals <- colensemble
  }

  # plot with wis components (panels A and B) ----------------------------------
  component_case <- component_plot(scores, "case", x = x, vals = vals, 
                                   facet_formula = facet_formula, 
                                   rel = relative, 
                                   colour_scheme = colour_scheme[1])
  component_death <- component_plot(scores, "death", x = x, vals = vals, 
                                    facet_formula = facet_formula, 
                                    rel = relative, 
                                    colour_scheme = colour_scheme[1])

  # plot with dispersion of forecasts in panel E -------------------------------
  p_sharpness <- 
    plot_helper(scores, y = "dispersion", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Dispersion", x = NULL)

  # bias plot in panel F -------------------------------------------------------
  # define appropriate y limits
  bias_limits <- max(abs(scores$bias)) * c(-1.1, 1.1) %>% 
    round(2)
  p_bias <- 
    plot_helper(scores, y = "bias", rel = FALSE, colour = vals, xvar = x,
                facet_formula = facet_formula) + 
    labs(y = "Bias", x = NULL) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    expand_limits(y=bias_limits)

  # plot with absolute error of performance in panel G -------------------------
  p_aem <-
    plot_helper(scores, y = "aem", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Absolute error", x = NULL)

  # plot with sd of performance in panel H -------------------------------------
  p_wis_sd <- 
    plot_helper(scores, y = "interval_score_sd", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "WIS - sd", x = NULL)

  # coverage plots (panel C and D) ---------------------------------------------
  # calculate scores that also include the interval range
  scores <- score_helper(data,  c("model", "target_type", x, "range"))

  # take difference to Hub ensemble if relative
  if (relative) {
    scores <- scores %>%
      arrange(target_type, range, !!sym(x)) %>%
      group_by(target_type, range, !!sym(x)) %>%
      mutate_at("coverage", ~ . - .[Model %in%  c("Hub-ensemble", 
                                                  "Hub-ensemble-mean")])
  }

  p_cov50 <- scores %>%
    filter(range == c(50)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                colour = vals, xvar = x) +
    labs(y = "Diff. cov. - 50% PI", x = "") + 
    scale_y_continuous(labels = scales::percent)

  # if not relative, add a dashed line at 50%
  if (!relative) {
    p_cov50 <- p_cov50 + 
      expand_limits(y=c(0, 1)) + 
      labs(y = "Coverage - 50% PI", x = "") + 
      geom_hline(yintercept = 0.5, linetype = "dashed") + 
      scale_y_continuous(labels = scales::percent)
    }

  p_cov90 <-  scores %>%
    filter(range == c(90)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                xvar = x, colour = vals) + 
    labs(y = "Diff. cov. - 90% PI") + 
    scale_y_continuous(labels = scales::percent)

  # if not relative, add a dashed line at 90%
  if (!relative) {
    p_cov90 <- p_cov90 + 
      expand_limits(y=c(0, 1)) + 
      labs(y = "Coverage - 90% PI", x = "") + 
      geom_hline(yintercept = 0.9, linetype = "dashed") + 
      scale_y_continuous() + 
      scale_y_continuous(labels = scales::percent)
  }

  # put all plots together -----------------------------------------------------
  p_combined4 <- 
    (component_case + component_death) / 
    (p_cov50 + p_cov90) /
    (p_sharpness + p_bias) /
    (p_aem + p_wis_sd) +
    plot_layout(guides = "collect", 
                heights = c(1.3, 1, 1, 1)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

    ggsave(here(output_dir, paste0(plotname, "-v4.png")), 
           height = 9.1, width = 10)
}

# create plots with aggregate performance --------------------------------------
# regular plot for Figure 1
filtered_data[model %in% regular_models] %>%
  aggregate_performance(plotname = "aggregate-performance-all")

# plot with restricted time period for sensitivity analysis for SI
filtered_data[model %in% regular_models] %>%
  filter(!(forecast_date %in% as.character(forecast_dates$death_not_scored))) %>%
  aggregate_performance(plotname = "aggregate-performance-all-late-period")

# plot version that compares locations for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all", 
                        x = "location_name", 
                        xlabel = NULL)

# plot version that compares locations in relative terms for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all-rel", 
                        x = "location_name", 
                        xlabel = NULL, relative = TRUE)

# plot for different median ensembles (in relative terms) for SI
filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  # filter(!(model == "Hub-ensemble-with-all" & target_type == "case")) %>%
  aggregate_performance(plotname = "aggregate-performance-rel-ensemble", 
                        relative = TRUE, colour_scheme = "ensemble")

# plot for different mean ensembles (in relative terms) for SI
filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  # filter(!(model == "Hub-ensemble-with-all" & target_type == "case")) %>%
  aggregate_performance(plotname = "aggregate-performance-rel-ensemble-mean", 
                        relative = TRUE, colour_scheme = "ensemble")

```

# Plot with Visualsiation of x week ahead forecasts and scores

```{r truth-vs-forecasts, fig.height=12, fig.width=15}

# helper function which creates panels A and C
truth_and_pred_plot <- function(df, ttype = "case") {

  # adjust distance between bars depending on case or death forecast
  if (ttype == "case") {
    dodgewidth = 3.6
  } else {
    dodgewidth = 3.7
  }

  # rename model column
  setnames(df, old = c("model"), new = c("Model"))

  # define date limits for the plot
  limits <- min(forecast_dates$full_hub_period) + 5 + 0:21 * 7

  # create plot
  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = Model), 
                   size = 1.1,
                   alpha = 1,
                   position = position_dodge(width = dodgewidth)) + 
    geom_line(data = unfiltered_data[target_type == ttype & 
                                       target_end_date %in% limits]) + 
    theme_minimal() + 
    theme(panel.grid.major.x = element_blank()) +
    facet_wrap(~ location_target, scales = "free", ncol = 1) + 
    theme(legend.position = "bottom") + 
    scale_x_date(limits = c(min(limits) - 2, max(limits) + 2), 
                 minor_breaks = limits) + 
      scale_y_continuous(labels = scale_fn)
  return(p)
}

# helper function which creates panels B and D
scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {

  # define plot limits on x axis
  limits <- min(forecast_dates$full_hub_period) + 5 + 0:20 * 7

  p <- scores %>%
    ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
    geom_line(aes(colour = model)) + 
    geom_point(aes(colour = model)) + 
    theme_minimal() + 
    facet_wrap( ~ location_target, ncol = 1, scales = "free") +
    theme(legend.position = "bottom") +
    theme(panel.grid.major.x = element_blank()) +
    scale_y_continuous(labels = scale_fn) + # , 
                       # trans = log_trans()) + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) 
  return(p)
}

# helper function that creates the final plot
forecasts_and_scores <- function(data, 
                                 output_dir = here("analysis", "plots"), 
                                 plotname = "figure-forecasts") {

  # calculate scores
  summarise_by <- c("model", "horizon", "location_name", "target_type", 
                    "forecast_date", "target_end_date", "location_target")

  scores_target_forecastdate <- eval_forecasts(
    data = data,
    summarise_by = summarise_by
  )[, forecast_date := as.Date(forecast_date)]

  # create data.frame for plotting predictions 
  df <- data[quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")

  # create plots for all horizons 
  for (i in 1:4) {

    # truth plot for cases 
    truth_vs_forecast_cases <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "case") + 
      scale_color_manual(values = colcases, name = "Model") + 
      labs(y = "Forecasts and observed values", x = "")  + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm")) + 
      scale_y_continuous(labels = scale_fn)

    # truth plot for deaths
    truth_vs_forecast_deaths <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "death")  + 
      scale_color_manual(values = coldeaths, name = "Model") + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 5 + (i-1) * 7,
               xmax = max(forecast_dates$death_not_scored) + 5 + (i-1) * 7,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Forecasts and observed values", x = "") +
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm"))

    # scores plot for cases
    scores_over_time_cases <- 
      scores_target_forecastdate[horizon == i & target_type == "case"] %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = colcases, name = "Model") + 
      labs(y = "Weighted interval score", x = "") + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none")

    # scores plot for deaths
    scores_over_time_deaths <- 
      scores_target_forecastdate[horizon == i & target_type == "death"] %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = coldeaths, name = "Model") + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Weighted interval score", x = "")

    # putting them all together
    p <- (truth_vs_forecast_cases + scores_over_time_cases) / 
      (truth_vs_forecast_deaths + scores_over_time_deaths) + 
      plot_layout(guides = "collect") &
      plot_layout(widths = c(1.9, 1)) &
      plot_annotation(tag_levels = 'A') &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin = margin())

    ggsave(here(output_dir, 
                paste0(plotname, "-", i, ".png")), width = 10, height = 11)

  }
}

# create plots -----------------------------------------------------------------
# regular forecasts
unfiltered_data[model %in% c(regular_models)] %>%
  forecasts_and_scores()
```

# Create tables with scores

```{r}

# helper function to create the tables
score_table <- function(scores, 
                        group = TRUE, 
                        full_width = TRUE) {

  # change column names, round column values and reduce to sign. digits
  scores <- as.data.table(scores)
  setnames(scores, 
           old = c("horizon", "model", "target_type", "interval_score", "aem", 
                   "relative_skill", "scaled_rel_skill", "sharpness", 
                   "underprediction", "overprediction", "bias", 
                   "50", "90", "location_target", "target_phase"), 
           new = c("Horizon", "Model", "Target", "WIS", "Abs. error", 
                   "rel. skill", "WIS - rel.", "Dispersion", 
                   "Underpred.", "Overpred.", "Bias", 
                   "50%-Cov.", "90%-Cov.", "Target", "Target"), 
           skip_absent = TRUE)

  # remove some cols, change column order, sort table
  if (all(c("coverage_deviation", "rel. skill") %in% names(scores))) {
    scores <- scores[, .SD,  .SDcols = !c("coverage_deviation", "rel. skill")]
  }

  if (all(c("WIS", "Horizon") %in% names(scores))) {
    wis <- names(scores)[grepl("WIS", names(scores))]
    setcolorder(scores, 
                neworder = c("Target", "Horizon", "Model", wis))  
  }

  if ("Horizon" %in% names(scores)) {
    scores <- scores[order(Target, Horizon)]
    scores[, Horizon := paste(Horizon, "wk ahead")]
  }

  # calculate where to group rows
  case_rows <- nrow(scores[Target == "case"])

  table <- scores[,!c("Target")] %>%
    kable(format = "latex", booktabs = TRUE, 
          align = c("l", "l", rep("c", 9)),
          col.names = c(" ", names(scores)[-(1:2)])) %>%
    collapse_rows(columns = 1) %>%
    kable_styling(latex_options = c("scale_down", 
                                    "hold_position")) %>%
    column_spec(1, background = "white")

  if (group) {
    table <- table %>%
      pack_rows("Cases", 1, case_rows, indent = FALSE, 
                hline_after = TRUE) %>% 
      pack_rows("Deaths", case_rows + 1, nrow(scores), indent = FALSE, 
                hline_after = TRUE)
  }

  return(table)
}

# function that actually makes the tables
make_table <- function(data, 
                       plotname = "table_scores",
                       output_dir = here("analysis", "plots")) {

  # get coverage values
  coverage <- eval_forecasts(
    data = data, 
    summarise_by = c("horizon", "model", "target_type", "range"), 
  )[range %in% c(50, 90), 
    .(model, target_type, coverage, range, horizon)] %>%
    dcast(formula = ... ~ range, value.var = "coverage") %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case"))

  # compute scores and merge with coverage values
  df <- eval_forecasts(
    data, 
    summarise_by = c("model", "horizon", "target_type"), 
    compute_relative_skill = FALSE,
    sd = TRUE) %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case")) %>%
    merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

  # round digits
  df <- df[, lapply(.SD, FUN = function(x) {
    if (is.numeric(x)) {
      return(round(signif(x, 3), 2))
    } else {
      return(x)
    }
  })]

  percentage_cols <- c("interval_score", 
                       "interval_score_sd", "sharpness", "underprediction", 
                       "overprediction", "aem")

  # add percentage values in brackets
  df <- df %>%
    dplyr::group_by(horizon, target_type) %>%
    dplyr::mutate_at(percentage_cols, 
                     ~ paste0(., " (", 
                              round(./.[model %in% c("Hub-ensemble", 
                                                     "Hub-ensemble-mean")], 2), 
                              ")"))

  setDT(df)

  # delete and rename a few columns
  del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                          !(grepl("interval", names(df)))]
  del_cols <- c(del_cols, "coverage_deviation")
  df[, (del_cols) := NULL] 
  setnames(df, old = c("interval_score_sd, sharpness"), 
           new = c("WIS - sd", "dispersion"))

  # create and save tables for 1-2 weeks ahead and 3-4 weeks ahead
  table <- score_table(df[horizon <= 2]) 
  saveRDS(object = table, 
          file = here("analysis", "plots", paste0(plotname, "_", 2, "_ahead.rds")))

  table <- score_table(df[horizon > 2]) 
  saveRDS(object = table, 
          file = here("analysis", "plots", paste0(plotname, "_", 4, "_ahead.rds")))
}

# create tables

# regular models
filtered_data[model %in% regular_models] %>%
  make_table()

# regular models with later period as sensitivity analysis in SI
filtered_data[model %in% regular_models] %>%
  filter(!(forecast_date %in% as.character(forecast_dates$death_not_scored))) %>%
  make_table(plotname = "table-late-period")

# table for mean ensembles for SI
filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_mean-ensemble_scores")

# table for median ensembles for SI
filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_median-ensemble_scores")
```

# Plot with Visualsiation of daily data

```{r}

# reformat daily truth data
df <- copy(dailytruth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
][, target_end_date := as.Date(target_end_date)
]

df[weekdays(target_end_date) == "Sunday", sundays := target_end_date]

# reformat weekly truth data
weeklytruth <- copy(truth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
]

# create plot
df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_col(fill = "lightsteelblue", color = "white") + 
  geom_line(data = weeklytruth, aes(y = true_value / 7)) + 
  geom_point(data = weeklytruth, aes(y = true_value / 7), size = 0.3) +
  # geom_vline(aes(xintercept = sundays), linetype = "dashed", alpha = 0.2) + 
  facet_wrap( ~ target_location, scales = "free", ncol = 1) + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                          max(forecast_dates$full_hub_period + 21))) + 
  theme_minimal()

ggsave(here("analysis", "plots", "daily_truth.png"), width = 10, height = 8)
```

# Number of member models in the Hub ensemble

```{r}
ens <- ensemble_members %>%
  melt(id.vars = c("Date", "model"))

epinow2_included <- ens[model == "epiforecasts-EpiNow2", "epinow" := value]
crowd_included <- ens[model == "epiforecasts-EpiExpert", "crowd" := value]

targetvars <- data.frame(
  variable = c("GM_inc_case", "GM_inc_death", "PL_inc_case", "PL_inc_death"), 
  target = c("Cases - Germany", "Deaths - Germany", "Cases - Poland", "Deaths - Poland")
)

ens <- ens[, .(value = sum(value), 
             epinow = sum(epinow, na.rm = TRUE), 
             crowd = sum(crowd, na.rm = TRUE)), by = c("variable", "Date")] %>%
  merge(targetvars)

# number of submissions from crowd and epinow2
ens[, .(crowd = sum(crowd) / 4, 
        epinow = sum(epinow) / 4)]

# how many models included in the ensemble? 
ens %>%
  summary()

p <- ens %>%
  ggplot(aes(y = value, x = Date)) + 
  facet_wrap(~ target) + 
  # geom_col(aes(fill = variable), position = position_dodge()) + 
  geom_point(size = 0.7) + 
  geom_line() + 
  scale_color_manual(values = colcases) + 
  theme_minimal() + 
  labs(y = "Number of ensemble members", x = "Date")

ggsave(here("analysis", "plots", "ensemble-members.png"), 
       plot = p)
```

# Plot with distribution of scores

```{r}
# score forecasts
scores <- eval_forecasts(
  filtered_data[(model %in% regular_models)], 
  summarise_by = c("model", "target_type", "location_name", "horizon", 
                   "forecast_date", "target_phase",
                   "location_target", "classification"),
  compute_relative_skill = FALSE
)

# reformat column
scores[, target_type := paste0(str_to_title(target_type), "s overall")]

# display standard deviation for different models
scores[, .(sd = sd(interval_score)), by = c("model", "target_type")]

# helper function to create distribution plot
distribution_plot_wis <- function(df, x = "interval_score", 
                                  ranks = FALSE) {

  xlabel = "Weighted interval score"
  dotsize = 1

  if (ranks) {
    df <- df %>%
      group_by(location_name, target_type, horizon, forecast_date) %>%
      arrange(location_name, target_type, horizon, forecast_date, interval_score) %>%
      mutate(!!x := rank(get(x))) %>%
      ungroup() 
    xlabel <- "Rank in terms of WIS"
    dotsize = 0.1

  }

  df %>%
    rename(Model = model) %>%
    ggplot(aes_string(y = "Model", x = x, 
                      group = "Model", fill = "Model")) + 
    ggdist::stat_slab(alpha = 0.3, scale = 0.8, normalize = "panels") +
    ggdist::stat_dots(aes(color = Model),
                      dotsize = dotsize) + 
    ggdist::stat_pointinterval(point_interval = mean_qi, size = 3, 
                               shape = 16, fill = "black", .width = NA) + 
    ggdist::stat_pointinterval(point_interval = median_qi, size = 3, 
                               shape = 15, fill = "black", .width = NA) + 
        facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) +
    scale_fill_manual(values = coldeaths) + 
    scale_color_manual(values = coldeaths) + 
    theme(legend.position = "bottom") + 
    labs(y = "", x = xlabel) + 
  scale_x_continuous(labels = scale_fn) + 
    theme_minimal()
}

# create plots for the distributions of WIS
for (i in 1:4) {
  # plot for locations combined
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score") +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

  # plot for locations separate
  p_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

  # combine plots and save
  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here("analysis", "plots", 
              paste0("distribution_scores_wis-", i, ".png")), 
         plot = pl_wis, height = 6, width = 9)
}

# create plots for the distributions of WIS rank
for (i in 1:4) {
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score", 
                                          ranks = TRUE) +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

  p_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                 ranks = TRUE) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here("analysis", "plots", 
              paste0("distribution_scores_wis-", i, "-ranks.png")), 
         plot = pl_wis, height = 6, width = 9)

}
```

# Number of expert forecasters

```{r num-forecasters0, results = 'asis'}
# number of experts and non-experts --------------------------------------------
crowdforecast_data %>%
  select(model, expert) %>%
  unique() %>%
  summarise(experts = sum(expert), 
            nonexpert = sum(!expert))
```

# Number of forecasts over time

```{r num-forecasters, results = 'asis'}
# filter out ensembles and model-based forecasts
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                               "EpiNow2_secondary", 
                               "EpiExpert-ensemble", 
                               "EpiNow2")), 
                .(`number of forecasters` = length(unique(model))), , 
                by = c("forecast_date", "location_name", "target_type")
][order(forecast_date)][
  !is.na(forecast_date)]

# display summary for the number of available forecasters
dt[, .(sd = sd(`number of forecasters`), 
       mean = mean(`number of forecasters`), 
       min  = min(`number of forecasters`), 
       max = max(`number of forecasters`), 
       median = median(`number of forecasters`))]

# display summary for the number of available forecasters by location and target
dt[, .(sd = sd(`number of forecasters`), 
       mean = mean(`number of forecasters`), 
       min  = min(`number of forecasters`), 
       max = max(`number of forecasters`), 
       median = median(`number of forecasters`)), 
   by = c("location_name", "target_type")] 
```

```{r num-forecasters2, results = 'asis'}
# filter out ensembles and model-based forecasts
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                                        "EpiNow2_secondary", 
                                        "EpiExpert-ensemble", 
                                        "EpiNow2")), 
                         .(`number of forecasters` = length(unique(model))), , 
                         by = c("forecast_date", "location_name", "target_type")
][order(forecast_date)][
  !is.na(forecast_date)]

dt[, location_target := paste0(str_to_title(target_type), 
                               "s", " in ", location_name)]

# plot number of forecasters over time
dt %>%
  ggplot(aes(y = `number of forecasters`, x = forecast_date)) + 
  geom_line() + 
  facet_wrap(~ location_target) + 
  theme_minimal() + 
  labs(y = "Number of forecasters", x = "Date")

ggsave(filename = here("analysis", "plots", "number-forecasters.png"))
```

# Number of submissions per forecaster

```{r, results = 'asis'}
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                                        "EpiNow2_secondary", 
                                        "EpiExpert-ensemble", 
                                        "EpiNow2")) & forecast_date %in% forecast_dates$unfiltered]

num_fc <- dt[, .(n_forecasts = length(unique(forecast_date))),
             by = c("model")][order(n_forecasts)]

median(num_fc$n_forecasts)
mean(num_fc$n_forecasts)

num_fc %>%
  ggplot(aes(x = n_forecasts)) + 
  geom_bar() + 
  theme_minimal() + 
  labs(x = "Number of available submissions", 
       y = "Number of forecasters")

ggsave("plots/crowdforecast_regularity.png")

```