### 2022-05-30 manuscript/manuscript.tex

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{lineno}
\usepackage[T1]{fontenc}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{London School of Hygiene \& Tropical Medicine,Karlsruhe
Institute of Technology Institute of Economic Theory and Statistics,Max
Planck Institute for Multidisciplinary Sciences,UK Health Security
Agency,Imperial College London}
}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Comparing human and model-based forecasts of COVID-19 in
Germany and
Poland} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Nikos I. Bosse\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}}\textsuperscript{*},
Sam Abbott\textsuperscript{\getAff{London School of Hygiene \& Tropical
Medicine}},
Johannes Bracher\textsuperscript{\getAff{Karlsruhe Institute of
Technology Institute of Economic Theory and Statistics}},
Habakuk Hain\textsuperscript{\getAff{Max Planck Institute for
Multidisciplinary Sciences}},
Billy J. Quilty\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}},
Mark Jit\textsuperscript{\getAff{London School of Hygiene \& Tropical
Medicine}},
Centre for the Mathematical Modelling of Infectious Diseases COVID-19
Working Group\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}},
Edwin van Leeuwen\textsuperscript{\getAff{UK Health Security
Agency}, \getAff{London School of Hygiene \& Tropical Medicine}},
Anne Cori\textsuperscript{\getAff{Imperial College London}},
Sebastian Funk\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}}\\
\bigskip
\textbf{\getAff{London School of Hygiene \& Tropical Medicine}}London
School of Hygiene \& Tropical Medicine, Keppel Street, London WC1E7HT,
United Kingdom\\
\textbf{\getAff{Karlsruhe Institute of Technology Institute of Economic
Theory and Statistics}}Karlsruhe Institute of Technology - Institute of
Economic Theory and Statistics, Post Box 6980, 76049 Karlsruhe,
Germany\\
\textbf{\getAff{Max Planck Institute for Multidisciplinary Sciences}}Max
Planck Institute for Multidisciplinary Sciences, Am Faßberg 11, 37077
Göttingen, Germany\\
\textbf{\getAff{UK Health Security Agency}}UK Health Security Agency,
London NW9 5EQ, United Kingdom\\
\textbf{\getAff{Imperial College London}}Imperial College London,
Norfolk Place, London W2 1PG, United Kingdom\\
\bigskip
* Corresponding author: nikos.bosse@lshtm.ac.uk\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Forecasts based on epidemiological modelling have played an important
role in shaping public policy throughout the COVID-19 pandemic. This
modelling combines knowledge about infectious disease dynamics with the
subjective opinion of the researcher who develops and refines the model
and often also adjusts model outputs. Developing a forecast model is
difficult, resource- and time-consuming. It is therefore worth asking
what modelling is able to add beyond the subjective opinion of the
researcher alone. To investigate this, we analysed different real-time
forecasts of cases of and deaths from COVID-19 in Germany and Poland
over a 1-4 week horizon submitted to the German and Polish Forecast Hub.
We compared crowd forecasts elicited from researchers and volunteers,
against a) forecasts from two semi-mechanistic models based on common
epidemiological assumptions and b) the ensemble of all other models
submitted to the Forecast Hub. We found crowd forecasts, despite being
overconfident, to outperform all other methods across all forecast
horizons when forecasting cases (weighted interval score relative to the
Hub ensemble 2 weeks ahead: 0.89). Forecasts based on computational
models performed comparably better when predicting deaths (rel. WIS
1.26), suggesting that epidemiological modelling and human judgement can
complement each other in important ways.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
Mathematical models of COVID-19 have played a key role in informing
governments across the world. While mathematical models are informed by
our knowledge of infectious disease dynamics, they are ultimately
developed and iteratively adjusted by the researchers and shaped by
their subjective opinions. To investigate what modelling is able to add
beyond the subjective opinion of the researcher alone, we compared human
forecasts with model-based predictions of COVID-19 cases and deaths
submitted to the so-called German/Polish Forecast Hub (which collates a
variety of models from a range of teams). \textbar{} We found that our
human forecasts consistently outperformed an aggregate of all available
model-based forecasts when predicting cases, but not when predicting
deaths. Our findings suggest that human insight may be most valuable
when forecasting highly uncertain quantities, which depend on many
factors that are hard to model using equations, while mathematical
models may be most useful in settings like predicting deaths, where
leading indicators with a clear connection to the target variable are
available. This potentially has very relevant policy implications, as
agencies informing policy-makers could benefit from routinely eliciting
human forecasts in addition to model-based predictions to inform
policies.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Infectious disease modelling has a long tradition and has helped inform
public health decisions both through scenario modelling, as well as
actual forecasts of (among others) influenza {[}e.g. 1,2--4{]}, dengue
fever {[}e.g. 5,6,7{]}, ebola {[}e.g. 8,9{]}, chikungunya {[}e.g.
10,11{]} and now COVID-19 {[}e.g. 12,13--17{]}. Applications of
epidemiological models differ in the way they make statements about the
future. Forecasts aim to predict the future as it will occur, while
scenario modelling and projections aim to represent what the future
could look like under certain scenario assumptions or if conditions
stayed the same as they were in the past. Forecasts can be judged by
comparing them against observed data. Since it is much harder to fairly
assess the accuracy and usefulness of projections and scenario modelling
in the same way, this work focuses on forecasts, which represent only a
subset of all epidemiological modelling.

Since March 2020, forecasts of COVID-19 from multiple teams have been
collected, aggregated and compared by Forecast Hubs such as the US
Forecast Hub {[}13,14{]}, the German and Polish Forecast Hub {[}15,16{]}
and the European Forecast Hub {[}17{]}. Often, different individual
forecasts are combined into a single forecast, e.g.~by taking the mean
or median of all forecasts. These ensemble forecasts usually tend to
perform better and more consistently than individual forecasts (see e.g.
{[}6{]}; {[}18{]}).

Individual computational models usually rely to varying degrees on
mechanistic assumptions about infectious disease dynamics (such as
SIR-type compartmental models that aim to represent how individuals move
from being susceptible to infected and then recovered or dead). Some are
more statistical in nature (such as time series models that detect
statistical patterns without explicitly modelling disease dynamics). How
exactly such a mathematical or computational model is constructed and
which assumptions are made depends on subjective opinion and judgement
of the researcher who develops and refines the model. Models are
commonly adjusted and improved based on whether the model output looks
plausible to the researchers involved.

The process of model construction and refinement is laborious and
time-consuming, and it is therefore worth asking what modelling can add
beyond the subjective judgment of the researcher alone. In this work, we
ask this question specifically in the context of predictive performance,
and set aside other advantages of epidemiological modelling (such as
reproducibility or the ability to obtain a deeper fundamental
understanding of how diseases spread). One natural way to do this is to
compare the predictive performance of forecasts based on computational
models (``model-based forecasts'') against forecasts made by individual
humans without explicit use of a computer model (``direct human
forecasts'') or a combination of multiple such forecasts (``crowd
forecasts'').

Previous work has examined such direct human forecasts in various
contexts, such as geopolitics {[}19,20{]}, meta-science {[}21,22{]},
sports {[}23{]} and epidemiology {[}11,24,25{]}. Several prediction
platforms {[}26--28{]} and prediction markets {[}29{]} have been created
to collate expert and non-expert predictions. However, with the notable
exception of {[}11{]}, these forecasts were not designed to be evaluated
alongside model-based forecasts and usually follow their own (often
binary) prediction formats. Direct human forecasts may be able to take
into account insights and relationships between variables which are hard
to specify using epidemiological models. However, it is not entirely
clear in which situations human forecasts perform well or badly. For
example, {[}11{]} found that humans could outperform computer models at
predicting the 2014/15 and 2015/16 flu season in the US, a setting where
the disease was well known and information about previous seasons was
available. However, humans tended to do slightly worse at predicting the
2014/15 outbreak of chikungunya in the Americas, a disease previously
largely unobserved and unknown in these regions at the time.

In this study, we analyse the performance of direct human forecasts
relative to model-based forecasts and discuss the added benefit of
epidemiological modelling over human judgement alone. As a case study,
we use different forecasts, involving varying degrees of human
intervention, which we submitted in real time to the German and Polish
Forecast Hub. In contrast to {[}11{]} we elicited not only point
predictions, but full predictive distributions (``probabilistic
forecasts'', see e.g. {[}30{]}) from participants. This allows us to
compare not only predictive accuracy, but also how well human
forecasters and model-based forecasts were able to quantify forecast
uncertainty.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We created and submitted the following forecasts to the German and
Polish Forecast Hub: 1) a direct human forecast (henceforth called
``crowd forecast''), elicited from participants through a web
application {[}31{]} and 2) two semi-mechanistic model-based forecasts
(``renewal model'' and ``convolution model'') informed by basic
assumptions about COVID-19 epidemiology. While the two semi-mechanistic
forecasts were necessarily shaped by our implicit assumptions and
decisions, they were designed such as to minimise the amount of human
intervention involved. For example, we refrained from adjusting model
outputs or refining the models based on past performance. Forecasts were
created in real time over a period of 21 weeks from October 12th 2020
until March 1st 2021 and submitted to the German and Polish Forecast hub
{[}15,16{]}. All code and tools necessary to generate the forecasts and
make a forecast submission are available in the
\texttt{covid.german.forecasts} R package {[}32{]}. This repository also
contains a record of all forecasts submitted to the German and Polish
Forecast Hub. Forecasts were evaluated using a variety of scoring
metrics and compared among each other and against an ensemble of all
other models submitted to the German and Polish Forecast Hub.

\hypertarget{forecast-targets-and-interaction-with-the-german-and-polish-forecast-hub}{%
\subsection{Forecast targets and interaction with the German and Polish
Forecast
Hub}\label{forecast-targets-and-interaction-with-the-german-and-polish-forecast-hub}}

The German and Polish Forecast Hub (now mostly merged into the {[}17{]})
elicits predictions for various COVID-19 related forecast targets from
different research groups every week. Forecasts had to be made every
Monday (with submissions allowed until Tuesday 3pm) and were permitted
to use any data that was available by Monday 11.59pm. We submitted
forecasts for incident and cumulative weekly reported numbers of cases
of and deaths from COVID-19 on a national level in Germany and Poland
over a one to four week forecast horizon. Forecasts were submitted on
Mondays, but weeks were defined as ending on a Saturday (and starting on
Sunday), meaning that forecast horizons were in fact 5, 12, 19 and 26
days. Submissions were required in a quantile-based format with 23
quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10,
0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub
were combined into different ensembles every week, with the median
ensemble (i.e., the \(\alpha\)-quantile of the ensemble is given by the
median of all submitted \(\alpha\)-quantiles) being the default ensemble
shown on all official Forecast hub visualisations
(https://kitmetricslab.github.io/forecasthub/forecast).

Data on daily reported test positive cases and deaths linked to COVID-19
were provided by the organisers of the German and Polish Forecast hub.
Until December 14th, 2020, these data were sourced from the European
Centre for Disease Control {[}33{]}. After ECDC stopped publishing daily
data, observations were sourced from the Robert Koch Institute (RKI) and
the Polish Ministry of Health for the remainder of the submission period
{[}34{]}. These data are subject to reporting artefacts, (such as for
example delayed case reporting in Poland on the 24th November,
{[}35{]}), changes in reporting over time, and variation in testing
regimes (for example in Germany from the 11th of November on, {[}36{]}).
The ECDC data as well as the data published by the Polish Ministry of
Health were also subject to data revisions, although most of them (with
a notable exception of a data update for October 12 2020 in Germany)
only affected daily, not weekly data (see Figures
\ref{fig:daily-truth-update} and \ref{fig:weekly-truth-update}).

\hypertarget{crowd-forecasts}{%
\subsection{Crowd forecasts}\label{crowd-forecasts}}

Our crowd forecasts were created as an ensemble of forecasts made by
individual participants every week through a web application
(https://cmmid-lshtm.shinyapps.io/crowd-forecast/). Weekly forecasts had
to be submitted before Tuesday 12pm every week, but participants were
asked to only use any information or data that was already available by
Monday night. The application was built using the \texttt{shiny} and
\texttt{golem} R packages {[}37,38{]} and is available in the
\texttt{crowdforecastr} R package {[}31{]}. To make a forecast in the
application participants could select a predictive distribution (with
the default being log-normal) to represent the probability that the
forecasted quantity took certain values. Median and width of the
uncertainty could be adjusted by either interacting with a figure
showing their forecast or providing numerical values (see screenshot in
Figure \ref{fig:screenshot} in the SI). The default shown was a
repetition of the last known observation with constant uncertainty
around it computed as the standard deviation of the last four changes in
weekly log observed forecasts (i.e.~as
\(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
A comparison of the crowd forecasts against the default baseline shown
in the application is displayed in Figure \ref{fig:compare-forecasters}
in the Appendix. Our interface also allowed participants to view past
observations based on the hub data, as well as their forecasts, on a
logarithmic scale and presented additional contextual COVID-19 data
sourced from {[}39{]}. These data included, for example, notifications
of both test positive COVID-19 cases and COVID-19 linked deaths and the
number of COVID-19 tests conducted over time. From November 26 2020 on
we displayed weekly small reports with a visualisation of past forecasts
and scores on our website, epiforecasts.io.

Forecasts were stored in a Google Sheet and downloaded, cleaned and
processed every week for submission to the Forecast Hub. If a forecaster
had submitted multiple predictions for a single target, only the latest
submission was kept. Information on the chosen distribution as well as
the parameters for median and width were used to obtain the required set
of 23 quantiles from that distribution. Forecasts from all forecasters
were then aggregated using an unweighted quantile-wise mean (i.e., the
\(\alpha\)-quantile of the ensemble is given by the mean of all
submitted \(\alpha\)-quantiles). To avoid issues with users trying out
the app and submitting a random forecast, we required that a forecaster
needed to make a forecast for at least two targets for a given forecast
in order to be included in the crowd forecast ensemble. On a few
occasions we deleted forecasts that were clearly the result of a user or
software error (such as for example forecasts that were zero
everywhere).

individual forecasts were assessed as clearly erroneous by visual
inspection and subsequently removed before aggregation and were excluded
from the submission as well as the analysis.

Participants were recruited mostly within the Centre of Mathematical
Modeling of Infectious Diseases at the London School of Hygiene \&
Tropical Medicine, but participants were also invited personally or via
social media to submit predictions. Depending on whether they had a
background in either statistics, forecasting or epidemiology,
participants were asked to self-identify as `experts' or `non-experts'.

The study was approved by the Observational / Interventions Research
Ethics Committee at the London School of Hygiene \& Tropical Medicine,
LSHTM Ethics Reference: 22290. Formal consent from participants was
taken in written form through the web application.

\hypertarget{model-based-forecasts}{%
\subsection{Model-based forecasts}\label{model-based-forecasts}}

We used two Bayesian semi-mechanistic models from the \texttt{EpiNow2} R
package (version 1.3.3) as our model-based forecasts {[}40{]}. The first
of these models, here called ``renewal model'', used the renewal
equation {[}41{]} to predict reported cases and deaths (see details in
the SI). It estimated the effective reproduction number \(R_t\) (the
average number of people each person infected at time t is expected to
infect in turn) and modelled future infections as a weighted sum of past
infection multiplied by \(R_t\). \(R_t\) was assumed to stay constant
beyond the forecast date, roughly corresponding to continuing the latest
exponential trend in infections. On the 9th of November we altered the
date when \(R_t\) was assumed to be constant from two weeks prior to the
date of the forecast to the forecast date, which we found to yield a
more stable \(R_t\) estimate. Reported case and death notifications were
obtained by convolving predicted infections over data-based delay
distributions {[}40,42--44{]} to model the time between infection and
report date. The renewal model was used to predict cases as well as
deaths with forecasts being generated for each target separately. Death
forecasts from the renewal model were therefore not informed by past
cases. One submission of the renewal model on December 28th 2020 was
delayed and therefore not included in the official Forecast hub
ensemble.

The second model (``convolution model'', see details in SI) was only
used to forecast deaths and was added later, starting December 7th 2020
(with the first forecast from December 7th suffering from a software bug
and therefore disregarded in all further analyses). The convolution
model was submitted, but never included in the official Forecast hub
ensemble due to concerns that it could be too similar to the renewal
model. The convolution model predicted deaths as a fraction of infected
people who would die with some delay, by using a convolution of reported
cases with a distribution that described the delay from case report to
death and a scaling factor (the case-fatality ratio). Both the renewal
and the convolution model used daily observations and assumed a negative
binomial observation model with a multiplicative day-of-the-week effect
{[}40{]}.

Line list data used to inform the prior for the delay from symptom onset
to test positive case report or death in the model-based forecasts was
sourced from {[}45{]} with data available up to the 1st of August. All
model fitting was done using Markov-chain Monte Carlo (MCMC) in stan
{[}46{]} with each location and forecast target being fitted separately.

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

For the main analysis we focused mostly on two week ahead forecasts, as
COVID-19 forecasts, especially for cases, were in the past found to have
poor predictive performance beyond this horizon {[}15{]}. Forecasts for
cases were scored using the full period from October 2020 until March
2021. To ensure comparability between models, all death forecasts were
scored using only the period from December 14th on, where all models
including the convolution model were available. To ensure robustness of
our results we conducted a sensitivity analysis where all forecasts
(including cases) were scored only over the later period for which all
forecasts were available (see Section \ref{sensitivity-analysis} in the
SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The
weighted interval score (WIS) {[}47{]}, the absolute error, relative
bias, and empirical coverage of the 50\% and 90\% prediction intervals.
The WIS is a proper scoring rule {[}48{]}, meaning that in expectation
the score is optimised by reporting a predictive distribution that is
identical to the true data-generating distribution. Forecasters are
therefore incentivised to report their true belief about the future. The
WIS can be understood as a generalisation of the absolute error to
quantile-based forecasts (also meaning that smaller values are better)
and can be decomposed into three separate penalties: forecast spread
(i.e.~uncertainty of forecasts), over-prediction and under-prediction.
While the over- and under-prediction components of the WIS capture the
amount of over-prediction and under-prediction in absolute terms, we
also look at a relative tendency to make biased forecasts. The bias
metric {[}9{]} we use captures how much probability mass of the forecast
was above or below the true value (mapped to values between -1 and 1)
and therefore represents a general tendency to over- or under-predict in
relative terms. A value of -1 implies that all quantiles of the
predictive distribution are below the observed value and a value of 1
that all quantiles are above the observed value. Empirical coverage is
the percentage of observed values that fall inside a given prediction
interval (e.g.~how many observed values fall inside all 50\% prediction
intervals). Scoring metrics are explained in more detail in Table
\ref{tab:scoring-metrics} in the SI. All scores were calculated using
the \texttt{scoringutils} R package {[}49{]}.

At all stages of the evaluation our forecasts were compared to the
median ensemble of all \emph{other} models submitted to the German and
Polish Forecast Hub (``Hub ensemble''). This ``Hub ensemble'' was
retrospectively computed and excludes all our models, leaving on average
five ensemble member models (see details in Section
\ref{contributions-hub}, as well as in Table
\ref{tab:table-ensemble-versions} and Figure
\ref{fig:num-ensemble-members} in the SI). What we call ``Hub ensemble''
in this article therefore differs from the ``official Hub ensemble''
(here called ``hub-ensemble-realised'') which included crowd forecasts
as well as renewal model forecasts. To enhance interpretability of
scores we mainly report WIS relative to the Hub ensemble in the main
text, i.e.~we divided the average scores for a given model by the
average score achieved by the Hub ensemble on the same set of forecasts
(with values \textgreater1 implying worse and values \textless1 implying
better performance than the Hub ensemble). In addition to comparing our
forecasts against the hub ensemble excluding our models, we also
assessed the impact of our forecasts on the performance of the
forecasting hub by recalculating separate versions of the Hub ensemble
with only some (or all) of our forecasts included. Versions that
included either all of our models (``hub-ensemble-with-all'') or only
one of them (``hub-ensemble-with-X'') were computed retrospectively.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{crowd-forecast-participation}{%
\subsection{Crowd forecast
participation}\label{crowd-forecast-participation}}

A total number of 32 participants submitted forecasts, 17 of those
self-identified as `expert' in either forecasting or epidemiology. The
median number of forecasters for any given forecast target was 6, the
minimum 2 and the maximum 10. The mean number of submissions from an
individual forecaster was 4.7 but the median number was only one - most
participants dropped out after their first submission. Only two
participants submitted a forecast every single week, both of whom are
authors on this study.

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-all-v4} \caption{Visualisation of aggregate performance metrics for forecasts one to four weeks into the future. A, B: mean weighted interval score (WIS, lower indicates better performance) across horizons. WIS is decomposed into its components dispersion, over-prediction and under-prediction. C: Empirical coverage of the 50\% prediction intervals (50\% coverage is perfect). D: Empirical coverage of the 90\% prediction intervals. E: Dispersion (same as in panel A, B). Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast (lower is better). H. Standard deviation of all WIS values for different horizons}\label{fig:agg-performance-all}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/figure-forecasts-2} \caption{Two week ahead forecasts and corresponding scores. A, C: Visualisation of 50\% prediction intervals of two week ahead forecasts against the reported values. Forecasts that were not scored (because there was no complete set of death forecasts available) are greyed out. B, D: Visualisation of corresponding WIS.}\label{fig:forecasts-and-truth}
\end{figure}

\hypertarget{case-forecasts}{%
\subsection{Case Forecasts}\label{case-forecasts}}

For cases, crowd forecasts had a lower mean weighted interval score
(WIS, lower values indicate better performance) than both the renewal
model and the Hub ensemble across all forecast horizons (Figure
\ref{fig:agg-performance-all}A) and locations (Figure
\ref{fig:performance-locations-rel}A). For two week ahead forecasts,
mean WIS relative to the Hub ensemble (= 1) was 0.89 for crowd forecasts
and 1.40 for the renewal model (Table \ref{tab:score-table-2}). Across
all forecasting approaches, locations and forecast horizons, the
distribution of WIS values was very right-skewed, and average
performance was heavily influenced by outliers (Figure
\ref{fig:distribution-scores}). Overall, low variance in forecast
performance was closely linked with good mean performance (Figures
\ref{fig:agg-performance-all}H and and \ref{fig:agg-performance-all}A),
suggesting that the ability to avoid large errors was an important
factor in determining overall performance. The impact of outlier values
was especially pronounced for the renewal model, which had more outliers
(Figure \ref{fig:distribution-scores}), as well as the highest standard
deviation of WIS values (standard deviation of the WIS relative to the
WIS sd of the Hub ensemble was 1.54 at the two weeks ahead horizon),
while the ensemble of crowd forecasts (rel. WIS sd 0.76) and the Hub
ensemble (= 1) showed more stable performance.

To varying degrees, all forecasts exhibited trend-following behaviour
and were rarely able to predict a change in trend before it had
happened. For example, all forecasts failed to predict the change in
trend from increase to decrease that happened in November in Germany and
severely overshot reported cases (Figure
\ref{fig:forecasts-and-truth}A). This was most striking for the renewal
model, which extrapolated unconstrained exponential growth based on the
recent past of observations. The Hub ensemble and the crowd forecast,
which had both been under-predicting throughout October, also failed to
predict the change in trend after cases peaked, but less severely so.
Human forecasters, possibly aware of the semi-lockdown announced on
November 2nd 2020 {[}50{]} and the change in the testing regime (with
stricter test criteria) on November 11th 2020 {[}36{]}, were fastest to
adapt to the new trend, and the Hub ensemble slowest. In December, cases
rose again in Germany, with all models under-predicting this growth to
varying extents. As in October, the renewal model captured the phase of
exponential growth in cases slightly better than other approaches, but
again overshot when reported case numbers fell over Christmas. The large
variance in predictions in January in Germany (severe under-prediction
followed by severe over-prediction) may in part be caused by the fact
that the renewal model operated on daily data and therefore was
susceptible to fluctuations in daily reporting around Christmas that
would not have influenced on weekly reporting. Similar trends in
performance were evident in Poland, with the crowd forecast quickest at
adapting to the change in trend in November. In general, there were
fewer large outlier forecasts in Poland and in particular the renewal
model performed more in line with other forecasts there.

All forecasting approaches, including the Hub ensemble, were
overconfident, i.e.~they showed lower than nominal coverage (meaning
that 50\% (90\%) prediction intervals generally covered less than 50\%
(90\%) of the actually observed values) (Figure
\ref{fig:agg-performance-all}C and \ref{fig:agg-performance-all}D).
Coverage for all forecasts deteriorated with increasing forecast
horizon, indicating that all forecasting approaches struggled to
quantify uncertainty appropriately for case forecasts. This was
especially an issue for crowd forecasts, which had markedly shorter
prediction intervals (i.e., narrower and more confident predictive
distributions) than other approaches (Figure
\ref{fig:agg-performance-all}E) and only showed a small increase in
uncertainty across forecast horizons. The crowd forecasts prediction
intervals were also noticeably narrower than the default baseline shown
to forecasters in the application (see Figure
\ref{fig:compare-forecasters}).

In spite of good performance in terms of the absolute error (Figure
\ref{fig:agg-performance-all}G), the narrow forecast intervals led to
forecasts which were severely overconfident (covering only 36\% and 55\%
of all observations with the 50\% and 90\% prediction intervals of all
forecasts made at a two week forecast horizon, and only 5\% and 38\%
four weeks ahead) (Figure \ref{fig:agg-performance-all}C,D and Tables
\ref{tab:score-table-2} and \ref{tab:score-table-4}). Despite worse
performance in terms of absolute error (Figure
\ref{fig:agg-performance-all}G), the renewal model achieved better
calibration (comparable to the Hub ensemble), as uncertainty increased
rapidly across forecast horizons. The crowd forecasts, on the other
hand, showed a smaller bias than the renewal model, but were
overconfident.

The renewal model exhibited a noticeable tendency towards
over-predicting reported cases across all horizons. The crowd forecast
tended to over-predict at longer forecast horizons, whereas the Hub
ensemble showed no systematic bias (Figure
\ref{fig:agg-performance-all}F). Regardless of a general relative
tendency to over-predict, all forecasting approaches incurred larger
absolute penalties from over- than from under-prediction (see
decomposition of the WIS into absolute penalties for over-prediction,
under-prediction and dispersion in Figures
\ref{fig:agg-performance-all}A and \ref{fig:agg-performance-all}B and
Tables \textbackslash ref\{tab:score-table-2) and
\ref{tab:score-table-4}).

Generally, trends in overall performance were broadly similar across
locations (Figures \textbackslash ref\{fig:performance-locations) and
\ref{fig:performance-locations-rel}). Due to the differing population
sizes and numbers of notifications in Germany and Poland absolute scores
were difficult to compare directly. However, relative to the Hub
ensemble, the crowd forecasts performed noticeably better in Germany
than in Poland and the renewal model better in Poland than in Germany
(Figures \ref{fig:performance-locations-rel}A,
\ref{fig:performance-locations-rel}G,
\ref{fig:agg-performance-all-Germany},
\ref{fig:agg-performance-all-Poland}).

\hypertarget{death-forecasts}{%
\subsection{Death Forecasts}\label{death-forecasts}}

For deaths, the Hub ensemble outperformed the crowd forecasts as well as
our model-based approaches across all forecast horizons and locations
(Figure \ref{fig:agg-performance-all}B, Figure
\ref{fig:performance-locations}B). Relative WIS values for the models
two weeks ahead were 1.22 (convolution model), 1.26 (crowd forecast), 1
(Hub ensemble) and 1.79 (renewal model). The crowd forecasts performed
better than the renewal model across all forecast horizons and locations
(Figure \ref{fig:agg-performance-all}B, Figure
\ref{fig:performance-locations}B), and also better than the convolution
model three and four weeks ahead. Poor performance of the renewal model,
especially at longer horizons, indicates that an approach that does not
know about past cases, but instead estimates and projects a separate
\(R_t\) trace from deaths, does not use the available information
efficiently. The convolution model was able to outperform both the
renewal model and the crowd forecasts at shorter forecast horizons
(where the delay between cases and deaths means that future deaths are
largely informed by present cases), but saw performance deteriorate at
three and four weeks ahead (where case predictions from the renewal
model were increasingly used to inform death predictions) (Figure
\ref{fig:agg-performance-all}B, Table \ref{tab:score-table-4}).

As past cases and hospitalisations can be used as predictors, predicting
a change in trend may be easier for deaths than for cases. Even though
all forecasts generally struggled with this, there were some instances
where changing trends were well captured or even anticipated. In Poland,
for example, the Hub ensemble was able to capture or even anticipate the
peak in deaths in December quite well (whereas the renewal model and
crowd forecast did not). The renewal model, which mostly exhibited
trend-following behaviour, correctly predicted another increase in
weekly deaths in mid-January (potentially based on changes in daily
deaths, as the renewal model did not know about past cases). In Germany
in early January, all models predicted a decrease in deaths two to three
weeks before it actually happened. Predictions from the renewal model at
that time were likely strongly influenced by an unexpected drop in
reported deaths in December. The other forecasting approaches and in
particular, the convolution model may have been affected by potentially
under-reported case numbers around Christmas. When the decrease that all
models had predicted to happen in early January failed to materialise,
the renewal model and the crowd forecast noticeably over-corrected and
over-predicted deaths in the following weeks, while the Hub ensemble,
and to a slightly lesser degree, the convolution model were able to
capture the downturn well when it finally happened at the end of
January.

Death forecasts, generally, showed greater coverage of the 50\% and 90\%
prediction intervals than case forecasts and no decrease in coverage
across forecast horizons, indicating that it might be easier to
appropriately quantify uncertainty for death forecasts. The Hub ensemble
had the greatest coverage, with empirical coverage of the 50\% and 90\%
prediction intervals exceeding 50\%, and 90\%, respectively, across all
forecast horizons. Coverage for the crowd forecasts and our model-based
approaches was generally lower than that of the Hub ensemble and mostly
slightly lower than nominal coverage (Figure
\ref{fig:agg-performance-all}C and \ref{fig:agg-performance-all}D). As
for cases, the crowd forecast tended to have the narrowest prediction
intervals and uncertainty increased most slowly across forecast
horizons, and the renewal model forecasts generally were widest. The
convolution model had relatively narrow prediction intervals for short
forecast horizons, but had rapidly (and non-linearly) increasing
uncertainty for longer forecast horizons, driven by increasing
uncertainty in the underlying case forecasts.

For deaths, the ensemble of crowd forecasts had a consistent tendency to
over-predict \ref{fig:agg-performance-all}F. The convolution model had a
strong tendency to under-predict, with the magnitude of under-prediction
steadily decreasing for longer forecast horizons. The renewal model
(which over-predicted for cases) and the Hub ensemble slightly tended
towards under-prediction. For deaths, absolute over- and
under-prediction penalties were more in line with a general relative
tendency to over- or under-predict than for cases (Figure
\ref{fig:agg-performance-all}A, \ref{fig:agg-performance-all}B and
Tables \ref{tab:score-table-2}, \ref{tab:score-table-4}).

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/distribution_scores_wis-2} \caption{Distribution of scores. A: Distribution of weighted interval scores for two week ahead forecasts of the different models and forecast targets. Points denote single forecasts scores, while the shaded area shows an estimated probability density. B: Distribution of WIS separate by country. Black squares indicate median and black circles mean scores.}\label{fig:distribution-scores}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-rel-ensemble-v4} \caption{Relative aggregate performance metrics across forecast horizons for different versions of the Hub median ensemble. “Hub-ensemble”    extit{excludes} all our models, Hub-ensemble-all    extit{includes} all of our models, “Hub-ensemble-realised” is the actual hub-ensemble observed in reality, which includes the renewal model and the crowd forecasts, but not the convolution model. A, B: mean weighted interval score (WIS) across horizons relative to the Hub ensemble (lower values indicate better performance). C, D: Empirical coverage of the 50\% and 90\% prediction intervals minus empirical coverage observed for the Hub ensemble. E: Dispersion relative to the dispersion of the Hub ensemble. Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- orunderpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast relative to the Hub ensemble. H. Standard deviation of all WIS values for different horizons relative to the Hub ensemble.}\label{fig:agg-performance-ensemble-rel}
\end{figure}

\hypertarget{contributions-hub}{%
\subsection{Contribution to the Forecast Hub}\label{contributions-hub}}

Of our three models, only the renewal model and the crowd forecast were
included in the official Forecast Hub median ensemble
(``hub-ensemble-realised''), while the convolution model was never
included as it was deemed too similar to the existing renewal model. In
the official Hub ensemble, there were on average 7.1 models included
(including our own), with a median of 7, a minimum of 4 (on 28 December
2020 over the Christmas period) and a maximum of 10. Versions that
included either all of our models (``hub-ensemble-with-all'') or only
one of them (``hub-ensemble-with-X'') were computed retrospectively. An
overview of all models and ensemble versions is shown in Table
\textbackslash ref\{tab:table-ensemble-versions) in the SI.

For cases, our contributions (compared to the Hub ensemble without our
contributions) consistently improved performance across all forecasting
horizons (rel. WIS 0.9 two weeks ahead, Table
\ref{tab:score-table-ensemble-2}). Contributions from the crowd
forecasts alone also improved performance of the Hub ensemble across all
forecast horizons, while contributions from the renewal model had a
negative effect for longer horizons (rel. WIS 1.02 three weeks ahead,
1.06 four weeks ahead). The realised ensemble including both models
performed better or equal compared to all versions with only one model
included for up to three weeks ahead, suggesting synergistic effects.
Only for predictions four weeks ahead would removing the renewal model
have improved performance (Table \ref{tab:score-table-ensemble-4}). The
realised ensemble performed comparably to the crowd forecasts for
predictions one to two weeks ahead, and worse for greater forecast
horizons.

For deaths, contributions from the renewal model and crowd forecast
together improved performance only for one week ahead predictions and
showed an increasingly negative impact on performance for longer
horizons (rel. WIS of the Hub-ensemble-realised 1.01 two weeks ahead,
1.05 four weeks ahead, Tables \ref{tab:score-table-ensemble-2} and
\ref{tab:score-table-ensemble-4}). Individual contributions from both
the renewal model and the crowd forecast were largely negative, while a
version of the Hub ensemble with only the convolution model included
would have performed consistently better across all forecast horizons
(with the positive impact increasing for longer horizons). This is
especially interesting as the convolution model performed consistently
worse than the pre-existing Hub ensemble (Figure
\ref{fig:agg-performance-all}) and especially worse for longer horizons.

We also considered the impact of our contributions on a version of the
Hub ensemble constructed by taking the quantile-wise mean, rather than
the median. General trends were similar, with the notable exception of
the convolution model, which had a consistently positive impact on the
median ensemble, but a mixed and mostly slightly negative impact on the
mean ensemble (Figures \ref{fig:agg-performance-ensemble-rel}B and
\ref{fig:agg-performance-ensemble-mean}B). This may happen if a model is
more correct directionally relative to the pre-existing ensemble, but
overshoots in absolute terms, thereby moving the ensemble too far. For
both the mean and the median ensemble, changes in performance from
adding or removing models were of a similar order of magnitude,
suggesting that at least in this instance, with a relatively small
ensemble size, the median ensemble was not necessarily more `robust' to
changes than the mean ensemble. However, the ensemble version with all
our forecasts included (``hub-ensemble-with-all'') tended to perform
relatively better for the median ensemble than the mean ensemble,
suggesting that adding more models may be more beneficial or `safer' for
the median than for the mean ensemble as directional errors can more
easily cancel out than errors in absolute terms.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Epidemiological forecasting modelling combines knowledge about
infectious disease dynamics with the subjective opinion of the
researcher who develops and refines the model. In this study, we
compared forecasts of cases of and deaths from COVID-19 in Germany and
Poland based purely on human judgement and elicited from a crowd of
researchers and volunteers against forecasts from two semi-mechanistic
epidemiological models. In spite of the small number of participants and
a general tendency to be overconfident, crowd forecasts consistently
outperformed our epidemiological models as well as the Hub ensemble when
forecasting cases but not when forecasting deaths. This suggests that
humans might be relatively good at foreseeing trends that are hard to
model but may struggle to form an intuition for the exact relationship
between cases and deaths.

Past studies have evaluated the performance of model-based forecasting
approaches as well as human experts and non-experts in various contexts.
However, most of these studies either focused only on the evaluation of
(expert-tuned) model-based approaches {[}e.g. 12,13,14{]}, or
exclusively on human forecasts {[}19,20,24,25{]}. In contrast, we
directly compared human and model-based forecasts. This is similar to
the approach taken by {[}11{]}, but extends it in several ways. While
Farrow et al.~only asked for point predictions and constructed a
predictive distribution from these, we asked participants to provide a
full predictive distribution, allowing us to compare human forecasts and
models without any further assumptions, as well as to analyse how humans
quantified their uncertainty. In addition, we compared crowd forecasts
to two semi-mechanistic models informed by basic epidemiological
knowledge of COVID-19, allowing us to assess not only relative
performance but also to analyse qualitative differences between human
judgement and model-based insight. In terms of interpretability of the
results, exact knowledge of our two models, as well as focus on a
limited set of targets and locations was a major advantage of our study
compared to larger studies conducted by the Forecast Hubs
{[}12--15,17{]}.

The strong performance of crowd forecasts in our study is in line with
results from Farrow et al.~who also report strong performance of human
predictions in past Flu challenges despite difficulties to recruit a
large number of participants. The advantage of crowd forecasts we
observed over our semi-mechanistic models is likely in part explained by
the fact that we compared an ensemble of crowd forecasts with single
models. However, this probably explains only part of the difference, and
performance relative to the Hub ensemble strongly suggests that human
insight is valuable when forecasting highly volatile and potentially
hard-to-predict quantities such as case numbers. One potential
explanation is that humans can have access to data that is not available
to or hard to integrate into model-based forecasts. Relatively good
performance of our semi-mechanistic models short-term, but not
longer-term, suggests that model-based forecasts are helpful to
extrapolate from current conditions, but require some form of human
intervention or additional assumptions to inform forecasts when
conditions change over time. This human intervention may be particularly
important when dealing with artefacts in reporting and data anomalies
(and especially when using daily, rather than weekly data). The large
variance in predictions in January in Germany for example (severe
under-prediction followed by severe over-prediction, see Figure
\ref{fig:forecasts-and-truth}A), may in part be caused by the fact that
the renewal model operated on daily data and therefore was susceptible
to fluctuations in daily reporting which have less of an influence on
weekly reporting.

Our results suggest that human intervention may be less beneficial when
forecasting deaths (especially at shorter horizons, when deaths are
largely dependent on already observed cases), which benefits from the
ability to model the delays and exact epidemiological relationships
between different leading and lagged indicators. Relatively good
performance of the convolution model, especially compared to the poor
performance of the renewal model on deaths (which used only deaths to
estimate and predict the effective reproduction number) underlines the
importance of including leading indicators such as cases as a predictor
for deaths.

Given the low number of participants in our study, it is difficult to
generalise conclusions about crowd predictions to other settings. Using
R shiny as a platform for the web application arguably created some
limits to user experience and performance, influencing the number of
participants and potentially creating a self-selection effect.
Motivating forecasters to contribute regularly proved challenging,
especially given that the majority of our participants were from the UK
and may not have been familiar with all relevant details of the
situation in Germany and Poland. On the other hand, R shiny facilitated
quick development and allowed us to provide our crowd forecasting
tooling as an open source R package, meaning that it is available for
others to use, for example in settings like early-stage outbreaks where
model-based forecasts are not available. In light of the relatively
small number of Hub ensemble models, performance of the Hub ensemble is
also difficult to generalise. More research is needed to replicate these
findings and investigate how crowd forecasts compare against the types
of models and model ensembles policy makers use to inform their
decisions.

Our work suggests that crowd forecasts and model-based forecasts could
have different strengths and may be able to complement each other. When
choosing a suitable approach for a given task it is important to take
into account how the output will be used. In this work we focused on
forecasts (which aim to predict future data points whilst accounting for
all factors that might influence them), whereas policy makers might be
more interested in projections (which show what would happen in the
absence of any events that could change the trend) or scenario
modelling. Forecasts may not be a suitable basis for informing policy
decisions, if forecasters already have factored in the expectation of a
future intervention. Model-based approaches can be either forecasts or
projections depending on the assumptions, whereas eliciting projections
that are not influenced by implicit assumptions about the future from
humans may be harder.

Further work should explore the effects of humans refining their
mathematical models or changing model outputs in more detail.
Model-based forecasts could be used as an input to human judgement, with
researchers adjusting predictions generated by models. Seeing a
model-based forecast could help humans calibrate uncertainty better,
while allowing for manual intervention to adapt spurious trend
predictions. Tools need to be developed to facilitate this process at a
larger scale. Human insight could also be used as an input to models.
Such a `hybrid' forecasting approach could for example ask humans to
predict the trend of the effective reproduction number \(R_t\) or the
doubling rate (i.e.~how the epidemic evolves) into the future and use
this to estimate the exact number of cases, hospitalisations or deaths
this would imply. In light of severe overconfidence, yet good
performance in terms of the absolute error, post-processing of human
forecasts to adjust and widen prediction intervals may be another
promising approach. Crowd forecasting in general could benefit greatly
from the availability of tools suitable to appeal to a greater audience.
Given the good performance we and previous authors observed in spite of
the limited resources available and the small number of participants,
this seems worthwhile to further develop and explore.

\clearpage

\textbf{Acknowledgements}

NIB received funding from the Health Protection Research Unit (grant
code NIHR200908, https://www.nihr.ac.uk/). SA's work was funded by the
Wellcome Trust (grant: 210758/Z/18/Z, https://wellcome.org/). The work
of JB was supported by the Helmholtz Foundation
(https://www.helmholtz.de/) via the SIMCARD Information and Data Science
Pilot Project. This research was partly funded by the National Institute
for Health Research (NIHR, https://www.nihr.ac.uk/) (16/137/109 \&
16/136/46) using UK aid from the UK Government to support global health
research. The views expressed in this publication are those of the
author(s) and not necessarily those of the NIHR or the UK Department of
Health and Social Care. BJQ is supported in part by a grant from the
Bill and Melinda Gates Foundation (OPP1139859,
https://www.gatesfoundation.org/). EvL acknowledges funding by the
National Institute for Health Research (NIHR) Health Protection Research
Unit (HPRU) in Modelling and Health Economics (grant number NIHR200908,
https://www.nihr.ac.uk/) and the European Union's Horizon 2020 research
and innovation programme - project EpiPose (101003688,
https://ec.europa.eu/programmes/horizon2020/). AC acknowledges funding
by the NIHR, the Sergei Brin foundation, USAID (https://www.usaid.gov/),
and the Academy of Medical Sciences (https://acmedsci.ac.uk/). SF's work
was supported by the Wellcome Trust (grant: 210758/Z/18/Z,
https://wellcome.org/), and the NIHR (NIHR200908,
https://www.nihr.ac.uk/).

We thank all forecasters who participated in this study for their
contribution.

We would also like to acknowledge (in a randomised order) the members of
Centre for the Mathematical Modelling of Infectious Diseases COVID-19
Working Group at the the London School of Hygiene \& Tropical Medicine:
Oliver Brady, Katharine Sherratt, Kaja Abbas, Kerry LM Wong, Charlie
Diamond, Katherine E. Atkins, Rein M G J Houben, Jiayao Lei, Rachel
Lowe, David Simons, Sophie R Meakin, Nicholas G. Davies, Timothy W
Russell, Kevin van Zandvoort, Quentin J Leclerc, Kathleen O'Reilly,
Stéphane Hué, Alicia Rosello, Emilie Finch, C Julian Villabona-Arenas,
Thibaut Jombart, W John Edmunds, Yalda Jafari, Jack Williams, Alicia
Showering, Damien C Tully, Jon C Emery, Carl A B Pearson, David Hodgson,
Frank G Sandmann, Petra Klepac, Adam J Kucharski, Graham Medley, Yang
Liu, Simon R Procter, Emily S Nightingale, William Waites, Rosanna C
Barnard, Joel Hellewell, Yung-Wai Desmond Chan, Fiona Yueqian Sun,
Hamish P Gibbs, Rosalind M Eggo, Lloyd A C Chapman, Stefan Flasche,
James W Rudge, Akira Endo, Naomi R Waterlow, Paul Mee, James D Munday,
Ciara V McCarthy, Mihaly Koltai, Amy Gimma, Christopher I Jarvis, Megan
Auzenbergs, Matthew Quaife, Fabienne Krauer, Samuel Clifford, Georgia R
Gore-Langton, Arminder K Deol, Kiesha Prem, Gwenan M Knight, Rachael
Pung, Anna M Foss.

\clearpage

\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\hypertarget{appendix-supplementary-information}{%
\section*{(APPENDIX) Supplementary
information}\label{appendix-supplementary-information}}
\addcontentsline{toc}{section}{(APPENDIX) Supplementary information}

\hypertarget{supplementary-information}{%
\section{Supplementary information}\label{supplementary-information}}

\hypertarget{scoring-metrics-used}{%
\subsection{Scoring metrics used}\label{scoring-metrics-used}}

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{9.3cm}}
\caption{\label{tab:scoring-metrics}Overview of the scoring metrics used.}\\
\toprule
Metric & Explanation\\
\midrule
\endfirsthead
\caption[]{Overview of the scoring metrics used. \textit{(continued)}}\\
\toprule
Metric & Explanation\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
WIS (Weighted) interval score & The weighted interval score (smaller values are better) is a proper scoring rule for quantile forecasts. It converges to the continuos ranked probability score (which itself is a generalisation of the absolute error to probabilistic forecasts) for an increasing number of intervals. The score can be decomposed into a dispersion (uncertainty) component and penalties for over- and underprediction. For a single interval, the score is computed as 
  $$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u), $$ 
  where $1()$ is the indicator function, $y$ is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e. the lower and upper bound of a single prediction interval. For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum, 
  $$WIS = \frac{1}{K + 0.5} \cdot \left( w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y) \right), $$
  where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.

\cellcolor{gray!6}{Its proximity to the absolute error means that when averaging across multiple targets (e.g. different weeks), it will be dominated by targets with higher absolute values.}\\
\addlinespace \addlinespace
Interval coverage & Interval coverage is a measure of marginal calibration and indicates the proportion of observed values that fall in a given prediction interval range. Nominal coverage represents the percentage of observed values that should ideally be covered (e.g. we would like a 50 percent prediction interval to cover on average 50 percent of the observations), while empirical coverage is the actual percentage of observations covered by a certain prediction interval.\\
\addlinespace \addlinespace
Bias & (Relative) bias is a measure of the general tendency of a forecaster to over- or underpredict. Values are between -1 and 1 and 0 ideally. For continuous forecasts, bias is given as 
$$B(F, y) = 1 - 2 \cdot (F (y)), $$ 
where $F$ is the CDF of the predictive distribution and $y$ is the observed value.

For quantile forecasts, $F(y)$ is replaced by a quantile rank. The appropriate quantile rank is determined by whether the median forecast is below  or above the true value. We then take the innermost quantile rank for which the quantile is still larger (under-prediction) or smaller (over-prediction) than the observed value.

\cellcolor{gray!6}{In contrast to the over- and underprediction penalties of the interval score it is bound between 0 and 1 and represents a general tendency of forecasts to be biased rather than the absolute amount of over- and underprediction. It is therefore a more robust measurement.}\\*
\end{longtable}

\clearpage

\hypertarget{the-crowdforecasting-app}{%
\subsection{The crowdforecasting app}\label{the-crowdforecasting-app}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../crowd-forecast/Screenshot-forecasting-app} \caption{Screenshot of the crowdforecasting app used to elicit predictions (made in June 2021). }\label{fig:screenshot}
\end{figure}

\clearpage

\hypertarget{text-s1.-further-details-on-the-semi-mechanistic-forecasting-models}{%
\subsection{Text S1. Further details on the semi-mechanistic forecasting
models}\label{text-s1.-further-details-on-the-semi-mechanistic-forecasting-models}}

\hypertarget{renewal-equation-model}{%
\subsubsection{Renewal equation model}\label{renewal-equation-model}}

The model was initialised prior to the first observed data point by
assuming constant exponential growth for the mean of assumed delays from
infection to case report.

\begin{align}
  I_{t} &= I_0 \exp  \left(r t \right)  \\
  I_0 &\sim \mathcal{LN}(\log I_{obs}, 0.2) \\
  r &\sim \mathcal{LN}(r_{obs}, 0.2) 
\end{align}

Where \(I_{obs}\) and \(r_{obs}\) are estimated from the first week of
observed data. For the time window of the observed data infections were
then modelled by weighting previous infections by the generation time
and scaling by the instantaneous reproduction number. These infections
were then convolved to cases by date (\(O_t\)) and cases by date of
report (\(D_t\)) using log-normal delay distributions. This model can be
defined mathematically as follows,

\begin{align}
  \log R_{t} &= \log R_{t-1} + \mathrm{GP}_t \\
  I_t &= R_t \sum_{\tau = 1}^{15} w(\tau | \mu_{w}, \sigma_{w}) I_{t - \tau} \\
  O_t &= \sum_{\tau = 0}^{15} \xi_{O}(\tau | \mu_{\xi_{O}}, \sigma_{\xi_{O}}) I_{t-\tau} \\
  D_t &= \alpha \sum_{\tau = 0}^{15} \xi_{D}(\tau | \mu_{\xi_{D}}, \sigma_{\xi_{D}}) O_{t-\tau} \\ 
  C_t &\sim \mathrm{NB}\left(\omega_{(t \mod 7)}D_t, \phi\right)
\end{align}

Where, \begin{align}
     w &\sim \mathcal{G}(\mu_{w}, \sigma_{w}) \\
    \xi_{O} &\sim \mathcal{LN}(\mu_{\xi_{O}}, \sigma_{\xi_{O}}) \\
    \xi_{D} &\sim \mathcal{LN}(\mu_{\xi_{D}}, \sigma_{\xi_{D}}) 
\end{align}

This model used the following priors for cases,

\begin{align}
     R_0 &\sim \mathcal{LN}(0.079, 0.18) \\
    \mu_w &\sim \mathcal{N}(3.6, 0.7) \\
    \sigma_w &\sim \mathcal{N}(3.1, 0.8) \\
    \mu_{\xi_{O}} &\sim \mathcal{N}(1.62, 0.064) \\
    \sigma_{\xi_{O}} &\sim \mathcal{N}(0.418, 0.069) \\
    \mu_{\xi_{D}} &\sim \mathcal{N}(0.614, 0.066) \\
    \sigma_{\xi_{D}} &\sim \mathcal{N}(1.51, 0.048) \\
    \alpha &\sim \mathcal{N}(0.25, 0.05) \\
    \frac{\omega}{7} &\sim \mathrm{Dirichlet}(1, 1, 1, 1, 1, 1, 1) \\
    \phi &\sim \frac{1}{\sqrt{\mathcal{N}(0, 1)}}
\end{align}

and updated the reporting process as follows when forecasting deaths,

\begin{align}
    \mu_{\xi_{D}} &\sim \mathcal{N}(2.29, 0.076) \\
    \sigma_{\xi_{D}} &\sim \mathcal{N}(0.76, 0.055) \\
    \alpha &\sim \mathcal{N}(0.005, 0.0025) 
\end{align}

\(\alpha\), \(\mu\), \(\sigma\), and \(\phi\) were truncated to be
greater than 0 and with \(\xi\), and \(w\) normalised to sum to 1.

The prior for the generation time was sourced from {[}51{]} but refit
using a log-normal incubation period with a mean of 5.2 days (SD 1.1)
and SD of 1.52 days (SD 1.1) with this incubation period also being used
as a prior {[}52{]} for \(\xi_{O}\). This resulted in a
gamma-distributed generation time with mean 3.6 days (standard deviation
(SD) 0.7), and SD of 3.1 days (SD 0.8) for all estimates. We estimated
the delay between symptom onset and case report or death required to
convolve latent infections to observations by fitting an integer
adjusted log-normal distribution to 10 subsampled bootstraps of a public
linelist for cases in Germany from April 2020 to June 2020 with each
bootstrap using 1\% or 1769 samples of the available data {[}45,53{]}
and combining the posteriors for the mean and standard deviation of the
log-normal distribution {[}40,42,46,54{]}.

\(GP_t\) is an approximate Hilbert space Gaussian process as defined in
{[}55{]} using a Matern 3/2 kernel using a boundary factor of 1.5 and 17
basis functions (20\% of the number of days used in fitting). The length
scale of the Gaussian process was given a log-normal prior with a mean
of 21 days, and a standard deviation of 7 days truncated to be greater
than 3 days and less than 60 days. The magnitude of the Gaussian process
was assumed to be normally distributed centred at 0 with a standard
deviation of 0.1.

From the forecast time horizon (\(T\)) and onwards the last value of the
Gaussian process was used (hence \(R_t\) was assumed to be fixed) and
latent infections were adjusted to account for the proportion of the
population that was susceptible to infection as follows,

\begin{equation}
    I_t = (N - I^c_{t-1}) \left(1 - \exp \left(\frac{-I'_t}{N - I^c_{T}}\right)\right),
\end{equation}

where \(I^c_t = \sum_{s< t} I_s\) are cumulative infections by \(t-1\)
and \(I'_t\) are the unadjusted infections defined above. This
adjustment is based on that implemented in the \texttt{epidemia} R
package {[}56,57{]}.

\hypertarget{convolution-model}{%
\paragraph{Convolution model}\label{convolution-model}}

The convolution model shares the same observation model as the renewal
model but rather than assuming that an observation is predicted by
itself using the renewal equation instead assumes that it is predicted
entirely by another observation after some parametric delay. It can be
defined mathematically as follows,

\begin{equation} 
    D_{t} \sim \mathrm{NB}\left(\omega_{(t \mod 7)} \alpha \sum_{\tau = 0}^{30} \xi(\tau | \mu, \sigma) C_{t-\tau},  \phi \right)
\end{equation}

with the following priors,

\begin{align}
    \frac{\omega}{7} &\sim \mathrm{Dirichlet}(1, 1, 1, 1, 1, 1, 1) \\
    \alpha &\sim \mathcal{N}(0.01, 0.02) \\
    \xi &\sim \mathcal{LN}(\mu, \sigma) \\
    \mu &\sim \mathcal{N}(2.5, 0.5) \\
\sigma &\sim \mathcal{N}(0.47, 0.2) \\
\phi &\sim \frac{1}{\sqrt{\mathcal{N}(0, 1)}}
\end{align}

with \(\alpha\), \(\mu\), \(\sigma\), and \(\phi\) truncated to be
greater than 0 and with \(\xi\) normalised such that
\(\sum_{\tau = 0}^{30} \xi(\tau | \mu, \sigma) = 1\).

\hypertarget{model-fitting}{%
\subsubsection{Model fitting}\label{model-fitting}}

Both models were implemented using the \texttt{EpiNow2} R package
(version 1.3.3) {[}40{]}. Each forecast target was fitted independently
for each model using Markov-chain Monte Carlo (MCMC) in stan {[}46{]}. A
minimum of 4 chains were used with a warmup of 250 samples for the
renewal equation-based model and 1000 samples for the convolution model.
2000 samples total post warmup were used for the renewal equation model
and 4000 samples for the convolution model. Different settings were
chosen for each model to optimise compute time contingent on
convergence. Convergence was assessed using the R hat diagnostic
{[}46{]}. For the convolution model forecast the case forecast from the
renewal equation model was used in place of observed cases beyond the
forecast horizon using 1000 posterior samples. 12 weeks of data was used
for both models though only 3 weeks of data were included in the
likelihood for the convolution model.

\clearpage

\hypertarget{tables-with-results-of-the-forecast-evaluation}{%
\subsection{Tables with results of the forecast
evaluation}\label{tables-with-results-of-the-forecast-evaluation}}

\begin{table}[!h]
\caption{\label{tab:score-table-2}Scores for one and two week ahead forecasts (cut to three significant digits and rounded). Note that scores for cases (which include the whole period from October 12th 2020 until March 1st 2021) and deaths (which include only forecasts from the 21st of December 2020 on) are computed on different subsets. Numbers in brackets show the metrics relative to the Hub ensemble (i.e. the median ensemble of all other models submitted to the German and Polish Forecast Hub, excluding our contributions). WIS is the mean weighted interval score (lower values are better), WIS - sd is the standard deviation of all scores achieved by a model. Dispersion, over-prediction and under-prediction together sum up to the weighted interval score. Bias (between -1 and 1, 0 is ideal) represents the general average tendency of a model to over- or underpredict. 50\% and 90\%-coverage are the percentage of observed values that fell within the 50\% and 90\% prediction intervals of a model.\\\hspace{\textwidth}}

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Crowd forecast & 7010 (0.8) & 7480 (0.64) & 2680 (0.73) & 1700 (1.38) & 2630 (0.68) & -0.01 & 10400 (0.82) & 0.55 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 8770 (1) & 11700 (1) & 3670 (1) & 1230 (1) & 3870 (1) & -0.04 & 12700 (1) & 0.57 & 0.81\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 1 wk ahead}} & Renewal & 8740 (1) & 11800 (1.01) & 2190 (0.6) & 2720 (2.21) & 3830 (0.99) & 0.18 & 12000 (0.94) & 0.48 & 0.71\\
\cmidrule{1-11}
\cellcolor{white}{} & Crowd forecast & 16200 (0.89) & 16600 (0.76) & 3660 (0.6) & 5930 (1.56) & 6600 (0.78) & -0.01 & 23300 (0.87) & 0.36 & 0.55\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 18300 (1) & 21900 (1) & 6140 (1) & 3800 (1) & 8410 (1) & -0.03 & 26800 (1) & 0.43 & 0.64\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 2 wk ahead}} & Renewal & 25600 (1.4) & 33800 (1.54) & 5420 (0.88) & 5920 (1.56) & 14200 (1.69) & 0.17 & 34600 (1.29) & 0.43 & 0.67\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Convolution & 255 (1.03) & 343 (1.01) & 82 (0.89) & 142 (1.23) & 31.1 (0.75) & -0.18 & 399 (1.19) & 0.42 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 265 (1.07) & 317 (0.94) & 78.2 (0.85) & 82 (0.71) & 105 (2.52) & 0.08 & 402 (1.2) & 0.38 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 248 (1) & 338 (1) & 92.2 (1) & 115 (1) & 41.6 (1) & -0.04 & 334 (1) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 1 wk ahead}} & Renewal & 298 (1.2) & 403 (1.19) & 87 (0.94) & 107 (0.93) & 105 (2.52) & -0.07 & 413 (1.24) & 0.50 & 0.79\\
\cmidrule{1-11}
\cellcolor{white}{} & Convolution & 357 (1.22) & 573 (1.49) & 104 (0.79) & 204 (1.89) & 48.8 (0.94) & -0.10 & 565 (1.32) & 0.33 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 368 (1.26) & 442 (1.15) & 107 (0.81) & 102 (0.94) & 160 (3.08) & 0.14 & 576 (1.34) & 0.38 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 292 (1) & 385 (1) & 132 (1) & 108 (1) & 51.9 (1) & 0.01 & 429 (1) & 0.62 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 2 wk ahead}} & Renewal & 524 (1.79) & 671 (1.74) & 155 (1.17) & 133 (1.23) & 236 (4.55) & -0.02 & 750 (1.75) & 0.50 & 0.71\\
\bottomrule
\end{tabular}}
\end{table}

\begin{table}[!h]
\caption{\label{tab:score-table-4}Scores for three and four week ahead forecasts (cut to three significant digits and rounded). Note that scores for cases (which include the whole period from October 12th 2020 until March 1st 2021) and deaths (which include only forecasts from the 21st of December 2020 on) are computed on different subsets. Numbers in brackets show the metrics relative to the Hub ensemble (i.e. the median ensemble of all other models submitted to the German and Polish Forecast Hub, excluding our contributions). WIS is the mean weighted interval score (lower values are better), WIS - sd is the standard deviation of all scores achieved by a model. Dispersion, over-prediction and under-prediction together sum up to the weighted interval score. Bias (between -1 and 1, 0 is ideal) represents the general average tendency of a model to over- or underpredict. 50\% and 90\%-coverage are the percentage of observed values that fell within the 50\% and 90\% prediction intervals of a model.\\\hspace{\textwidth}}

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Crowd forecast & 27000 (0.81) & 26200 (0.64) & 4750 (0.52) & 11000 (1.43) & 11200 (0.67) & 0.02 & 39000 (0.83) & 0.14 & 0.48\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 33400 (1) & 40700 (1) & 9130 (1) & 7690 (1) & 16600 (1) & -0.01 & 46900 (1) & 0.29 & 0.62\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 3 wk ahead}} & Renewal & 50600 (1.51) & 70000 (1.72) & 10800 (1.18) & 7710 (1) & 32100 (1.93) & 0.13 & 68700 (1.46) & 0.29 & 0.55\\
\cmidrule{1-11}
\cellcolor{white}{} & Crowd forecast & 39200 (0.7) & 38600 (0.52) & 5970 (0.49) & 15600 (1.26) & 17600 (0.56) & 0.07 & 54800 (0.74) & 0.05 & 0.38\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 55900 (1) & 73700 (1) & 12200 (1) & 12400 (1) & 31300 (1) & 0.01 & 74400 (1) & 0.24 & 0.52\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 4 wk ahead}} & Renewal & 91700 (1.64) & 135000 (1.83) & 19500 (1.6) & 8990 (0.72) & 63200 (2.02) & 0.09 & 125000 (1.68) & 0.31 & 0.48\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Convolution & 541 (1.7) & 802 (2.45) & 157 (0.91) & 279 (3.01) & 105 (1.91) & -0.04 & 747 (1.53) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 414 (1.3) & 526 (1.6) & 137 (0.8) & 82 (0.88) & 194 (3.52) & 0.12 & 648 (1.33) & 0.42 & 0.83\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 319 (1) & 328 (1) & 172 (1) & 92.7 (1) & 55.1 (1) & -0.03 & 488 (1) & 0.54 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 3 wk ahead}} & Renewal & 724 (2.27) & 916 (2.79) & 249 (1.45) & 158 (1.7) & 317 (5.75) & -0.01 & 1040 (2.13) & 0.46 & 0.83\\
\cmidrule{1-11}
\cellcolor{white}{} & Convolution & 763 (1.8) & 932 (2.1) & 268 (1.26) & 331 (2.63) & 164 (1.91) & 0.01 & 985 (1.46) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 498 (1.17) & 633 (1.43) & 168 (0.79) & 83.6 (0.66) & 246 (2.87) & 0.14 & 756 (1.12) & 0.38 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 424 (1) & 443 (1) & 212 (1) & 126 (1) & 85.7 (1) & -0.06 & 675 (1) & 0.58 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 4 wk ahead}} & Renewal & 959 (2.26) & 1210 (2.73) & 337 (1.59) & 200 (1.59) & 421 (4.91) & -0.05 & 1350 (2) & 0.50 & 0.79\\
\bottomrule
\end{tabular}}
\end{table}

\clearpage

\hypertarget{aggregate-performance-by-location}{%
\subsection{Aggregate performance by
location}\label{aggregate-performance-by-location}}

\hypertarget{performance-in-germany}{%
\subsubsection{Performance in Germany}\label{performance-in-germany}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-all-Germany-v4} \caption{Visualisation of aggregate performance metrics for forecasts one to four weeks into the future in Germany. A, B: mean weighted interval score (WIS, lower indicates better performance) across horizons. WIS is decomposed into its components dispersion, over-prediction and under-prediction. C: Empirical coverage of the 50\% prediction intervals (50\% coverage is perfect). D: Empirical coverage of the 90\% prediction intervals. E: Dispersion (same as in panel A, B). Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast (lower is better). H. Standard deviation of all WIS values for different horizons}\label{fig:agg-performance-all-Germany}
\end{figure}

\hypertarget{performance-in-poland}{%
\subsubsection{Performance in Poland}\label{performance-in-poland}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-all-Poland-v4} \caption{Visualisation of aggregate performance metrics for forecasts one to four weeks into the future in Poland. A, B: mean weighted interval score (WIS, lower indicates better performance) across horizons. WIS is decomposed into its components dispersion, over-prediction and under-prediction. C: Empirical coverage of the 50\% prediction intervals (50\% coverage is perfect). D: Empirical coverage of the 90\% prediction intervals. E: Dispersion (same as in panel A, B). Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast (lower is better). H. Standard deviation of all WIS values for different horizons}\label{fig:agg-performance-all-Poland}
\end{figure}

\hypertarget{performance-across-locations-in-absolute-terms}{%
\subsubsection{Performance across locations in absolute
terms}\label{performance-across-locations-in-absolute-terms}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-2-weeks-locations-all-v4} \caption{Visualisation of aggregate performance metrics across locations. A, B: mean weighted interval score (WIS, lower indicates better performance) across horizons. WIS is decomposed into its components dispersion, over-prediction and under-prediction. C: Empirical coverage of the 50\% prediction intervals (50\% coverage is perfect). D: Empirical coverage of the 90\% prediction intervals. E: Dispersion (same as in panel A, B). Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast (lower is better). H. Standard deviation of WIS values.}\label{fig:performance-locations}
\end{figure}

\hypertarget{performance-across-locations-in-relative-terms}{%
\subsection{Performance across locations in relative
terms}\label{performance-across-locations-in-relative-terms}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-2-weeks-locations-all-rel-v4} \caption{Visualisation of relative aggregate performance metrics across locations. A, B: mean weighted interval score (WIS) across locations (lower values indicate better performance). C, D: Empirical coverage of the 50\% and 90\% prediction intervals. E: Dispersion. Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- orunderpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast. H. Standard deviation of WIS values.}\label{fig:performance-locations-rel}
\end{figure}

\clearpage

\hypertarget{visualisation-of-daily-reported-cases-and-deaths}{%
\subsection{Visualisation of daily reported cases and
deaths}\label{visualisation-of-daily-reported-cases-and-deaths}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/daily_truth} \caption{Visualisation of daily report data. The black line represents weekly data divided by seven. Data were last accessed through the German and Polish Forecast Hub on August 21 2021.}\label{fig:daily-truth}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/daily-data-updates} \caption{Visualisation of the absolute difference between the daily report data at the time and the data now. In Germany, there were zero cases and deaths reported on 2020-10-12, and only later 2467 cases and 6 deaths were added. Data were last accessed through the German and Polish Forecast Hub on May 10 2022.}\label{fig:daily-truth-update}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/weekly-data-updates} \caption{Visualisation of the relative difference between the weekly report data at the time and the data now. Apart from the data that was retrospectively added on 2020-10-12, data updates did not have a noticeable effect on weekly data (as shown in the forecasting application). Data were last accessed through the German and Polish Forecast Hub on May 10 2022.}\label{fig:weekly-truth-update}
\end{figure}

\clearpage

\hypertarget{visualisation-of-scores-and-forecasts-1-3-4-weeks-ahead}{%
\subsection{Visualisation of scores and forecasts 1, 3, 4 weeks
ahead}\label{visualisation-of-scores-and-forecasts-1-3-4-weeks-ahead}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/figure-forecasts-1} \caption{A, C: Visualisation of 50\% prediction intervals of one week ahead forecasts against the reported values. Forecasts that were not scored (because there was no complete set of death forecasts available) are greyed out. B, D: Visualisation of corresponding WIS.}\label{fig:forecasts-and-truth-1}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/figure-forecasts-3} \caption{A, C: Visualisation of 50\% prediction intervals of three week ahead forecasts against the reported values. Forecasts that were not scored (because there was no complete set of death forecasts available) are greyed out. B, D: Visualisation of corresponding WIS.}\label{fig:forecasts-and-truth-3}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/figure-forecasts-4} \caption{A, C: Visualisation of 50\% prediction intervals of four week ahead forecasts against the reported values. Forecasts that were not scored (because there was no complete set of death forecasts available) are greyed out. B, D: Visualisation of corresponding WIS.}\label{fig:forecasts-and-truth-4}
\end{figure}

\clearpage

\hypertarget{distribution-of-scores}{%
\subsection{Distribution of scores}\label{distribution-of-scores}}

\hypertarget{absolute-scores}{%
\subsubsection{Absolute scores}\label{absolute-scores}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/distribution_scores_wis-1} \caption{A: Distribution of weighted interval scores for one week ahead forecasts of the different models and forecast targets. B: Distribution of WIS separate by country.}\label{fig:distribution-scores-1}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/distribution_scores_wis-3} \caption{A: Distribution of weighted interval scores for three week ahead forecasts of the different models and forecast targets. B: Distribution of WIS separate by country.}\label{fig:distribution-scores-3}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/distribution_scores_wis-4} \caption{A: Distribution of weighted interval scores for four week ahead forecasts of the different models and forecast targets. B: Distribution of WIS separate by country.}\label{fig:distribution-scores-4}
\end{figure}

\hypertarget{ranks-achieved-by-forecasts}{%
\subsubsection{Ranks achieved by
forecasts}\label{ranks-achieved-by-forecasts}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/distribution_scores_wis-1-ranks} \caption{A: Distribution of the ranks (determined by the weighted interval score) for one week ahead forecasts of the different models and forecast targets. B: Distribution of ranks separate by country.}\label{fig:distribution-scores-ranks-1}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/distribution_scores_wis-2-ranks} \caption{A: Distribution of the ranks (determined by the weighted interval score) for two week ahead forecasts of the different models and forecast targets. B: Distribution of ranks separate by country.}\label{fig:distribution-scores-ranks-2}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/distribution_scores_wis-3-ranks} \caption{A: Distribution of the ranks (determined by the weighted interval score) for three week ahead forecasts of the different models and forecast targets. B: Distribution of ranks separate by country.}\label{fig:distribution-scores-ranks-3}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/distribution_scores_wis-4-ranks} \caption{A: Distribution of the ranks (determined by the weighted interval score) for four week ahead forecasts of the different models and forecast targets. B: Distribution of ranks separate by country.}\label{fig:distribution-scores-ranks-4}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/difference-wis-crowd-ensemble} \caption{Density plot with the difference in WIS between the Crowd forecast and the Hub ensemble (values below zero mean better performance of the Crowd forecasts) for a 2 week ahead forecast horizon.}\label{fig:distribution-scores-differences}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/difference-wis-crowd-renewal} \caption{Density plot with the difference in WIS between the Crowd forecast and the Renewal model (values below zero mean better performance of the Crowd forecasts) for a 2 week ahead forecast horizon.}\label{fig:distribution-scores-differences-renewal}
\end{figure}

\newpage

\hypertarget{comparison-of-ensembles}{%
\subsection{Comparison of ensembles}\label{comparison-of-ensembles}}

\hypertarget{performance-visualisation-mean-ensemble}{%
\subsubsection{Performance visualisation mean
ensemble}\label{performance-visualisation-mean-ensemble}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-rel-ensemble-mean-v4} \caption{Visualisation of aggregate performance metrics across forecast horizons for the different versions of the Hub mean ensemble. “Hub-ensemble”     extit{excludes} all our models, Hub-ensemble-all    extit{includes} all of our models, “Hub-ensemble-realised” is the actual hub-ensemble observed in reality, which includes the renewal model and the crowd forecasts, but ont the convolution model. Values (except for Bias) are computed as differences to the Hub ensemble which excludes our contributions. For Coverage, this is an absolute difference, for other metrics this is a percentage difference. A, B: mean weighted interval score (WIS) across horizons relative to the Hub ensemble (lower values indicate better performance). C, D: Empirical coverage of the 50\% and 90\% prediction intervals minus empirical coverage observed for the Hub ensemble. E: Dispersion relative to the dispersion of the Hub ensemble. Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- orunderpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast relative to the Hub ensemble. H. Standard deviation of all WIS values for different horizons relative to the Hub ensemble.}\label{fig:agg-performance-ensemble-mean}
\end{figure}

\hypertarget{tables-median-ensemble}{%
\subsubsection{Tables median ensemble}\label{tables-median-ensemble}}

\begin{table}[!h]
\caption{\label{tab:score-table-ensemble-2}Scores for one and two week ahead forecasts (cut to three significant digits and rounded) for the different versions of the median ensemble. Note that scores for cases (which include the whole period from October 12th 2020 until March 1st 2021) and deaths (which include only forecasts from the 21st of December 2020 on) are computed on different subsets. Numbers in brackets show the metrics relative to the Hub ensemble (i.e. the median ensemble of all other models submitted to the German and Polish Forecast Hub, excluding our contributions). WIS is the mean weighted interval score (lower values are better), WIS - sd is the standard deviation of all scores achieved by a model. Dispersion, over-prediction and under-prediction together sum up to the weighted interval score. Bias (between -1 and 1, 0 is ideal) represents the general average tendency of a model to over- or underpredict. 50\% and 90\%-coverage are the percentage of observed values that fell within the 50\% and 90\% prediction intervals of a model.\\\hspace{\textwidth}}

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Hub-ensemble & 8770 (1) & 11700 (1) & 3670 (1) & 1230 (1) & 3870 (1) & -0.04 & 12700 (1) & 0.57 & 0.81\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised & 6970 (0.79) & 8260 (0.71) & 3060 (0.83) & 943 (0.77) & 2970 (0.77) & 0.04 & 10800 (0.85) & 0.55 & 0.83\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd & 7820 (0.89) & 9630 (0.82) & 3270 (0.89) & 1210 (0.98) & 3330 (0.86) & -0.02 & 12000 (0.94) & 0.48 & 0.81\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 1 wk ahead}} & Hub-ensemble-with-renewal & 7960 (0.91) & 10300 (0.88) & 3190 (0.87) & 1020 (0.83) & 3760 (0.97) & 0.04 & 12100 (0.95) & 0.57 & 0.83\\
\cmidrule{1-11}
\cellcolor{white}{} & Hub-ensemble & 18300 (1) & 21900 (1) & 6140 (1) & 3800 (1) & 8410 (1) & -0.03 & 26800 (1) & 0.43 & 0.64\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised & 16400 (0.9) & 19600 (0.89) & 5350 (0.87) & 3290 (0.87) & 7730 (0.92) & 0.02 & 24200 (0.9) & 0.43 & 0.69\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd & 16900 (0.92) & 19600 (0.89) & 5230 (0.85) & 4310 (1.13) & 7370 (0.88) & 0.00 & 24600 (0.92) & 0.38 & 0.64\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 2 wk ahead}} & Hub-ensemble-with-renewal & 17500 (0.96) & 21400 (0.98) & 5830 (0.95) & 2880 (0.76) & 8770 (1.04) & 0.00 & 25500 (0.95) & 0.45 & 0.71\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Hub-ensemble & 248 (1) & 338 (1) & 92.2 (1) & 115 (1) & 41.6 (1) & -0.04 & 334 (1) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised & 235 (0.95) & 332 (0.98) & 88.6 (0.96) & 90.4 (0.79) & 55.5 (1.33) & -0.01 & 323 (0.97) & 0.62 & 0.88\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-all & 234 (0.94) & 331 (0.98) & 85.2 (0.92) & 98.1 (0.85) & 50.2 (1.21) & -0.05 & 329 (0.99) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-convolution & 234 (0.94) & 329 (0.97) & 90.7 (0.98) & 118 (1.03) & 25.3 (0.61) & -0.08 & 333 (1) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd & 239 (0.96) & 337 (1) & 85.2 (0.92) & 99.6 (0.87) & 54.2 (1.3) & -0.03 & 322 (0.96) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-6}{*}{\raggedright\arraybackslash 1 wk ahead}} & Hub-ensemble-with-renewal & 246 (0.99) & 342 (1.01) & 91.5 (0.99) & 106 (0.92) & 48.6 (1.17) & -0.06 & 342 (1.02) & 0.67 & 0.92\\
\cmidrule{1-11}
\cellcolor{white}{} & Hub-ensemble & 292 (1) & 385 (1) & 132 (1) & 108 (1) & 51.9 (1) & 0.01 & 429 (1) & 0.62 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised & 296 (1.01) & 398 (1.03) & 125 (0.95) & 91 (0.84) & 80.2 (1.55) & 0.05 & 486 (1.13) & 0.58 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-all & 303 (1.04) & 423 (1.1) & 115 (0.87) & 122 (1.13) & 66.1 (1.27) & 0.00 & 483 (1.13) & 0.62 & 0.88\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-convolution & 270 (0.92) & 385 (1) & 121 (0.92) & 119 (1.1) & 29.9 (0.58) & -0.04 & 403 (0.94) & 0.58 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd & 303 (1.04) & 392 (1.02) & 122 (0.92) & 106 (0.98) & 74.6 (1.44) & 0.03 & 499 (1.16) & 0.58 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-6}{*}{\raggedright\arraybackslash 2 wk ahead}} & Hub-ensemble-with-renewal & 296 (1.01) & 397 (1.03) & 128 (0.97) & 97.1 (0.9) & 71.2 (1.37) & -0.01 & 462 (1.08) & 0.67 & 0.92\\
\bottomrule
\end{tabular}}
\end{table}

\begin{table}[!h]
\caption{\label{tab:score-table-ensemble-4}Scores for three and four week ahead forecasts (cut to three significant digits and rounded) for the different versions of the median ensemble. Note that scores for cases (which include the whole period from October 12th 2020 until March 1st 2021) and deaths (which include only forecasts from the 21st of December 2020 on) are computed on different subsets. Numbers in brackets show the metrics relative to the Hub ensemble (i.e. the median ensemble of all other models submitted to the German and Polish Forecast Hub, excluding our contributions). WIS is the mean weighted interval score (lower values are better), WIS - sd is the standard deviation of all scores achieved by a model. Dispersion, over-prediction and under-prediction together sum up to the weighted interval score. Bias (between -1 and 1, 0 is ideal) represents the general average tendency of a model to over- or underpredict. 50\% and 90\%-coverage are the percentage of observed values that fell within the 50\% and 90\% prediction intervals of a model.\\\hspace{\textwidth}}

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Hub-ensemble & 33400 (1) & 40700 (1) & 9130 (1) & 7690 (1) & 16600 (1) & -0.01 & 46900 (1) & 0.29 & 0.62\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised & 30800 (0.92) & 38600 (0.95) & 7910 (0.87) & 6890 (0.9) & 16000 (0.96) & 0.03 & 44200 (0.94) & 0.29 & 0.62\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd & 30800 (0.92) & 34100 (0.84) & 7500 (0.82) & 8960 (1.17) & 14300 (0.86) & 0.02 & 44100 (0.94) & 0.24 & 0.55\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 3 wk ahead}} & Hub-ensemble-with-renewal & 34000 (1.02) & 43100 (1.06) & 8860 (0.97) & 6300 (0.82) & 18900 (1.14) & 0.02 & 48100 (1.03) & 0.29 & 0.60\\
\cmidrule{1-11}
\cellcolor{white}{} & Hub-ensemble & 55900 (1) & 73700 (1) & 12200 (1) & 12400 (1) & 31300 (1) & 0.01 & 74400 (1) & 0.24 & 0.52\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised & 51200 (0.92) & 69900 (0.95) & 10900 (0.89) & 11100 (0.9) & 29300 (0.94) & 0.04 & 69600 (0.94) & 0.19 & 0.57\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd & 48800 (0.87) & 58600 (0.8) & 9700 (0.8) & 13700 (1.1) & 25400 (0.81) & 0.00 & 65800 (0.88) & 0.19 & 0.48\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 4 wk ahead}} & Hub-ensemble-with-renewal & 59100 (1.06) & 84100 (1.14) & 12600 (1.03) & 10100 (0.81) & 36400 (1.16) & 0.01 & 78900 (1.06) & 0.29 & 0.55\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Hub-ensemble & 319 (1) & 328 (1) & 172 (1) & 92.7 (1) & 55.1 (1) & -0.03 & 488 (1) & 0.54 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised & 332 (1.04) & 388 (1.18) & 158 (0.92) & 78.7 (0.85) & 95 (1.72) & -0.02 & 547 (1.12) & 0.46 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-all & 321 (1.01) & 385 (1.17) & 153 (0.89) & 100 (1.08) & 68.1 (1.24) & -0.01 & 535 (1.1) & 0.54 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-convolution & 298 (0.93) & 337 (1.03) & 155 (0.9) & 106 (1.14) & 37.5 (0.68) & -0.04 & 441 (0.9) & 0.67 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd & 319 (1) & 342 (1.04) & 160 (0.93) & 85.1 (0.92) & 73.6 (1.34) & -0.02 & 547 (1.12) & 0.54 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-6}{*}{\raggedright\arraybackslash 3 wk ahead}} & Hub-ensemble-with-renewal & 332 (1.04) & 363 (1.11) & 168 (0.98) & 86.1 (0.93) & 78.2 (1.42) & -0.02 & 528 (1.08) & 0.58 & 0.96\\
\cmidrule{1-11}
\cellcolor{white}{} & Hub-ensemble & 424 (1) & 443 (1) & 212 (1) & 126 (1) & 85.7 (1) & -0.06 & 675 (1) & 0.58 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised & 445 (1.05) & 532 (1.2) & 193 (0.91) & 107 (0.85) & 144 (1.68) & -0.03 & 700 (1.04) & 0.54 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-all & 399 (0.94) & 438 (0.99) & 195 (0.92) & 105 (0.83) & 97.9 (1.14) & -0.05 & 692 (1.03) & 0.46 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-convolution & 384 (0.91) & 387 (0.87) & 196 (0.92) & 122 (0.97) & 65.9 (0.77) & -0.06 & 602 (0.89) & 0.54 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd & 407 (0.96) & 456 (1.03) & 202 (0.95) & 105 (0.83) & 101 (1.18) & -0.03 & 669 (0.99) & 0.67 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-6}{*}{\raggedright\arraybackslash 4 wk ahead}} & Hub-ensemble-with-renewal & 457 (1.08) & 527 (1.19) & 208 (0.98) & 129 (1.02) & 121 (1.41) & -0.06 & 744 (1.1) & 0.50 & 0.96\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{tables-mean-ensemble}{%
\subsubsection{Tables mean ensemble}\label{tables-mean-ensemble}}

\begin{table}[!h]
\caption{\label{tab:score-table-ensemble-mean-2}Scores for one and two week ahead forecasts (cut to three significant digits and rounded) for the different versions of the mean ensemble. Note that scores for cases (which include the whole period from October 12th 2020 until March 1st 2021) and deaths (which include only forecasts from the 21st of December 2020 on) are computed on different subsets. Numbers in brackets show the metrics relative to the Hub mean ensemble (i.e. the mean ensemble of all other models submitted to the German and Polish Forecast Hub, excluding our contributions). WIS is the mean weighted interval score (lower values are better), WIS - sd is the standard deviation of all scores achieved by a model. Dispersion, over-prediction and under-prediction together sum up to the weighted interval score. Bias (between -1 and 1, 0 is ideal) represents the general average tendency of a model to over- or underpredict. 50\% and 90\%-coverage are the percentage of observed values that fell within the 50\% and 90\% prediction intervals of a model.\\\hspace{\textwidth}}

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Hub-ensemble-mean & 8680 (1) & 10300 (1) & 3700 (1) & 1460 (1) & 3520 (1) & -0.02 & 13400 (1) & 0.50 & 0.86\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised-mean & 7600 (0.88) & 8770 (0.85) & 3360 (0.91) & 1090 (0.75) & 3140 (0.89) & 0.01 & 11900 (0.89) & 0.52 & 0.90\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd-mean & 8050 (0.93) & 9070 (0.88) & 3520 (0.95) & 1410 (0.97) & 3120 (0.89) & -0.02 & 12600 (0.94) & 0.48 & 0.88\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 1 wk ahead}} & Hub-ensemble-with-renewal-mean & 8090 (0.93) & 9780 (0.95) & 3490 (0.94) & 1110 (0.76) & 3490 (0.99) & 0.02 & 12700 (0.95) & 0.57 & 0.88\\
\cmidrule{1-11}
\cellcolor{white}{} & Hub-ensemble-mean & 19000 (1) & 22100 (1) & 5960 (1) & 3690 (1) & 9340 (1) & 0.02 & 28800 (1) & 0.33 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised-mean & 17100 (0.9) & 20600 (0.93) & 5550 (0.93) & 2850 (0.77) & 8660 (0.93) & 0.05 & 26000 (0.9) & 0.38 & 0.76\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd-mean & 17600 (0.93) & 20000 (0.9) & 5540 (0.93) & 3790 (1.03) & 8230 (0.88) & 0.01 & 26800 (0.93) & 0.36 & 0.76\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 2 wk ahead}} & Hub-ensemble-with-renewal-mean & 18300 (0.96) & 22600 (1.02) & 5910 (0.99) & 2640 (0.72) & 9720 (1.04) & 0.06 & 27700 (0.96) & 0.38 & 0.76\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Hub-ensemble-mean & 229 (1) & 292 (1) & 101 (1) & 90.4 (1) & 36.7 (1) & -0.07 & 315 (1) & 0.71 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised-mean & 219 (0.96) & 289 (0.99) & 96.8 (0.96) & 79.8 (0.88) & 42.6 (1.16) & -0.04 & 297 (0.94) & 0.71 & 0.88\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-all-mean & 217 (0.95) & 287 (0.98) & 95.3 (0.94) & 83.1 (0.92) & 38.7 (1.05) & -0.07 & 300 (0.95) & 0.67 & 0.88\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-convolution-mean & 225 (0.98) & 292 (1) & 98.7 (0.98) & 94.2 (1.04) & 32 (0.87) & -0.09 & 314 (1) & 0.71 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd-mean & 222 (0.97) & 289 (0.99) & 98 (0.97) & 84.1 (0.93) & 39.6 (1.08) & -0.04 & 295 (0.94) & 0.71 & 0.88\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-6}{*}{\raggedright\arraybackslash 1 wk ahead}} & Hub-ensemble-with-renewal-mean & 225 (0.98) & 290 (0.99) & 99.7 (0.99) & 84.7 (0.94) & 40.5 (1.1) & -0.05 & 314 (1) & 0.71 & 0.88\\
\cmidrule{1-11}
\cellcolor{white}{} & Hub-ensemble-mean & 256 (1) & 306 (1) & 138 (1) & 64.5 (1) & 53.2 (1) & 0.04 & 374 (1) & 0.67 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised-mean & 270 (1.05) & 338 (1.1) & 136 (0.99) & 65.2 (1.01) & 68.1 (1.28) & 0.06 & 413 (1.1) & 0.75 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-all-mean & 268 (1.05) & 346 (1.13) & 133 (0.96) & 78.7 (1.22) & 57.1 (1.07) & 0.05 & 408 (1.09) & 0.75 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-convolution-mean & 259 (1.01) & 322 (1.05) & 133 (0.96) & 81.7 (1.27) & 44.4 (0.83) & 0.03 & 380 (1.02) & 0.67 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd-mean & 264 (1.03) & 315 (1.03) & 133 (0.96) & 70.1 (1.09) & 60 (1.13) & 0.06 & 404 (1.08) & 0.71 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-6}{*}{\raggedright\arraybackslash 2 wk ahead}} & Hub-ensemble-with-renewal-mean & 264 (1.03) & 332 (1.08) & 141 (1.02) & 60.1 (0.93) & 63.1 (1.19) & 0.06 & 390 (1.04) & 0.79 & 0.92\\
\bottomrule
\end{tabular}}
\end{table}

\begin{table}[!h]
\caption{\label{tab:score-table-ensemble-mean-4}Scores for three and four week ahead forecasts (cut to three significant digits and rounded) for the different versions of the mean ensemble. Note that scores for cases (which include the whole period from October 12th 2020 until March 1st 2021) and deaths (which include only forecasts from the 21st of December 2020 on) are computed on different subsets. Numbers in brackets show the metrics relative to the Hub mean ensemble (i.e. the mean ensemble of all other models submitted to the German and Polish Forecast Hub, excluding our contributions). WIS is the mean weighted interval score (lower values are better), WIS - sd is the standard deviation of all scores achieved by a model. Dispersion, over-prediction and under-prediction together sum up to the weighted interval score. Bias (between -1 and 1, 0 is ideal) represents the general average tendency of a model to over- or underpredict. 50\% and 90\%-coverage are the percentage of observed values that fell within the 50\% and 90\% prediction intervals of a model.\\\hspace{\textwidth}}

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Hub-ensemble-mean & 35600 (1) & 42100 (1) & 9340 (1) & 7050 (1) & 19200 (1) & 0.03 & 51200 (1) & 0.26 & 0.62\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised-mean & 32100 (0.9) & 40500 (0.96) & 8830 (0.95) & 4920 (0.7) & 18300 (0.95) & 0.07 & 47200 (0.92) & 0.29 & 0.64\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd-mean & 32200 (0.9) & 36700 (0.87) & 8430 (0.9) & 7190 (1.02) & 16500 (0.86) & 0.04 & 46900 (0.92) & 0.24 & 0.64\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 3 wk ahead}} & Hub-ensemble-with-renewal-mean & 35200 (0.99) & 46000 (1.09) & 9630 (1.03) & 4600 (0.65) & 20900 (1.09) & 0.08 & 51000 (1) & 0.38 & 0.67\\
\cmidrule{1-11}
\cellcolor{white}{} & Hub-ensemble-mean & 60300 (1) & 79300 (1) & 15700 (1) & 10400 (1) & 34100 (1) & 0.04 & 78600 (1) & 0.29 & 0.57\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised-mean & 55000 (0.91) & 77100 (0.97) & 14600 (0.93) & 6620 (0.64) & 33800 (0.99) & 0.11 & 75200 (0.96) & 0.33 & 0.64\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd-mean & 53400 (0.89) & 66600 (0.84) & 13700 (0.87) & 10600 (1.02) & 29200 (0.86) & 0.06 & 70400 (0.9) & 0.26 & 0.60\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 4 wk ahead}} & Hub-ensemble-with-renewal-mean & 61700 (1.02) & 89800 (1.13) & 16400 (1.04) & 6400 (0.62) & 38900 (1.14) & 0.12 & 82900 (1.05) & 0.31 & 0.64\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Hub-ensemble-mean & 289 (1) & 293 (1) & 178 (1) & 45.9 (1) & 65.7 (1) & 0.01 & 443 (1) & 0.58 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised-mean & 310 (1.07) & 348 (1.19) & 182 (1.02) & 42 (0.92) & 86.5 (1.32) & 0.08 & 502 (1.13) & 0.58 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-all-mean & 315 (1.09) & 339 (1.16) & 178 (1) & 62.2 (1.36) & 74 (1.13) & 0.07 & 507 (1.14) & 0.62 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-convolution-mean & 297 (1.03) & 292 (1) & 174 (0.98) & 67.7 (1.47) & 55 (0.84) & 0.01 & 452 (1.02) & 0.67 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd-mean & 294 (1.02) & 299 (1.02) & 172 (0.97) & 48 (1.05) & 74.2 (1.13) & 0.03 & 476 (1.07) & 0.58 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-6}{*}{\raggedright\arraybackslash 3 wk ahead}} & Hub-ensemble-with-renewal-mean & 310 (1.07) & 349 (1.19) & 189 (1.06) & 39.4 (0.86) & 81.9 (1.25) & 0.05 & 482 (1.09) & 0.62 & 1.00\\
\cmidrule{1-11}
\cellcolor{white}{} & Hub-ensemble-mean & 437 (1) & 568 (1) & 232 (1) & 72 (1) & 134 (1) & 0.00 & 702 (1) & 0.62 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-realised-mean & 445 (1.02) & 598 (1.05) & 237 (1.02) & 56.4 (0.78) & 152 (1.13) & 0.06 & 707 (1.01) & 0.58 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-all-mean & 421 (0.96) & 520 (0.92) & 239 (1.03) & 49.9 (0.69) & 132 (0.99) & 0.05 & 678 (0.97) & 0.58 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-convolution-mean & 398 (0.91) & 465 (0.82) & 235 (1.01) & 55.6 (0.77) & 107 (0.8) & 0.00 & 628 (0.89) & 0.67 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble-with-crowd-mean & 418 (0.96) & 533 (0.94) & 222 (0.96) & 66.8 (0.93) & 129 (0.96) & 0.03 & 662 (0.94) & 0.58 & 1.00\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-6}{*}{\raggedright\arraybackslash 4 wk ahead}} & Hub-ensemble-with-renewal-mean & 467 (1.07) & 636 (1.12) & 248 (1.07) & 61 (0.85) & 158 (1.18) & 0.05 & 755 (1.08) & 0.67 & 1.00\\
\bottomrule
\end{tabular}}
\end{table}

\clearpage

\hypertarget{sensitivity-analysis}{%
\subsection{Sensitivity analysis}\label{sensitivity-analysis}}

In the original analysis, cases and deaths were scored on different
periods, as the convolution model was only added later. This sensitivity
shows performance of all models restricted to the period from December
14 2020 until March 1st 2021 where all models were available.

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-all-late-period-v4} \caption{Visualisation of aggregate performance metrics across forecast horizons only for the period from December 14th 2020 on where all models were available. A, B: mean weighted interval score (WIS, lower indicates better performance) across horizons. WIS is decomposed into its components dispersion, over-prediction and under-prediction. C: Empirical coverage of the 50\% prediction intervals (50\% coverage is perfect). D: Empirical coverage of the 90\% prediction intervals. E: Dispersion (same as in panel A, B). Higher values mean greater dispersion of the forecast and imply ceteris paribus a worse score. F: Bias, i.e. general (relative) tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Absolute error of the median forecast (lower is better). H. Standard deviation of all WIS values for different horizons}\label{fig:agg-performance-all-late}
\end{figure}

\begin{table}[!h]
\caption{\label{tab:score-table-late-2}Scores for one and two week ahead forecasts (cut to three significant digits and rounded) calculated on forecasts made between December 14th 2020 and March 1st 2021. Numbers in brackets show the metrics relative to the Hub ensemble (i.e. the median ensemble of all other models submitted to the German and Polish Forecast Hub, excluding our contributions). WIS is the mean weighted interval score (lower values are better), WIS - sd is the standard deviation of all scores achieved by a model. Dispersion, over-prediction and under-prediction together sum up to the weighted interval score. Bias (between -1 and 1, 0 is ideal) represents the general average tendency of a model to over- or underpredict. 50\% and 90\%-coverage are the percentage of observed values that fell within the 50\% and 90\% prediction intervals of a model.\\\hspace{\textwidth}}

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Crowd forecast & 4980 (0.74) & 5730 (0.64) & 2070 (0.6) & 728 (0.74) & 2190 (0.94) & 0.09 & 7810 (0.82) & 0.54 & 0.88\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 6730 (1) & 8960 (1) & 3430 (1) & 978 (1) & 2330 (1) & -0.09 & 9550 (1) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 1 wk ahead}} & Renewal & 9640 (1.43) & 13300 (1.48) & 1970 (0.57) & 4170 (4.26) & 3500 (1.5) & 0.09 & 12700 (1.33) & 0.46 & 0.71\\
\cmidrule{1-11}
\cellcolor{white}{} & Crowd forecast & 10700 (0.99) & 13800 (1.1) & 2880 (0.58) & 2350 (0.85) & 5430 (1.79) & 0.08 & 15400 (1.07) & 0.46 & 0.62\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 10800 (1) & 12500 (1) & 4940 (1) & 2780 (1) & 3030 (1) & -0.13 & 14400 (1) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 2 wk ahead}} & Renewal & 25000 (2.31) & 34000 (2.72) & 4780 (0.97) & 8710 (3.13) & 11500 (3.8) & 0.05 & 32000 (2.22) & 0.50 & 0.67\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Convolution & 255 (1.03) & 343 (1.01) & 82 (0.89) & 142 (1.23) & 31.1 (0.75) & -0.18 & 399 (1.19) & 0.42 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 265 (1.07) & 317 (0.94) & 78.2 (0.85) & 82 (0.71) & 105 (2.52) & 0.08 & 402 (1.2) & 0.38 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 248 (1) & 338 (1) & 92.2 (1) & 115 (1) & 41.6 (1) & -0.04 & 334 (1) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 1 wk ahead}} & Renewal & 298 (1.2) & 403 (1.19) & 87 (0.94) & 107 (0.93) & 105 (2.52) & -0.07 & 413 (1.24) & 0.50 & 0.79\\
\cmidrule{1-11}
\cellcolor{white}{} & Convolution & 357 (1.22) & 573 (1.49) & 104 (0.79) & 204 (1.89) & 48.8 (0.94) & -0.10 & 565 (1.32) & 0.33 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 368 (1.26) & 442 (1.15) & 107 (0.81) & 102 (0.94) & 160 (3.08) & 0.14 & 576 (1.34) & 0.38 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 292 (1) & 385 (1) & 132 (1) & 108 (1) & 51.9 (1) & 0.01 & 429 (1) & 0.62 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 2 wk ahead}} & Renewal & 524 (1.79) & 671 (1.74) & 155 (1.17) & 133 (1.23) & 236 (4.55) & -0.02 & 750 (1.75) & 0.50 & 0.71\\
\bottomrule
\end{tabular}}
\end{table}

\begin{table}[!h]
\caption{\label{tab:score-table-late-4}Scores for three and four week ahead forecasts (cut to three significant digits and rounded) calculated on forecasts made between December 14th 2020 and March 1st 2021. Numbers in brackets show the metrics relative to the Hub ensemble (i.e. the median ensemble of all other models submitted to the German and Polish Forecast Hub, excluding our contributions). WIS is the mean weighted interval score (lower values are better), WIS - sd is the standard deviation of all scores achieved by a model. Dispersion, over-prediction and under-prediction together sum up to the weighted interval score. Bias (between -1 and 1, 0 is ideal) represents the general average tendency of a model to over- or underpredict. 50\% and 90\%-coverage are the percentage of observed values that fell within the 50\% and 90\% prediction intervals of a model.\\\hspace{\textwidth}}

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Crowd forecast & 17200 (1) & 16000 (0.98) & 3800 (0.63) & 5660 (0.85) & 7770 (1.74) & 0.07 & 26800 (1.1) & 0.08 & 0.58\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 17200 (1) & 16300 (1) & 6030 (1) & 6670 (1) & 4470 (1) & -0.16 & 24400 (1) & 0.33 & 0.67\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 3 wk ahead}} & Renewal & 37700 (2.19) & 55900 (3.43) & 8840 (1.47) & 10700 (1.6) & 18100 (4.05) & -0.03 & 49800 (2.04) & 0.33 & 0.58\\
\cmidrule{1-11}
\cellcolor{white}{} & Crowd forecast & 26100 (0.95) & 21000 (0.84) & 4810 (0.7) & 11300 (0.83) & 10100 (1.43) & 0.04 & 39400 (1.05) & 0.00 & 0.46\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 27600 (1) & 25000 (1) & 6900 (1) & 13600 (1) & 7060 (1) & -0.19 & 37400 (1) & 0.29 & 0.54\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 4 wk ahead}} & Renewal & 48900 (1.77) & 77800 (3.11) & 13800 (2) & 11900 (0.88) & 23200 (3.29) & -0.10 & 65500 (1.75) & 0.38 & 0.58\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Convolution & 541 (1.7) & 802 (2.45) & 157 (0.91) & 279 (3.01) & 105 (1.91) & -0.04 & 747 (1.53) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 414 (1.3) & 526 (1.6) & 137 (0.8) & 82 (0.88) & 194 (3.52) & 0.12 & 648 (1.33) & 0.42 & 0.83\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 319 (1) & 328 (1) & 172 (1) & 92.7 (1) & 55.1 (1) & -0.03 & 488 (1) & 0.54 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 3 wk ahead}} & Renewal & 724 (2.27) & 916 (2.79) & 249 (1.45) & 158 (1.7) & 317 (5.75) & -0.01 & 1040 (2.13) & 0.46 & 0.83\\
\cmidrule{1-11}
\cellcolor{white}{} & Convolution & 763 (1.8) & 932 (2.1) & 268 (1.26) & 331 (2.63) & 164 (1.91) & 0.01 & 985 (1.46) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 498 (1.17) & 633 (1.43) & 168 (0.79) & 83.6 (0.66) & 246 (2.87) & 0.14 & 756 (1.12) & 0.38 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 424 (1) & 443 (1) & 212 (1) & 126 (1) & 85.7 (1) & -0.06 & 675 (1) & 0.58 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 4 wk ahead}} & Renewal & 959 (2.26) & 1210 (2.73) & 337 (1.59) & 200 (1.59) & 421 (4.91) & -0.05 & 1350 (2) & 0.50 & 0.79\\
\bottomrule
\end{tabular}}
\end{table}

\clearpage

\hypertarget{overview-of-models-and-forecasters}{%
\subsection{Overview of models and
forecasters}\label{overview-of-models-and-forecasters}}

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{4.5cm}>{\raggedright\arraybackslash}p{7.3cm}}
\caption{\label{tab:table-ensemble-versions}Overview of the models and ensembles used.}\\
\toprule
Name & Explanation\\
\midrule
\endfirsthead
\caption[]{Overview of the models and ensembles used. \textit{(continued)}}\\
\toprule
Name & Explanation\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{Hub-ensemble-realised} & \cellcolor{gray!6}{Official Forecast Hub median ensemble. Created by the Forecast Hub officially under the name 'KITCOVIDhub-median\_ensemble' and used as the default ensemble. Included are our crowd forecasts as well as the renewal model (with one missed submission on December 28 2020, but not the convolution model which was deemed to similar to the renewal model.}\\
\addlinespace \addlinespace
Hub-ensemble-realised-mean & Official Forecast Hub mean ensemble. Created by the Forecast Hub officially under the name 'KITCOVIDhub-mean\_ensemble'.\\
\addlinespace \addlinespace
\cellcolor{gray!6}{ \vphantom{1}} & \cellcolor{gray!6}{}\\
\addlinespace \addlinespace
Hub-ensemble & Version of the official Hub median ensemble which excludes all our contributions.\\
\addlinespace \addlinespace
\cellcolor{gray!6}{Hub-ensemble-mean} & \cellcolor{gray!6}{Version of the official Hub mean ensemble which excludes all our contributions.}\\
\addlinespace \addlinespace
Hub-ensemble-with-renewal, 
    Hub-ensemble-with-renewal-mean & Versions of the official Hub ensembles which of our contributions includes only the Renewal model.\\
\addlinespace \addlinespace
Hub-ensemble-with-crowd, 
\cellcolor{gray!6}{    Hub-ensemble-with-crowd-mean} & \cellcolor{gray!6}{Versions of the official Hub ensembles which of our contributions includes only the Crowd forecast.}\\
\addlinespace \addlinespace
Hub-ensemble-with-convolution, 
    Hub-ensemble-with-convolution-mean & Versions of the official Hub ensembles which of our contributions includes only the Convolution model (which originally was never included in any official Hub ensemble).\\
\addlinespace \addlinespace
Hub-ensemble-with-all, 
\cellcolor{gray!6}{    Hub-ensemble-with-all-mean} & \cellcolor{gray!6}{Versions of the official Hub ensembles which includes all our contributions. For cases, this is identical to the official Hub ensembles, but for deaths the convolution model was added.}\\
\addlinespace \addlinespace
 & \\
\addlinespace \addlinespace
\cellcolor{gray!6}{Crowd forecast} & \cellcolor{gray!6}{Submitted to the Forecast Hub as 'epiforecasts-EpiExpert'}\\
\addlinespace \addlinespace
Renewal model & Submitted to the Forecast Hub as 'epiforecasts-EpiNow2'\\
\addlinespace \addlinespace
\cellcolor{gray!6}{Convolution model} & \cellcolor{gray!6}{Submitted to the Forecast Hub as 'epiforecasts-EpiNow2\_secondary'}\\*
\end{longtable}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/number-forecasters} \caption{Number of participants who submitted a forecast over time.}\label{fig:num-forecasters}
\end{figure}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/ensemble-members} \caption{Number of member models (including our crowd forecasts and the renewal model) in the official Hub ensemble. Note that the renewal model was not included in the ensemble on December 28th 2020.}\label{fig:num-ensemble-members}
\end{figure}

\clearpage

\hypertarget{comparison-of-crowd-forecasts-and-application-baseline}{%
\subsection{Comparison of crowd forecasts and application
baseline}\label{comparison-of-crowd-forecasts-and-application-baseline}}

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/comparison-forecast-interals} \caption{Crowd forecasts and baseline shown in the application for a two week horizon. Shown are the 90\% and 50\% prediction intervals as well as the median (in order of increasing opacity). For any given point in time, the baseline shown in red is what forecasters saw when they opened the app (the baseline shown was constant across all forecast horizons).}\label{fig:compare-forecasters}
\end{figure}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-mcgowanCollaborativeEffortsForecast2019}{}}%
\CSLLeftMargin{1. }
\CSLRightInline{McGowan CJ, Biggerstaff M, Johansson M, Apfeldorf KM,
Ben-Nun M, Brooks L, et al. Collaborative efforts to forecast seasonal
influenza in the {United States}, 2015--2016. Scientific Reports.
2019;9: 683.
doi:\href{https://doi.org/10.1038/s41598-018-36361-9}{10.1038/s41598-018-36361-9}}

\leavevmode\vadjust pre{\hypertarget{ref-reichCollaborativeMultiyearMultimodel2019}{}}%
\CSLLeftMargin{2. }
\CSLRightInline{Reich NG, Brooks LC, Fox SJ, Kandula S, McGowan CJ,
Moore E, et al. A collaborative multiyear, multimodel assessment of
seasonal influenza forecasting in the {United States}. PNAS. 2019;116:
3146--3154.
doi:\href{https://doi.org/10.1073/pnas.1812594116}{10.1073/pnas.1812594116}}

\leavevmode\vadjust pre{\hypertarget{ref-shamanForecastingSeasonalOutbreaks2012}{}}%
\CSLLeftMargin{3. }
\CSLRightInline{Shaman J, Karspeck A. Forecasting seasonal outbreaks of
influenza. PNAS. 2012;109: 20425--20430.
doi:\href{https://doi.org/10.1073/pnas.1208772109}{10.1073/pnas.1208772109}}

\leavevmode\vadjust pre{\hypertarget{ref-biggerstaffResultsCentersDisease2016}{}}%
\CSLLeftMargin{4. }
\CSLRightInline{Biggerstaff M, Alper D, Dredze M, Fox S, Fung IC-H,
Hickmann KS, et al. Results from the centers for disease control and
prevention's predict the 2013--2014 {Influenza Season Challenge}. BMC
Infectious Diseases. 2016;16: 357.
doi:\href{https://doi.org/10.1186/s12879-016-1669-x}{10.1186/s12879-016-1669-x}}

\leavevmode\vadjust pre{\hypertarget{ref-johanssonOpenChallengeAdvance2019}{}}%
\CSLLeftMargin{5. }
\CSLRightInline{Johansson MA, Apfeldorf KM, Dobson S, Devita J, Buczak
AL, Baugher B, et al. An open challenge to advance probabilistic
forecasting for dengue epidemics. PNAS. 2019;116: 24268--24274.
doi:\href{https://doi.org/10.1073/pnas.1909865116}{10.1073/pnas.1909865116}}

\leavevmode\vadjust pre{\hypertarget{ref-yamanaSuperensembleForecastsDengue2016}{}}%
\CSLLeftMargin{6. }
\CSLRightInline{Yamana TK, Kandula S, Shaman J. Superensemble forecasts
of dengue outbreaks. Journal of The Royal Society Interface. 2016;13:
20160410.
doi:\href{https://doi.org/10.1098/rsif.2016.0410}{10.1098/rsif.2016.0410}}

\leavevmode\vadjust pre{\hypertarget{ref-colon-gonzalezProbabilisticSeasonalDengue2021}{}}%
\CSLLeftMargin{7. }
\CSLRightInline{Colón-González FJ, Bastos LS, Hofmann B, Hopkin A,
Harpham Q, Crocker T, et al. Probabilistic seasonal dengue forecasting
in {Vietnam}: {A} modelling study using superensembles. PLOS Medicine.
2021;18: e1003542.
doi:\href{https://doi.org/10.1371/journal.pmed.1003542}{10.1371/journal.pmed.1003542}}

\leavevmode\vadjust pre{\hypertarget{ref-viboudRAPIDDEbolaForecasting2018}{}}%
\CSLLeftMargin{8. }
\CSLRightInline{Viboud C, Sun K, Gaffey R, Ajelli M, Fumanelli L, Merler
S, et al. The {RAPIDD} ebola forecasting challenge: {Synthesis} and
lessons learnt. Epidemics. 2018;22: 13--21.
doi:\href{https://doi.org/10.1016/j.epidem.2017.08.002}{10.1016/j.epidem.2017.08.002}}

\leavevmode\vadjust pre{\hypertarget{ref-funkAssessingPerformanceRealtime2019}{}}%
\CSLLeftMargin{9. }
\CSLRightInline{Funk S, Camacho A, Kucharski AJ, Lowe R, Eggo RM,
Edmunds WJ. Assessing the performance of real-time epidemic forecasts:
{A} case study of {Ebola} in the {Western Area} region of {Sierra
Leone}, 2014-15. PLOS Computational Biology. 2019;15: e1006785.
doi:\href{https://doi.org/10.1371/journal.pcbi.1006785}{10.1371/journal.pcbi.1006785}}

\leavevmode\vadjust pre{\hypertarget{ref-delvalleSummaryResults201420152018}{}}%
\CSLLeftMargin{10. }
\CSLRightInline{Del Valle SY, McMahon BH, Asher J, Hatchett R, Lega JC,
Brown HE, et al. Summary results of the 2014-2015 {DARPA Chikungunya}
challenge. BMC Infectious Diseases. 2018;18: 245.
doi:\href{https://doi.org/10.1186/s12879-018-3124-7}{10.1186/s12879-018-3124-7}}

\leavevmode\vadjust pre{\hypertarget{ref-farrowHumanJudgmentApproach2017}{}}%
\CSLLeftMargin{11. }
\CSLRightInline{Farrow DC, Brooks LC, Hyun S, Tibshirani RJ, Burke DS,
Rosenfeld R. A human judgment approach to epidemiological forecasting.
PLOS Computational Biology. 2017;13: e1005248.
doi:\href{https://doi.org/10.1371/journal.pcbi.1005248}{10.1371/journal.pcbi.1005248}}

\leavevmode\vadjust pre{\hypertarget{ref-funkShorttermForecastsInform2020}{}}%
\CSLLeftMargin{12. }
\CSLRightInline{Funk S, Abbott S, Atkins BD, Baguelin M, Baillie JK,
Birrell P, et al. Short-term forecasts to inform the response to the
{Covid-19} epidemic in the {UK}. medRxiv. 2020; 2020.11.11.20220962.
doi:\href{https://doi.org/10.1101/2020.11.11.20220962}{10.1101/2020.11.11.20220962}}

\leavevmode\vadjust pre{\hypertarget{ref-cramerCOVID19ForecastHub2020}{}}%
\CSLLeftMargin{13. }
\CSLRightInline{Cramer E, Reich NG, Wang SY, Niemi J, Hannan A, House K,
et al. {COVID-19 Forecast Hub}: 4 {December} 2020 snapshot. {Zenodo};
2020.
doi:\href{https://doi.org/10.5281/zenodo.3963371}{10.5281/zenodo.3963371}}

\leavevmode\vadjust pre{\hypertarget{ref-cramerEvaluationIndividualEnsemble2021}{}}%
\CSLLeftMargin{14. }
\CSLRightInline{Cramer E, Ray EL, Lopez VK, Bracher J, Brennen A,
Rivadeneira AJC, et al. Evaluation of individual and ensemble
probabilistic forecasts of {COVID-19} mortality in the {US}. medRxiv.
2021; 2021.02.03.21250974.
doi:\href{https://doi.org/10.1101/2021.02.03.21250974}{10.1101/2021.02.03.21250974}}

\leavevmode\vadjust pre{\hypertarget{ref-bracherShorttermForecastingCOVID192021}{}}%
\CSLLeftMargin{15. }
\CSLRightInline{Bracher J, Wolffram D, Deuschel J, Görgen K, Ketterer
JL, Ullrich A, et al. Short-term forecasting of {COVID-19} in {Germany}
and {Poland} during the second wave -- a preregistered study. medRxiv.
2021; 2020.12.24.20248826.
doi:\href{https://doi.org/10.1101/2020.12.24.20248826}{10.1101/2020.12.24.20248826}}

\leavevmode\vadjust pre{\hypertarget{ref-bracherNationalSubnationalShortterm2021}{}}%
\CSLLeftMargin{16. }
\CSLRightInline{Bracher J, Wolffram D, Deuschel J, Görgen K, Ketterer
JL, Ullrich A, et al. National and subnational short-term forecasting of
{COVID-19} in {Germany} and {Poland}, early 2021. 2021;
2021.11.05.21265810.
doi:\href{https://doi.org/10.1101/2021.11.05.21265810}{10.1101/2021.11.05.21265810}}

\leavevmode\vadjust pre{\hypertarget{ref-europeancovid-19forecasthubEuropeanCovid19Forecast2021}{}}%
\CSLLeftMargin{17. }
\CSLRightInline{European Covid-19 Forecast Hub. European {Covid-19
Forecast Hub}. 2021 {[}cited 30 May 2021{]}. Available:
\url{https://covid19forecasthub.eu/}}

\leavevmode\vadjust pre{\hypertarget{ref-reichAccuracyRealtimeMultimodel2019}{}}%
\CSLLeftMargin{18. }
\CSLRightInline{Reich NG, McGowan CJ, Yamana TK, Tushar A, Ray EL,
Osthus D, et al. Accuracy of real-time multi-model ensemble forecasts
for seasonal influenza in the {U}.{S}. PLOS Computational Biology.
2019;15: e1007486.
doi:\href{https://doi.org/10.1371/journal.pcbi.1007486}{10.1371/journal.pcbi.1007486}}

\leavevmode\vadjust pre{\hypertarget{ref-tetlockForecastingTournamentsTools2014}{}}%
\CSLLeftMargin{19. }
\CSLRightInline{Tetlock PE, Mellers BA, Rohrbaugh N, Chen E. Forecasting
{Tournaments}: {Tools} for {Increasing Transparency} and {Improving} the
{Quality} of {Debate}. Curr Dir Psychol Sci. 2014;23: 290--295.
doi:\href{https://doi.org/10.1177/0963721414534257}{10.1177/0963721414534257}}

\leavevmode\vadjust pre{\hypertarget{ref-atanasovDistillingWisdomCrowds2016}{}}%
\CSLLeftMargin{20. }
\CSLRightInline{Atanasov P, Rescober P, Stone E, Swift SA,
Servan-Schreiber E, Tetlock P, et al. Distilling the {Wisdom} of
{Crowds}: {Prediction Markets} vs. {Prediction Polls}. Management
Science. 2016;63: 691--706.
doi:\href{https://doi.org/10.1287/mnsc.2015.2374}{10.1287/mnsc.2015.2374}}

\leavevmode\vadjust pre{\hypertarget{ref-hoogeveenLaypeopleCanPredict2020}{}}%
\CSLLeftMargin{21. }
\CSLRightInline{Hoogeveen S, Sarafoglou A, Wagenmakers E-J. Laypeople
{Can Predict Which Social-Science Studies Will Be Replicated
Successfully}. Advances in Methods and Practices in Psychological
Science. 2020;3: 267--285.
doi:\href{https://doi.org/10.1177/2515245920919667}{10.1177/2515245920919667}}

\leavevmode\vadjust pre{\hypertarget{ref-replicationmarketsReplicationMarketsReliable2020}{}}%
\CSLLeftMargin{22. }
\CSLRightInline{ReplicationMarkets. Replication {Markets} -- {Reliable}
research replicates\ldots you can bet on it. 2020 {[}cited 13 Oct
2021{]}. Available: \url{https://www.replicationmarkets.com/}}

\leavevmode\vadjust pre{\hypertarget{ref-servan-schreiberPredictionMarketsDoes2004}{}}%
\CSLLeftMargin{23. }
\CSLRightInline{Servan-Schreiber E, Wolfers J, Pennock DM, Galebach B.
Prediction {Markets}: {Does Money Matter}? Electronic Markets. 2004;14:
243--251.
doi:\href{https://doi.org/10.1080/1019678042000245254}{10.1080/1019678042000245254}}

\leavevmode\vadjust pre{\hypertarget{ref-mcandrewExpertJudgmentModel2020}{}}%
\CSLLeftMargin{24. }
\CSLRightInline{McAndrew TC, Reich NG. An expert judgment model to
predict early stages of the {COVID-19} outbreak in the {United States}.
medRxiv. 2020; 2020.09.21.20196725.
doi:\href{https://doi.org/10.1101/2020.09.21.20196725}{10.1101/2020.09.21.20196725}}

\leavevmode\vadjust pre{\hypertarget{ref-recchiaHowWellDid2021}{}}%
\CSLLeftMargin{25. }
\CSLRightInline{Recchia G, Freeman ALJ, Spiegelhalter D. How well did
experts and laypeople forecast the size of the {COVID-19} pandemic? PLOS
ONE. 2021;16: e0250935.
doi:\href{https://doi.org/10.1371/journal.pone.0250935}{10.1371/journal.pone.0250935}}

\leavevmode\vadjust pre{\hypertarget{ref-metaculusPreliminaryLookMetaculus2020}{}}%
\CSLLeftMargin{26. }
\CSLRightInline{Metaculus. A {Preliminary Look} at {Metaculus} and
{Expert Forecasts}. 22 Jun 2020 {[}cited 30 May 2021{]}. Available:
\url{https://www.metaculus.com/news/2020/06/02/LRT/}}

\leavevmode\vadjust pre{\hypertarget{ref-hypermindHypermindSupercollectiveIntelligence2021}{}}%
\CSLLeftMargin{27. }
\CSLRightInline{Hypermind. Hypermind \textbar{} {Supercollective}
intelligence for decision makers. {Hypermind}; 2021 {[}cited 13 Oct
2021{]}. Available: \url{https://www.hypermind.com/en/}}

\leavevmode\vadjust pre{\hypertarget{ref-csetforetellCSETForetell2021}{}}%
\CSLLeftMargin{28. }
\CSLRightInline{CSET Foretell. {CSET Foretell}. 2021 {[}cited 13 Oct
2021{]}. Available: \url{https://www.cset-foretell.com/}}

\leavevmode\vadjust pre{\hypertarget{ref-predictitPredictIt2021}{}}%
\CSLLeftMargin{29. }
\CSLRightInline{PredictIt. {PredictIt}. 2021 {[}cited 13 Oct 2021{]}.
Available: \url{https://www.predictit.org/}}

\leavevmode\vadjust pre{\hypertarget{ref-heldProbabilisticForecastingInfectious2017}{}}%
\CSLLeftMargin{30. }
\CSLRightInline{Held L, Meyer S, Bracher J. Probabilistic forecasting in
infectious disease epidemiology: The 13th {Armitage} lecture. Statistics
in Medicine. 2017;36: 3443--3460.
doi:\href{https://doi.org/10.1002/sim.7363}{10.1002/sim.7363}}

\leavevmode\vadjust pre{\hypertarget{ref-crowdforecastr}{}}%
\CSLLeftMargin{31. }
\CSLRightInline{Bosse NI, Abbott S, EpiForecasts, Funk S.
Crowdforecastr: Eliciting crowd forecasts in r shiny. 2020.
doi:\href{https://doi.org/10.5281/zenodo.4618519}{10.5281/zenodo.4618519}}

\leavevmode\vadjust pre{\hypertarget{ref-covidgermanforecasts}{}}%
\CSLLeftMargin{32. }
\CSLRightInline{Bosse NI, Abbott S, EpiForecasts, Funk S.
Covid.german.forecasts: Forecasting covid-19 related metrics for the
german/poland forecast hub. 2020. }

\leavevmode\vadjust pre{\hypertarget{ref-ecdcDownloadHistoricalData2020}{}}%
\CSLLeftMargin{33. }
\CSLRightInline{ECDC. Download historical data (to 14 {December} 2020)
on the daily number of new reported {COVID-19} cases and deaths
worldwide. {European Centre for Disease Prevention and Control}; 14 Dec
2020 {[}cited 30 May 2021{]}. Available:
\url{https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide}}

\leavevmode\vadjust pre{\hypertarget{ref-rkiRKICoronavirusSARSCoV22021}{}}%
\CSLLeftMargin{34. }
\CSLRightInline{RKI. {RKI} - {Coronavirus SARS-CoV-2} - {Aktueller
Lage-}/{Situationsbericht} des {RKI} zu {COVID-19}. 2021 {[}cited 30 May
2021{]}. Available:
\url{https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Situationsberichte/Gesamt.html}}

\leavevmode\vadjust pre{\hypertarget{ref-forsal.plRozbieznosciStatystykachKoronawirusa2020}{}}%
\CSLLeftMargin{35. }
\CSLRightInline{Forsal.pl. Rozbieżności w statystykach koronawirusa. 22
tys. przypadków będą doliczone do ogólnej liczby wyników. 2020 {[}cited
30 May 2021{]}. Available:
\url{https://forsal.pl/lifestyle/zdrowie/artykuly/8017628,rozbieznosci-w-statystykach-koronawirusa-22-tys-przypadkow-beda-doliczone-do-ogolnej-liczby-wynikow.html}}

\leavevmode\vadjust pre{\hypertarget{ref-aerzteblattSARSCoV2DiagnostikRKIPasst2020}{}}%
\CSLLeftMargin{36. }
\CSLRightInline{Ärzteblatt DÄG Redaktion Deutsches.
SARS-CoV-2-Diagnostik: RKI passt Testempfehlungen an. {Deutsches
Ärzteblatt}; 3 Nov 2020 {[}cited 30 May 2021{]}. Available:
\url{https://www.aerzteblatt.de/nachrichten/118001/SARS-CoV-2-Diagnostik-RKI-passt-Testempfehlungen-an}}

\leavevmode\vadjust pre{\hypertarget{ref-golem}{}}%
\CSLLeftMargin{37. }
\CSLRightInline{Fay C, Guyader V, Rochette S, Girard C. Golem: A
framework for robust shiny applications. 2021. Available:
\url{https://github.com/ThinkR-open/golem}}

\leavevmode\vadjust pre{\hypertarget{ref-shiny}{}}%
\CSLLeftMargin{38. }
\CSLRightInline{Chang W, Cheng J, Allaire J, Sievert C, Schloerke B, Xie
Y, et al. Shiny: Web application framework for r. 2021. Available:
\url{https://CRAN.R-project.org/package=shiny}}

\leavevmode\vadjust pre{\hypertarget{ref-ourworldindataCOVID19DataExplorer2020}{}}%
\CSLLeftMargin{39. }
\CSLRightInline{Our World in Data. {COVID-19 Data Explorer}. {Our World
in Data}; 2020 {[}cited 30 May 2021{]}. Available:
\url{https://ourworldindata.org/coronavirus-data-explorer}}

\leavevmode\vadjust pre{\hypertarget{ref-epinow2}{}}%
\CSLLeftMargin{40. }
\CSLRightInline{Abbott S, Hellewell J, Hickson J, Munday J, Gostic K,
Ellis P, et al. EpiNow2: Estimate real-time case counts and time-varying
epidemiological parameters. -. 2020;-: --.
doi:\href{https://doi.org/10.5281/zenodo.3957489}{10.5281/zenodo.3957489}}

\leavevmode\vadjust pre{\hypertarget{ref-fraserEstimatingIndividualHousehold2007}{}}%
\CSLLeftMargin{41. }
\CSLRightInline{Fraser C. Estimating {Individual} and {Household
Reproduction Numbers} in an {Emerging Epidemic}. PLOS ONE. 2007;2: e758.
doi:\href{https://doi.org/10.1371/journal.pone.0000758}{10.1371/journal.pone.0000758}}

\leavevmode\vadjust pre{\hypertarget{ref-epiforecasts.ioux2fcovidCovid19TemporalVariation2020}{}}%
\CSLLeftMargin{42. }
\CSLRightInline{epiforecasts.io/covid. Covid-19: {Temporal} variation in
transmission during the {COVID-19} outbreak. {Covid-19}; 2020 {[}cited
30 May 2021{]}. Available: \url{https://epiforecasts.io/covid/}}

\leavevmode\vadjust pre{\hypertarget{ref-sherrattExploringSurveillanceData2021}{}}%
\CSLLeftMargin{43. }
\CSLRightInline{Sherratt K, Abbott S, Meakin SR, Hellewell J, Munday JD,
Bosse N, et al. Exploring surveillance data biases when estimating the
reproduction number: With insights into subpopulation transmission of
{Covid-19} in {England}. 2021; 2020.10.18.20214585.
doi:\href{https://doi.org/10.1101/2020.10.18.20214585}{10.1101/2020.10.18.20214585}}

\leavevmode\vadjust pre{\hypertarget{ref-abbottEstimatingTimevaryingReproduction2020a}{}}%
\CSLLeftMargin{44. }
\CSLRightInline{Abbott S, Hellewell J, Thompson RN, Sherratt K, Gibbs
HP, Bosse NI, et al. Estimating the time-varying reproduction number of
{SARS-CoV-2} using national and subnational case counts. 2020 {[}cited
14 Oct 2021{]}.
doi:\href{https://doi.org/10.12688/wellcomeopenres.16006.1}{10.12688/wellcomeopenres.16006.1}}

\leavevmode\vadjust pre{\hypertarget{ref-kraemer2020epidemiological}{}}%
\CSLLeftMargin{45. }
\CSLRightInline{Xu B, Gutierrez B, Hill S, Scarpino S, Loskill A, Wu J,
et al. Epidemiological data from the nCoV-2019 outbreak: Early
descriptions from publicly available data. 2020. Available:
\url{http://virological.org/t/epidemiological-data-from-the-ncov-2019-outbreak-early-descriptions-from-publicly-available-data/337}}

\leavevmode\vadjust pre{\hypertarget{ref-rstan}{}}%
\CSLLeftMargin{46. }
\CSLRightInline{Stan Development Team. RStan: The r interface to stan.
2020. Available: \url{http://mc-stan.org/}}

\leavevmode\vadjust pre{\hypertarget{ref-bracherEvaluatingEpidemicForecasts2021}{}}%
\CSLLeftMargin{47. }
\CSLRightInline{Bracher J, Ray EL, Gneiting T, Reich NG. Evaluating
epidemic forecasts in an interval format. PLoS Comput Biol. 2021;17:
e1008618.
doi:\href{https://doi.org/10.1371/journal.pcbi.1008618}{10.1371/journal.pcbi.1008618}}

\leavevmode\vadjust pre{\hypertarget{ref-gneiting_strictly_2007}{}}%
\CSLLeftMargin{48. }
\CSLRightInline{Gneiting T, Raftery AE. Strictly proper scoring rules,
prediction, and estimation. Journal of the American Statistical
Association. 2007;102: 359--378.
doi:\href{https://doi.org/10.1198/016214506000001437}{10.1198/016214506000001437}}

\leavevmode\vadjust pre{\hypertarget{ref-scoringutils}{}}%
\CSLLeftMargin{49. }
\CSLRightInline{Bosse NI, Abbott S, EpiForecasts, Funk S. Scoringutils:
Utilities for scoring and assessing predictions. 2020.
doi:\href{https://doi.org/10.5281/zenodo.4618017}{10.5281/zenodo.4618017}}

\leavevmode\vadjust pre{\hypertarget{ref-deutschewelleCoronavirusGermanyImpose2020}{}}%
\CSLLeftMargin{50. }
\CSLRightInline{Deutsche Welle. Coronavirus: {Germany} to impose
one-month partial lockdown \textbar{} {DW} \textbar{} 28.10.2020. 2020
{[}cited 29 Jun 2021{]}. Available:
\url{https://www.dw.com/en/coronavirus-germany-to-impose-one-month-partial-lockdown/a-55421241}}

\leavevmode\vadjust pre{\hypertarget{ref-generationinterval}{}}%
\CSLLeftMargin{51. }
\CSLRightInline{Ganyani T, Kremer C, Chen D, Torneri A, Faes C, Wallinga
J, et al. Estimating the generation interval for coronavirus disease
(COVID-19) based on symptom onset data, march 2020. Eurosurveillance.
2020;25. }

\leavevmode\vadjust pre{\hypertarget{ref-incubationperiod}{}}%
\CSLLeftMargin{52. }
\CSLRightInline{Lauer SA, Grantz KH, Bi Q, Jones FK, Zheng Q, Meredith
HR, et al. The incubation period of coronavirus disease 2019 (COVID-19)
from publicly reported confirmed cases: Estimation and application.
Annals of Internal Medicine. 2020;172: 577--582. }

\leavevmode\vadjust pre{\hypertarget{ref-covidregionaldata}{}}%
\CSLLeftMargin{53. }
\CSLRightInline{Abbott S, Sherratt K, Bevan J, Gibbs H, Hellewell J,
Munday J, et al. Covidregionaldata: Subnational data for the covid-19
outbreak. -. 2020;-: --.
doi:\href{https://doi.org/10.5281/zenodo.3957539}{10.5281/zenodo.3957539}}

\leavevmode\vadjust pre{\hypertarget{ref-EvaluatingUseReproduction}{}}%
\CSLLeftMargin{54. }
\CSLRightInline{Evaluating the use of the reproduction number as an
epidemiological tool, using spatio-temporal trends of the {Covid-19}
outbreak in {England} \textbar{} {medRxiv}. {[}cited 30 May 2021{]}.
Available:
\url{https://www.medrxiv.org/content/10.1101/2020.10.18.20214585v1}}

\leavevmode\vadjust pre{\hypertarget{ref-approxGP}{}}%
\CSLLeftMargin{55. }
\CSLRightInline{Riutort-Mayol G, Bürkner P-C, Andersen MR, Solin A,
Vehtari A. Practical hilbert space approximate bayesian gaussian
processes for probabilistic programming. 2020. Available:
\url{https://arxiv.org/abs/2004.11408}}

\leavevmode\vadjust pre{\hypertarget{ref-epidemia}{}}%
\CSLLeftMargin{56. }
\CSLRightInline{Scott JA, Gandy A, Mishra S, Unwin J, Flaxman S, Bhatt
S. Epidemia: Modeling of epidemics using hierarchical bayesian models.
2020. Available:
\url{https://imperialcollegelondon.github.io/epidemia/}}

\leavevmode\vadjust pre{\hypertarget{ref-bhattSemiMechanisticBayesianModeling}{}}%
\CSLLeftMargin{57. }
\CSLRightInline{Bhatt S, Ferguson N, Flaxman S, Gandy A, Mishra S, Scott
JA. Semi-{Mechanistic Bayesian} modeling of {COVID-19} with {Renewal
Processes}. : 14. }

\end{CSLReferences}

\nolinenumbers

\end{document}