### 2023-12-11 analysis/analysis.Rmd

---
title: "Analysis paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
options(knitr.kable.NA = '')
library(here)
library(magrittr)
library(scoringutils) #needs v0.18
# remotes::install_github("epiforecasts/scoringutils@v0.1.8")
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)
library(covid.german.forecasts)
library(grates)
library(ggdist)
library(ggthemes)
library(scales)
library(dplyr)
library(tidyr)
library(forcats)

# define colors for deaths and cases
colcases <- brewer.pal(name = "Accent", n = 8)[-c(1, 4)]
coldeaths <- brewer.pal(name = "Accent", n = 8)[-4]
colensemble <- c("#FDC086", brewer.pal(name = "Set2", n = 8))

# helper function for labeling plots
scale_fn <- function(x) {
  ifelse(abs(x) >= 5000,
         paste0(x / 1000, "k"), x)
}
```

# Visualisation of aggregate performance across horizons

``` {r}
# helper function to score forecasts
score_helper <- function(data, summarise_by) {
  scores <- eval_forecasts(data, 
                           summarise_by = summarise_by, 
                           quantiles = 0.5, 
                           sd = TRUE)

  # for cases, filter out redundant ensembles as convolution model only does 
  # deaths (e.g. ensemble with all is identical to realised ensemble for cases)
  scores <- scores %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case")
    )

  setnames(scores, old = c("model", "sharpness"), 
           new = c("Model", "dispersion"))
}

# helper function to create the WIS plot with its sub-components in panels A and B
component_plot <- function(scores, type, colour_scheme, x = x, rel = relative,
                           facet_formula, vals = NULL) {

  # define colour values for plot
  # for cases and ensembles, filter out identical ensembles
  if (colour_scheme == "ensemble" & type == "case") {
    scores <- scores %>%
      filter(!(Model %in% c("Hub-ensemble-with-all", 
                            "Hub-ensemble-with-convolution")))
    vals <- vals[-c(3, 4)]
  } else if (type == "case") {
    vals <- vals[-1]
  }

  # pivot scores and rename WIS components
  scores <- scores %>%
    select(Model, target_type, dispersion, interval_score,
           underprediction, overprediction, !!x) %>%
    rename("Dispersion" = dispersion, 
           "Over-prediction" = overprediction, 
           "Under-prediction" = underprediction) %>%
    filter(target_type == type) %>%
    pivot_longer(names_to = "WIS components",
                 cols = c(`Dispersion`, `Over-prediction`, `Under-prediction`)) %>%
    mutate(`WIS components` = fct_relevel(`WIS components`, 
                                          c("Dispersion", "Under-prediction", 
                                            "Over-prediction")))

  # compute relative version of the scores
  if (rel) {
    scores <- scores %>%
      arrange(!!sym(x), target_type, interval_score) %>%
      group_by(!!sym(x), target_type) %>%
      mutate_at("value", ~ . / interval_score[Model %in% c("Hub-ensemble", 
                                                           "Hub-ensemble-mean")])
  }

  # create plot
  p <- scores %>%
    ggplot(aes(x = reorder(Model, (interval_score)), y = value, fill = Model, 
               group = Model, alpha = `WIS components`))

  if (rel) {
    p <- p + 
      geom_hline(yintercept = 1, linetype = "dashed") +
      scale_y_continuous(breaks = seq(0, max(scores$value), 0.25))
  }

  p <- p + 
    scale_alpha_manual(values = c(1, 0.1, 0.6)) +
    scale_fill_manual(values = vals) + 
    geom_bar(stat = "identity", position = "stack",
             color = "black",
             size = 0.1,
             width = 0.8) +
    facet_wrap(as.formula(paste("~", x)), ncol = 4,
               strip.position="bottom") + 
    theme_minimal() + 
    theme(axis.text.x = element_blank(), 
          plot.title = element_text(hjust = 0.5, 
                                    size = 8), 
          legend.position = c(0.12, 0.83)) + 
    labs(x = NULL, title = type, y = "WIS")

  return(p)
}

# helper function to make subplots E - H
plot_helper <- function(tmp = scores, xvar = x, y, rel = relative, 
                        facet_formula, colour, scalefree = NULL) {
  if (is.null(scalefree)) {
    scalefree <- "free_y"
  }
  if (rel) {
    scalefree <- "fixed"
    tmp <- tmp %>%
      arrange(target_type, !!sym(xvar)) %>%
      group_by(target_type, !!sym(xvar)) %>%
      mutate_at(y, ~ . / .[Model %in% c("Hub-ensemble", 
                                        "Hub-ensemble-mean")])
  }

  out <- tmp %>%
    ggplot(aes_string(x = xvar, y = y, group = "Model", color = "Model")) +
    geom_point(position = position_dodge(width = 0.1)) +
    geom_line(position = position_dodge(width = 0.1)) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_color_manual(values = colour) + 
    scale_y_continuous(labels = scale_fn) + 
    facet_wrap(facet_formula, scales = scalefree) +
    guides(fill = "none", colour = "none")

  if (!rel) {
    out <- out + 
      expand_limits(y = 0)
  }

  return(out)
}

# function to create Figure 1
aggregate_performance <- function(data, 
                                  x = "horizon",
                                  xlabel = "Forecast horizon (weeks)",
                                  output_dir = here("analysis", "plots"), 
                                  plotname,
                                  relative = FALSE,
                                  facet_formula = ~ target_type, 
                                  colour_scheme = c("normal", "ensemble"),
                                  vals = NULL) {

  # calculate scores
  scores <- score_helper(data, c("model", "target_type", x))

  # define colour values for the plot
  if (colour_scheme[1] == "normal") {
    vals <- coldeaths
  } else {
    vals <- colensemble
  }

  # plot with wis components (panels A and B) ----------------------------------
  component_case <- component_plot(scores, "case", x = x, vals = vals, 
                                   facet_formula = facet_formula, 
                                   rel = relative, 
                                   colour_scheme = colour_scheme[1]) + 
    guides(fill = "none")
  component_death <- component_plot(scores, "death", x = x, vals = vals, 
                                    facet_formula = facet_formula, 
                                    rel = relative, 
                                    colour_scheme = colour_scheme[1])

  # plot with dispersion of forecasts in panel E -------------------------------
  p_sharpness <- 
    plot_helper(scores, y = "dispersion", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Dispersion", x = NULL)

  # bias plot in panel F -------------------------------------------------------
  # define appropriate y limits
  bias_limits <- max(abs(scores$bias)) * c(-1.1, 1.1) %>% 
    round(2)
  p_bias <- 
    plot_helper(scores, y = "bias", rel = FALSE, colour = vals, xvar = x,
                facet_formula = facet_formula, scalefree = "fixed") + 
    labs(y = "Bias", x = NULL) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    expand_limits(y=bias_limits)

  # plot with absolute error of performance in panel G -------------------------
  p_aem <-
    plot_helper(scores, y = "aem", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Absolute error", x = xlabel)

  # plot with sd of performance in panel H -------------------------------------
  p_wis_sd <- 
    plot_helper(scores, y = "interval_score_sd", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "WIS - sd", x = xlabel)

  # coverage plots (panel C and D) ---------------------------------------------
  # calculate scores that also include the interval range
  scores <- score_helper(data,  c("model", "target_type", x, "range"))

  # take difference to Hub ensemble if relative
  if (relative) {
    scores <- scores %>%
      arrange(target_type, range, !!sym(x)) %>%
      group_by(target_type, range, !!sym(x)) %>%
      mutate_at("coverage", ~ . - .[Model %in%  c("Hub-ensemble", 
                                                  "Hub-ensemble-mean")])
  }

  p_cov50 <- scores %>%
    filter(range == c(50)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                colour = vals, xvar = x, scalefree = "fixed") +
    labs(y = "Diff. cov. - 50% PI", x = "") + 
    scale_y_continuous(labels = scales::percent)

  # if not relative, add a dashed line at 50%
  if (!relative) {
    p_cov50 <- p_cov50 + 
      expand_limits(y=c(0, 1)) + 
      labs(y = "Coverage - 50% PI", x = "") + 
      geom_hline(yintercept = 0.5, linetype = "dashed") + 
      scale_y_continuous(labels = scales::percent)
    }

  p_cov90 <-  scores %>%
    filter(range == c(90)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                xvar = x, colour = vals, scalefree = "fixed") + 
    labs(y = "Diff. cov. - 90% PI", x = "") + 
    scale_y_continuous(labels = scales::percent)

  # if not relative, add a dashed line at 90%
  if (!relative) {
    p_cov90 <- p_cov90 + 
      expand_limits(y=c(0, 1)) + 
      labs(y = "Coverage - 90% PI", x = "") + 
      geom_hline(yintercept = 0.9, linetype = "dashed") + 
      scale_y_continuous() + 
      scale_y_continuous(labels = scales::percent)
  }

  # put all plots together -----------------------------------------------------
  p_combined4 <- 
    (component_case + component_death) / 
    (p_cov50 + p_cov90) /
    (p_sharpness + p_bias) /
    (p_aem + p_wis_sd) +
    plot_layout(guides = "collect", 
                heights = c(1.3, 1, 1, 1)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin()) &
    labs(x = xlabel)

    ggsave(here(output_dir, paste0(plotname, "-v4.tiff")), 
           height = 9.1, width = 10, device = "tiff", dpi = 300)
}

# create plots with aggregate performance --------------------------------------
# regular plot for Figure 1
filtered_data[model %in% regular_models] %>%
  aggregate_performance(plotname = "aggregate-performance-all")

# plot with restricted time period for sensitivity analysis for SI
filtered_data[model %in% regular_models] %>%
  filter(!(forecast_date %in% as.character(forecast_dates$death_not_scored))) %>%
  aggregate_performance(plotname = "aggregate-performance-all-late-period")

# plot version that compares locations for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all", 
                        x = "location_name", 
                        xlabel = "Location")

# plot version that compares locations in relative terms for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all-rel", 
                        x = "location_name", 
                        xlabel = "Location", relative = TRUE)

# plot for different median ensembles (in relative terms) Figure 4
filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  # filter(!(model == "Hub-ensemble-with-all" & target_type == "case")) %>%
  aggregate_performance(plotname = "aggregate-performance-rel-ensemble", 
                        relative = TRUE, colour_scheme = "ensemble")

# plot for different mean ensembles (in relative terms) for SI
filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  # filter(!(model == "Hub-ensemble-with-all" & target_type == "case")) %>%
  aggregate_performance(plotname = "aggregate-performance-rel-ensemble-mean", 
                        relative = TRUE, colour_scheme = "ensemble")

# make SI version of Figure 1 for Germany and Poland separately
filtered_data[(model %in% regular_models) & location == "GM" ] %>%
  aggregate_performance(plotname = "aggregate-performance-all-Germany")

filtered_data[(model %in% regular_models) & location == "PL" ] %>%
  aggregate_performance(plotname = "aggregate-performance-all-Poland")

```

# Plot with Visualisation of x week ahead forecasts and scores

```{r truth-vs-forecasts, fig.height=12, fig.width=15}

# helper function which creates panels A and C
truth_and_pred_plot <- function(df, ttype = "case") {

  # adjust distance between bars depending on case or death forecast
  if (ttype == "case") {
    dodgewidth = 3.6
  } else {
    dodgewidth = 3.7
  }

  # rename model column
  setnames(df, old = c("model"), new = c("Model"))

  # define date limits for the plot
  limits <- min(forecast_dates$full_hub_period) + 5 + 0:21 * 7

  # create plot
  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = Model), 
                   size = 1.1,
                   alpha = 1,
                   position = position_dodge(width = dodgewidth)) + 
    geom_line(data = unfiltered_data[target_type == ttype & 
                                       target_end_date %in% limits]) + 
    theme_minimal() + 
    theme(panel.grid.major.x = element_blank()) +
    facet_wrap(~ location_target, scales = "free", ncol = 1) + 
    theme(legend.position = "bottom") + 
    scale_x_date(limits = c(min(limits) - 2, max(limits) + 2), 
                 minor_breaks = limits) + 
      scale_y_continuous(labels = scale_fn)
  return(p)
}

# helper function which creates panels B and D
scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {

  # define plot limits on x axis
  limits <- min(forecast_dates$full_hub_period) + 5 + 0:20 * 7

  p <- scores %>%
    ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
    geom_line(aes(colour = model)) + 
    geom_point(aes(colour = model)) + 
    theme_minimal() + 
    facet_wrap( ~ location_target, ncol = 1, scales = "free") +
    theme(legend.position = "bottom") +
    theme(panel.grid.major.x = element_blank()) +
    scale_y_continuous(labels = scale_fn) + # , 
                       # trans = log_trans()) + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) 
  return(p)
}

# helper function that creates the final plot
forecasts_and_scores <- function(data, 
                                 output_dir = here("analysis", "plots"), 
                                 plotname = "figure-forecasts") {

  # calculate scores
  summarise_by <- c("model", "horizon", "location_name", "target_type", 
                    "forecast_date", "target_end_date", "location_target")

  scores_target_forecastdate <- eval_forecasts(
    data = data,
    summarise_by = summarise_by
  )[, forecast_date := as.Date(forecast_date)]

  # create data.frame for plotting predictions 
  df <- data[quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")

  # create plots for all horizons 
  for (i in 1:4) {

    # truth plot for cases 
    truth_vs_forecast_cases <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "case") + 
      scale_color_manual(values = colcases, name = "Model") + 
      labs(y = "Forecasts and observed values", x = "")  + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm")) + 
      scale_y_continuous(labels = scale_fn)

    # truth plot for deaths
    truth_vs_forecast_deaths <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "death")  + 
      scale_color_manual(values = coldeaths, name = "Model") + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 5 + (i-1) * 7,
               xmax = max(forecast_dates$death_not_scored) + 5 + (i-1) * 7,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Forecasts and observed values", x = "") +
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm"))

    # scores plot for cases
    scores_over_time_cases <- 
      scores_target_forecastdate[horizon == i & target_type == "case"] %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = colcases, name = "Model") + 
      labs(y = "Weighted interval score", x = "") + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none")

    # scores plot for deaths
    scores_over_time_deaths <- 
      scores_target_forecastdate[horizon == i & target_type == "death"] %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = coldeaths, name = "Model") + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Weighted interval score", x = "")

    # putting them all together
    p <- (truth_vs_forecast_cases + scores_over_time_cases) / 
      (truth_vs_forecast_deaths + scores_over_time_deaths) + 
      plot_layout(guides = "collect") &
      plot_layout(widths = c(1.9, 1)) &
      plot_annotation(tag_levels = 'A') &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin = margin())

    ggsave(here(output_dir, 
                paste0(plotname, "-", i, ".tiff")), width = 10, height = 11, 
           device = "tiff", dpi = 300)

  }
}

# create plots -----------------------------------------------------------------
# regular forecasts
unfiltered_data[model %in% c(regular_models)] %>%
  forecasts_and_scores()
```

# Create tables with scores

```{r}

# helper function to create the tables
score_table <- function(scores, 
                        group = TRUE, 
                        full_width = TRUE) {

  # change column names, round column values and reduce to sign. digits
  scores <- as.data.table(scores)
  setnames(scores, 
           old = c("horizon", "model", "target_type", "interval_score", "aem", 
                   "relative_skill", "scaled_rel_skill", "sharpness", 
                   "underprediction", "overprediction", "bias", 
                   "50", "90", "location_target", "target_phase"), 
           new = c("Horizon", "Model", "Target", "WIS", "Abs. error", 
                   "rel. skill", "WIS - rel.", "Dispersion", 
                   "Underpred.", "Overpred.", "Bias", 
                   "50%-Cov.", "90%-Cov.", "Target", "Target"), 
           skip_absent = TRUE)

  # remove some cols, change column order, sort table
  if (all(c("coverage_deviation", "rel. skill") %in% names(scores))) {
    scores <- scores[, .SD,  .SDcols = !c("coverage_deviation", "rel. skill")]
  }

  if (all(c("WIS", "Horizon") %in% names(scores))) {
    wis <- names(scores)[grepl("WIS", names(scores))]
    setcolorder(scores, 
                neworder = c("Target", "Horizon", "Model", wis))  
  }

  if ("Horizon" %in% names(scores)) {
    scores <- scores[order(Target, Horizon)]
    scores[, Horizon := paste(Horizon, "wk ahead")]
  }

  # calculate where to group rows
  case_rows <- nrow(scores[Target == "case"])

  table <- scores[,!c("Target")] %>%
    kable(format = "latex", booktabs = TRUE, 
          align = c("l", "l", rep("c", 9)),
          col.names = c(" ", names(scores)[-(1:2)])) %>%
    collapse_rows(columns = 1) %>%
    kable_styling(latex_options = c("scale_down", 
                                    "hold_position")) %>%
    column_spec(1, background = "white")

  if (group) {
    table <- table %>%
      pack_rows("Cases", 1, case_rows, indent = FALSE, 
                hline_after = TRUE) %>% 
      pack_rows("Deaths", case_rows + 1, nrow(scores), indent = FALSE, 
                hline_after = TRUE)
  }

  return(table)
}

# function that actually makes the tables
make_table <- function(data, 
                       plotname = "table_scores",
                       output_dir = here("analysis", "plots")) {

  # get coverage values
  coverage <- eval_forecasts(
    data = data, 
    summarise_by = c("horizon", "model", "target_type", "range"), 
  )[range %in% c(50, 90), 
    .(model, target_type, coverage, range, horizon)] %>%
    dcast(formula = ... ~ range, value.var = "coverage") %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case"))

  # compute scores and merge with coverage values
  df <- eval_forecasts(
    data, 
    summarise_by = c("model", "horizon", "target_type"), 
    compute_relative_skill = FALSE,
    sd = TRUE) %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case")) %>%
    merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

  # round digits
  df <- df[, lapply(.SD, FUN = function(x) {
    if (is.numeric(x)) {
      return(round(signif(x, 3), 2))
    } else {
      return(x)
    }
  })]

  percentage_cols <- c("interval_score", 
                       "interval_score_sd", "sharpness", "underprediction", 
                       "overprediction", "aem")

  # add percentage values in brackets
  df <- df %>%
    dplyr::group_by(horizon, target_type) %>%
    dplyr::mutate_at(percentage_cols, 
                     ~ paste0(., " (", 
                              round(./.[model %in% c("Hub-ensemble", 
                                                     "Hub-ensemble-mean")], 2), 
                              ")"))

  setDT(df)

  # delete and rename a few columns
  setnames(df, 
           old = c("interval_score_sd", "sharpness"), 
           new = c("WIS - sd", "dispersion"))
  del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                          !(grepl("interval", names(df)))]
  del_cols <- c(del_cols, "coverage_deviation")
  df[, (del_cols) := NULL]

  # create and save tables for 1-2 weeks ahead and 3-4 weeks ahead
  table <- score_table(df[horizon <= 2]) 
  saveRDS(object = table, 
          file = here("analysis", "plots", paste0(plotname, "_", 2, "_ahead.rds")))

  table <- score_table(df[horizon > 2]) 
  saveRDS(object = table, 
          file = here("analysis", "plots", paste0(plotname, "_", 4, "_ahead.rds")))
}

# create tables
# regular models
filtered_data[model %in% regular_models] %>%
  make_table()

# regular models with later period as sensitivity analysis in SI
filtered_data[model %in% regular_models] %>%
  filter(!(forecast_date %in% as.character(forecast_dates$death_not_scored))) %>%
  make_table(plotname = "table-late-period")

# table for mean ensembles for SI
filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_mean-ensemble_scores")

# table for median ensembles for SI
filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_median-ensemble_scores")
```

# Plot with Visualsiation of daily data

```{r}

# reformat daily truth data
df <- copy(dailytruth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
][, target_end_date := as.Date(target_end_date)
]

df[weekdays(target_end_date) == "Sunday", sundays := target_end_date]

# reformat weekly truth data
weeklytruth <- copy(truth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
]

# create plot
df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_col(fill = "lightsteelblue", color = "white") + 
  geom_line(data = weeklytruth, aes(y = true_value / 7)) + 
  geom_point(data = weeklytruth, aes(y = true_value / 7), size = 0.3) +
  # geom_vline(aes(xintercept = sundays), linetype = "dashed", alpha = 0.2) + 
  facet_wrap( ~ target_location, scales = "free", ncol = 1) + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                          max(forecast_dates$full_hub_period + 21))) + 
  theme_minimal()

ggsave(here("analysis", "plots", "daily_truth.png"), width = 10, height = 8)
```

# Number of member models in the Hub ensemble

```{r}
ens <- ensemble_members %>%
  melt(id.vars = c("Date", "model"))

epinow2_included <- ens[model == "epiforecasts-EpiNow2", "epinow" := value]
crowd_included <- ens[model == "epiforecasts-EpiExpert", "crowd" := value]

targetvars <- data.frame(
  variable = c("GM_inc_case", "GM_inc_death", "PL_inc_case", "PL_inc_death"), 
  target = c("Cases - Germany", "Deaths - Germany", "Cases - Poland", "Deaths - Poland")
)

ens <- ens[, .(value = sum(value), 
             epinow = sum(epinow, na.rm = TRUE), 
             crowd = sum(crowd, na.rm = TRUE)), by = c("variable", "Date")] %>%
  merge(targetvars)

# number of submissions from crowd and epinow2
ens[, .(crowd = sum(crowd) / 4, 
        epinow = sum(epinow) / 4)]

# how many models included in the ensemble? 
ens %>%
  summary()

p <- ens %>%
  ggplot(aes(y = value, x = Date)) + 
  facet_wrap(~ target) + 
  # geom_col(aes(fill = variable), position = position_dodge()) + 
  geom_point(size = 0.7) + 
  geom_line() + 
  scale_color_manual(values = colcases) + 
  theme_minimal() + 
  labs(y = "Number of ensemble members", x = "Date")

ggsave(here("analysis", "plots", "ensemble-members.png"), 
       plot = p)
```

# Plot with distribution of scores

```{r}
# score forecasts
scores <- eval_forecasts(
  filtered_data[(model %in% regular_models)], 
  summarise_by = c("model", "target_type", "location_name", "horizon", 
                   "forecast_date", "target_phase",
                   "location_target", "classification"),
  compute_relative_skill = FALSE
)

# reformat column
scores[, target_type := paste0(str_to_title(target_type), "s overall")]

# display standard deviation for different models
scores[, .(sd = sd(interval_score)), by = c("model", "target_type")]

# helper function to create distribution plot
distribution_plot_wis <- function(df, x = "interval_score", 
                                  ranks = FALSE) {

  xlabel = "Weighted interval score"
  dotsize = 1

  if (ranks) {
    df <- df %>%
      group_by(location_name, target_type, horizon, forecast_date) %>%
      arrange(location_name, target_type, horizon, forecast_date, interval_score) %>%
      mutate(!!x := rank(get(x))) %>%
      ungroup() 
    xlabel <- "Rank in terms of WIS"
    dotsize = 0.1

  }

  df %>%
    rename(Model = model) %>%
    ggplot(aes_string(y = "Model", x = x, 
                      group = "Model", fill = "Model")) + 
    ggdist::stat_slab(alpha = 0.3, scale = 0.8, normalize = "panels") +
    ggdist::stat_dots(aes(color = Model),
                      dotsize = dotsize) + 
    ggdist::stat_pointinterval(point_interval = mean_qi, size = 3, 
                               shape = 16, fill = "black", .width = NA) + 
    ggdist::stat_pointinterval(point_interval = median_qi, size = 3, 
                               shape = 15, fill = "black", .width = NA) + 
        facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) +
    scale_fill_manual(values = coldeaths) + 
    scale_color_manual(values = coldeaths) + 
    theme(legend.position = "bottom") + 
    labs(y = "", x = xlabel) + 
  scale_x_continuous(labels = scale_fn) + 
    theme_minimal()
}

# create plots for the distributions of WIS
for (i in 1:4) {
  # plot for locations combined
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score") +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

  # plot for locations separate
  p_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

  # combine plots and save
  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here("analysis", "plots", 
              paste0("distribution_scores_wis-", i, ".tiff")), 
         plot = pl_wis, height = 6, width = 9, device = "tiff", dpi = 300)
}

# create plots for the distributions of WIS rank
for (i in 1:4) {
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score", 
                                          ranks = TRUE) +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

  p_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                 ranks = TRUE) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here("analysis", "plots", 
              paste0("distribution_scores_wis-", i, "-ranks.png")), 
         plot = pl_wis, height = 6, width = 9)

}
```

# Number of expert forecasters

```{r num-forecasters0, results = 'asis'}
# number of experts and non-experts --------------------------------------------
crowdforecast_data %>%
  select(model, expert) %>%
  unique() %>%
  summarise(experts = sum(expert), 
            nonexpert = sum(!expert))
```

# Number of forecasts over time

```{r num-forecasters, results = 'asis'}
# filter out ensembles and model-based forecasts
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                               "EpiNow2_secondary", 
                               "EpiExpert-ensemble", 
                               "EpiNow2")), 
                .(`number of forecasters` = length(unique(model))), , 
                by = c("forecast_date", "location_name", "target_type")
][order(forecast_date)][
  !is.na(forecast_date)]

# display summary for the number of available forecasters
dt[, .(sd = sd(`number of forecasters`), 
       mean = mean(`number of forecasters`), 
       min  = min(`number of forecasters`), 
       max = max(`number of forecasters`), 
       median = median(`number of forecasters`))]

# display summary for the number of available forecasters by location and target
dt[, .(sd = sd(`number of forecasters`), 
       mean = mean(`number of forecasters`), 
       min  = min(`number of forecasters`), 
       max = max(`number of forecasters`), 
       median = median(`number of forecasters`)), 
   by = c("location_name", "target_type")] 
```

```{r num-forecasters2, results = 'asis'}
# filter out ensembles and model-based forecasts
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                                        "EpiNow2_secondary", 
                                        "EpiExpert-ensemble", 
                                        "EpiNow2")), 
                         .(`number of forecasters` = length(unique(model))), , 
                         by = c("forecast_date", "location_name", "target_type")
][order(forecast_date)][
  !is.na(forecast_date)]

dt[, location_target := paste0(str_to_title(target_type), 
                               "s", " in ", location_name)]

# plot number of forecasters over time
dt %>%
  ggplot(aes(y = `number of forecasters`, x = forecast_date)) + 
  geom_line() + 
  facet_wrap(~ location_target) + 
  theme_minimal() + 
  labs(y = "Number of forecasters", x = "Date")

ggsave(filename = here("analysis", "plots", "number-forecasters.png"))
```

# Number of submissions per forecaster

```{r, results = 'asis'}
dt <- crowdforecast_data[!(model %in% c("Crowd-Rt-Forecast",
                                        "EpiNow2_secondary", 
                                        "EpiExpert-ensemble", 
                                        "EpiNow2")) & forecast_date %in% forecast_dates$unfiltered]

num_fc <- dt[, .(n_forecasts = length(unique(forecast_date))),
             by = c("model")][order(n_forecasts)]

median(num_fc$n_forecasts)
mean(num_fc$n_forecasts)

num_fc %>%
  ggplot(aes(x = n_forecasts)) + 
  geom_bar() + 
  theme_minimal() + 
  labs(x = "Number of available submissions", 
       y = "Number of forecasters")

ggsave("plots/crowdforecast_regularity.png")

```

Create comparison of crowd forecasts and baseline model - 2 Week ahead forecasts

```{r}

crowdforecasts <- covid.german.forecasts::prediction_data %>% 
  filter(model == "Crowd forecast")

qs <- c(0.05, 0.25, 0.5, 0.75, 0.95)

baseline_forecasts <- truth_data %>%
  mutate(log_true_value = log(pmax(true_value, 1)),
         difference = c(NA, diff(log_true_value)), 
         target_end_date = as.Date(target_end_date))

calc_baseline <- function(baseline_forecasts) {
  baseline_forecasts %>%
    mutate(id = 1:n()) %>%
    rowwise() %>%
    mutate(sd = sd(baseline_forecasts$difference[pmax(0, id - (4:1))]), 
           quantile = list(qs)) %>%
    tidyr::unnest(cols = quantile) %>%
    mutate(prediction = exp(qnorm(quantile, mean = log(true_value), sd = sd)), 
           model = "Baseline", 
           forecast_date = target_end_date + 2) %>%
    unique()
}

baseline_forecasts <- 
  rbind(
    calc_baseline(filter(baseline_forecasts, location_name == "Germany" & target_type == "case")), 
    calc_baseline(filter(baseline_forecasts, location_name == "Germany" & target_type == "death")), 
    calc_baseline(filter(baseline_forecasts, location_name == "Poland" & target_type == "case")), 
    calc_baseline(filter(baseline_forecasts, location_name == "Poland" & target_type == "death"))
  )

plot_df <- prediction_data %>%
  filter(model == "Crowd forecast", 
         quantile %in% qs, 
         grepl("2", target)) %>%
  unique() %>%
  rbind(baseline_forecasts, fill = TRUE) %>%
  filter(forecast_date %in% forecast_dates$unfiltered) %>%
  pivot_wider(names_from = quantile, values_from = prediction) %>%
  rename(Model = model) %>%
  mutate(target_type = ifelse(target_type == "case", "Case", "Death"))

plot_df %>%
  ggplot(aes(x = forecast_date, colour = NULL, fill = Model)) +
  geom_line(aes(y = `0.5`)) +
  facet_wrap(location_name ~ target_type, scales = "free_y") +
  geom_ribbon(aes(ymin = `0.05`, ymax = `0.95`), alpha = 0.2) +
  geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`), alpha = 0.5) +
  theme_minimal() +
  scale_y_continuous(labels = scale_fn) +
  labs(y = "Prediction intervals", x = "Forecast date") +
  theme(legend.position = "bottom")

ggsave(here("analysis", "plots", "comparison-forecast-interals.png"), 
       height = 6, width = 9)

```

Analysis of how data changed over time

```{r}
current_data <- 
  rbind(
    fread(here("data-raw", "daily-incidence-cases.csv")) %>%
      mutate(target_type = "Case"), 
    fread(here("data-raw", "daily-incidence-deaths.csv")) %>%
      mutate(target_type = "Death")
  ) %>%
  filter(location_name %in% c("Germany", "Poland")) %>%
  select(region = location_name, date, value, target_type) %>%
  unique()

original_cases <- original_deaths <- list()
for (target_date in as.character(forecast_dates$unfiltered)) {
  # Get cases  ---------------------------------------------------------------
  original_cases[[target_date]] <- 
    fread(file.path("rt-forecast", "data", "summary", 
                    "cases", target_date, "reported_cases.csv")) %>%
    mutate(fit_date = target_date)

  original_deaths[[target_date]] <- 
    fread(file.path("rt-forecast", "data", "summary", 
                    "deaths", target_date, "reported_cases.csv")) %>%
    mutate(fit_date = target_date)
}
original_cases <- rbindlist(original_cases) %>%
  mutate(target_type = "Case") 
original_deaths <- rbindlist(original_deaths) %>%
  mutate(target_type = "Death")

original <- 
  rbind(original_cases, original_deaths) %>%
  filter(region %in% c("Germany", "Poland")) %>%
  unique() %>%
  rename(value_original = confirm)

dailycomparison <- full_join(current_data, original) %>%
  filter(!is.na(value_original)) %>%
  mutate(difference = value - value_original, 
         rel_difference = difference / value_original)

dailycomparison%>%
  filter(date >= min(forecast_dates$unfiltered -2), 
         date <= max(forecast_dates$unfiltered)) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = difference)) +
  theme_minimal() +
  facet_wrap(region ~ target_type, scales = "free_y") +
  labs(y = "Absolute change in data", x = "Date")

ggsave(here("analysis", "plots", "daily-data-updates.png"), 
       height = 6, width = 9)

# weekly data
weeklycomparison <- dailycomparison %>%
  group_by(date, region, target_type) %>%
  # only take the oldest one for comparison if an update has taken place
  slice_head(n = 1) %>%
  ungroup() %>%
  select(-fit_date) %>%
  unique() %>%
  mutate(week = grates::as_yearweek(as.character(date), firstday = 7)) %>%
  group_by(week, target_type, region) %>%
  summarise(value = sum(value), 
            date = max(date),
            value_original = sum(value_original)) %>%
  mutate(difference = value - value_original, 
         rel_difference = difference / value_original)

weeklycomparison %>%
  filter(date >= min(forecast_dates$unfiltered -2), 
         date <= max(forecast_dates$unfiltered)) %>%
  ggplot(aes(x = date, y = rel_difference)) +
  geom_line() +
  facet_wrap(region ~ target_type, scales = "fixed") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(y = "Relative change in weekly data", x = "Date")

ggsave(here("analysis", "plots", "weekly-data-updates.png"), 
       height = 6, width = 9)

```

Plot with differences hub ensemble vs. crowd forecasts

```{r}
scores <- eval_forecasts(
  filtered_data[(model %in% regular_models)], 
  summarise_by = c("model", "target_type", "location_name", "horizon", 
                   "forecast_date", "target_phase",
                   "location_target", "classification"),
  compute_relative_skill = FALSE
)

# reformat column
scores[, target_type := str_to_title(target_type)]

diff_plot <- function(scores) {
  p <- scores %>%
    select(target_type, difference, location_name, forecast_date) %>%
    ggplot(aes(x = difference)) +
    ggdist::stat_slab(alpha = 1, scale = 1, normalize = "panels") +
    ggdist::stat_dots(fill = "black", color = "black", dotsize = 0.5) + 
    scale_x_continuous(labels = scale_fn) +
    theme(axis.ticks.y = element_blank(), 
          axis.text.y = element_blank(), 
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = NULL) +
    lemon::scale_x_symmetric(labels = scale_fn) +
    theme_minimal()
p1 <- p +
  facet_wrap(~ target_type, nrow = 2, scales = "free")
p2 <- p +
  facet_wrap(location_name ~ target_type, scales = "free")

return(p1 + p2)
}

p <- scores[model %in% c("Crowd forecast", "Hub-ensemble")] %>%
  filter(horizon == 2) %>%
  select(target_type, model, horizon, interval_score, location_name, forecast_date) %>%
  pivot_wider(names_from = model, values_from = interval_score) %>%
  mutate(difference = `Crowd forecast` - `Hub-ensemble`) %>%
  diff_plot()

ggsave(here("analysis", "plots", "difference-wis-crowd-ensemble.png"), 
        plot = p,
       height = 3.5, width = 9) +
  labs(y = "Density", x = "WIS difference Crowd forecast - Hub-ensemble")

p <- scores[model %in% c("Crowd forecast", "Renewal")] %>%
  filter(horizon == 2) %>%
  select(target_type, model, interval_score, location_name, forecast_date) %>%
  pivot_wider(names_from = model, values_from = interval_score) %>%
  mutate(difference = `Crowd forecast` - `Renewal`) %>%
  diff_plot() +
  labs(y = "Density", x = "WIS difference Crowd forecast - renewal")

ggsave(here("analysis", "plots", "difference-wis-crowd-renewal.png"), plot = p,
       height = 3.5, width = 9)

# significant results
significance <- function(scores, 
                         A = "Crowd forecast", B = "Hub-ensemble", 
                         horizons = 1:4) {
  scores[model %in% c(A, B)] %>%
    select(target_type, horizon, model, interval_score, location_name, forecast_date) %>%
    filter(horizon %in% horizons) %>%
    pivot_wider(names_from = model, values_from = interval_score) %>%
    mutate(difference = .data[[A]] - .data[[B]]) %>%
    pull(difference) %>%
    wilcox.test()
}

significance(scores, B = "Hub-ensemble")
significance(scores, B = "Renewal",  horizon = 4)
significance(scores, B = "Convolution")

```