### 2023-12-11 manuscript/S1-Acknowledgements.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\author{}
\date{\vspace{-2.5em}}

\begin{document}

We would like to thank (in a randomised order) the members of
Centre for the Mathematical Modelling of Infectious Diseases COVID-19
Working Group at the the London School of Hygiene \& Tropical Medicine:

Oliver Brady, Katharine Sherratt, Kaja Abbas, Kerry LM Wong, Charlie
Diamond, Katherine E. Atkins, Rein M G J Houben, Jiayao Lei, Rachel
Lowe, David Simons, Sophie R Meakin, Nicholas G. Davies, Timothy W
Russell, Kevin van Zandvoort, Quentin J Leclerc, Kathleen O'Reilly,
Stéphane Hué, Alicia Rosello, Emilie Finch, C Julian Villabona-Arenas,
Thibaut Jombart, W John Edmunds, Yalda Jafari, Jack Williams, Alicia
Showering, Damien C Tully, Jon C Emery, Carl A B Pearson, David Hodgson,
Frank G Sandmann, Petra Klepac, Adam J Kucharski, Graham Medley, Yang
Liu, Simon R Procter, Emily S Nightingale, William Waites, Rosanna C
Barnard, Joel Hellewell, Yung-Wai Desmond Chan, Fiona Yueqian Sun,
Hamish P Gibbs, Rosalind M Eggo, Lloyd A C Chapman, Stefan Flasche,
James W Rudge, Akira Endo, Naomi R Waterlow, Paul Mee, James D Munday,
Ciara V McCarthy, Mihaly Koltai, Amy Gimma, Christopher I Jarvis, Megan
Auzenbergs, Matthew Quaife, Fabienne Krauer, Samuel Clifford, Georgia R
Gore-Langton, Arminder K Deol, Kiesha Prem, Gwenan M Knight, Rachael
Pung, Anna M Foss.

\end{document}